---
title: "Multivariave distributions"
output: flexdashboard::flex_dashboard
runtime: shiny
---

<!-- -------------------------------------------------------------------------------- -->

<!-- Copyright 2019 Georgios Karagiannis -->

<!-- georgios.karagiannis@durham.ac.uk -->
<!-- Assistant Professor -->
<!-- Department of Mathematical Sciences, Durham University, Durham,  UK  -->

<!-- This file is part of Bayesian_Statistics (MATH3341/4031 Bayesian Statistics III/IV) -->
<!-- which is the material of the course (MATH3341/4031 Bayesian Statistics III/IV) -->
<!-- taught by Georgios P. Katagiannis in the Department of Mathematical Sciences   -->
<!-- in the University of Durham  in Michaelmas term in 2019 -->

<!-- Bayesian_Statistics is free software: you can redistribute it and/or modify -->
<!-- it under the terms of the GNU General Public License as published by -->
<!-- the Free Software Foundation version 3 of the License. -->

<!-- Bayesian_Statistics is distributed in the hope that it will be useful, -->
<!-- but WITHOUT ANY WARRANTY; without even the implied warranty of -->
<!-- MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the -->
<!-- GNU General Public License for more details. -->

<!-- You should have received a copy of the GNU General Public License -->
<!-- along with Bayesian_Statistics  If not, see <http://www.gnu.org/licenses/>. -->

<!-- -------------------------------------------------------------------------------- -->

<!-- Conjugate prior Applet -->
<!-- Copyright (C) 2018  Georgios Karagiannis -->
<!-- georgios.karagiannis@durham.ac.uk -->
<!-- Assistant Professor -->

<!-- This program is free software: you can redistribute it and/or modify -->
<!-- it under the terms of the GNU General Public License as published by -->
<!-- the Free Software Foundation, either version 3 of the License -->

<!-- This program is distributed in the hope that it will be useful, -->
<!-- but WITHOUT ANY WARRANTY; without even the implied warranty of -->
<!-- MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the -->
<!-- GNU General Public License for more details. -->

<!-- You should have received a copy of the GNU General Public License -->
<!-- along with this program. If not, see <http://www.gnu.org/licenses/>. -->

```{r}
library('mvtnorm')
```


# Normal

```{r}

G_x1_sl_def <- 0.0
G_x2_sl_def <- 0.0



G_x1_min_def <- -10.0
G_x1_max_def <- 10.0
G_x2_min_def <- -10.0
G_x2_max_def <- 10.0

#G_x1_def <- 0.0
#G_x2_def <- 0.0
G_m1_def <- 0.0 
G_m2_def <-  0.0
G_s1_def <- 1.0
G_s2_def <- 1.0
G_r_def <- 0.0

theta_def <- 45
phi_def <- 45


G_PDF1 <- function(x1,m1,s1) {
  return (dnorm(x1, mean = m1, sd = s1, log = FALSE))
}

G_PDF2 <- function(x1,m1,s1) {
  return (dnorm(x1, mean = m1, sd = s1, log = FALSE))
}

G_PDF12 <- function(x1,x2,m1,m2,s1,s2,r) {
  mu <- m1+s1/s2*r*(x2-m2)
  sig <- sqrt((1-r^2)*s1^2)
  return (dnorm(x1, mean = mu, sd = sig, log = FALSE))
}

G_PDF21 <- function(x2,x1,m1,m2,s1,s2,r) {
  mu <- m2+s2/s1*r*(x1-m1)
  sig <- sqrt((1-r^2)*s2^2)
  return (dnorm(x2, mean = mu, sd = sig, log = FALSE))
}

G_PDF <- function(x1,x2,m1,m2,s1,s2,r) {
  
  # x <- x1
  # y <- x2
  # mu1 <- m1
  # mu2 <- m2
  # sig1 <- s1
  # sig2 <- s2
  # rho <-r
  # 
  # term1 <- 1 / (2 * pi * sig1 * sig2 * sqrt(1 - rho^2))
  # term2 <- (x - mu1)^2 / sig1^2
  # term3 <- -(2 * rho * (x - mu1)*(y - mu2))/(sig1 * sig2)
  # term4 <- (y - mu2)^2 / sig2^2
  # z <- term2 + term3 + term4
  # term5 <- term1 * exp((-z / (2 *(1 - rho^2))))
  # return (term5)
  
  return( G_PDF1(x1,m1,s1)*G_PDF21(x2,x1,m1,m2,s1,s2,r) )
}

G_plot<- function(x1_sl=G_x1_sl_def,
                  x2_sl=G_x2_sl_def,
                  m1=G_m1_def,
                  m2=G_m2_def,
                  s1=G_s1_def,
                  s2=G_s2_def,
                  r=G_r_def,
                  x1_min=G_x1_min_def, 
                  x1_max=G_x1_max_def,
                  x2_min=G_x2_min_def, 
                  x2_max=G_x2_max_def,
                  theta = theta_def,
                  phi = phi_def
                  ) {


# x1_sl=G_x1_sl_def
# x2_sl=G_x2_sl_def
# m1=G_m1_def
# m2=G_m2_def
# s1=G_s1_def
# s2=G_s2_def
# r=G_r_def
# x1_min=G_x1_min_def
# x1_max=G_x1_max_def
# x2_min=G_x2_min_def
# x2_max=G_x2_max_def
# theta = theta_def
# phi = phi_def
  
  
  fntsz <- 1.5
  
  n_grid <- 50
  x1_grid <- seq(from = x1_min, to = x1_max, length.out=n_grid)
  x2_grid <- seq(from = x2_min, to = x2_max, length.out=n_grid)
  
  
  aa <-matrix (rep(5,times=2),nrow=1,ncol=2)
  bb <-matrix (rep(3,times=2),nrow=1,ncol=2)
  cc <-matrix (rep(2,times=4),nrow=2,ncol=2)
  dd <-matrix (rep(1,times=4),nrow=2,ncol=2)
  ee <-matrix (rep(6,times=2),nrow=2,ncol=1)
  ff <-matrix (rep(4,times=2),nrow=2,ncol=1)
  eeff <-cbind(ee,ff)
  AA <- rbind(aa,bb,cc)
  BB <- rbind(dd,eeff)
  AAA <- cbind(BB,AA)
  layout( AAA )
  rm(aa,bb,cc,ee,ff,eeff,AA,BB,AAA)
  

# PERSPECTIVE...
  
  PDF_local <- function(x1,x2,
                          m1_=m1,
                          m2_=m2,
                          s1_=s1,
                          s2_=s2,
                          r_=r) {
    G_PDF(x1,x2,m1_,m2_,s1_,s2_,r_)
  }
  
  z_grid <- outer(x1_grid, x2_grid, PDF_local)
  
  persp(x = x1_grid,
    y = x2_grid,
    z = z_grid, 
    xlim = c(x1_min,x1_max), 
    ylim = c(x2_min,x2_max), 
    xlab = expression('x1'),
    ylab = expression('x2'),
    zlab = "PDF",
    #main = "Gaussian",
    phi = phi, 
    theta = theta,
    cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz
  )
 
#COUNTORPLOT..
  
  contour(x = x1_grid,
    y = x2_grid,
    z = z_grid, 
    nlevels = 20,
    xlim = c(x1_min,x1_max), 
    ylim = c(x2_min,x2_max), 
    xlab = expression('x1'),
    ylab = expression('x2'),
    zlab = "PDF",
    #main = "Gaussian",
    cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz,
    method = "edge"
  )
  abline(v=x1_sl, col='blue')
  abline(h=x2_sl, col='green')
  legend('topright',
       bg="transparent",
       c('joint', 'x1', 'x2'),
       lty=c(1,1),
       lwd=c(2.5,2.5),
       col=c('black','blue','green')) 
  
#MARGINAL X1
  
    PDF_local <- function(x1,
                          m1_=m1,
                          s1_=s1) {
          G_PDF1(x1,m1_,s1_)
        }
  
  z_grid <- apply(matrix(x1_grid,ncol=1), 1, PDF_local)
  
  plot(x = x1_grid,
       y = z_grid,
       type='l',
       main="Marg. X1",
       xlab="",
       ylab='',
       xlim=c(x1_min,x1_max),
       cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz)
  
#MARGINAL X2
  
    PDF_local <- function(x2,
                          m2_=m2,
                          s2_=s2) {
          G_PDF2(x2,m2_,s2_)
        }
  
  z_grid <- apply(matrix(x2_grid,ncol=1), 1, PDF_local)
  
  plot(y = x2_grid,
       x = z_grid,
       type='l',
       main="Marg. X2",
       ylab="",
       xlab="",
       ylim=c(x2_min,x2_max),
       xlim=rev(range(z_grid)),
       cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz)
  
#CONDITIONAL X1|X2=x2
  
    PDF_local <- function(x1, 
                            x2_=x2_sl,
                          m1_=m1,
                          m2_=m2,
                          s1_=s1,
                          s2_=s2,
                          r_=r) {
          return(G_PDF12(x1,x2_,m1_,m2_,s1_,s2_,r_))
        }
  
  z_grid <- apply(matrix(x1_grid,ncol=1), 1, PDF_local)
  
  plot(x = x1_grid,
       y = z_grid,
       type='l',
       main="Cond. X1|X2=x2",
       xlab="",
       ylab='',
       xlim=c(x1_min,x1_max),
       col=c('green'),
       cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz)

  #CONDITIONAL X2|X1=x1
  
    PDF_local <- function(x2, 
                            x1_=x1_sl,
                          m1_=m1,
                          m2_=m2,
                          s1_=s1,
                          s2_=s2,
                          r_=r) {
          G_PDF21(x2,x1_,m1_,m2_,s1_,s2_,r_)
        }
  
  z_grid <- apply(matrix(x2_grid,ncol=1), 1, PDF_local)
  
  plot(y = x2_grid,
       x = z_grid,
       type='l',
       main="Cond. X2|X1=x1",
       ylab="",
       xlab='',
       ylim=c(x2_min,x2_max),
       xlim=rev(range(z_grid)),
       col=c('blue'),
       cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz)
    
}



```



Inputs {.sidebar} 
-----------------------



***The distribution***  

\begin{align*}
X|\mu, \Sigma \sim \text{N}_d(\mu, \Sigma) 
\end{align*}




***Distr. parameters***

```{r,echo=FALSE}
sliderInput("G_m1",
                 "$$\\mu_1:$$",
                  min = -5.0,
                  max = 5.0,
                  step = 0.05,
                  value = 0.0)
sliderInput("G_m2",
                 "$$\\mu_2:$$",
                  min = -5.0,
                  max = 5.0,
                  step = 0.05,
                  value = 0.0)
sliderInput("G_sig1",
                 "$$\\sigma_1:$$",
                  min = 0.001,
                  max = 3.0,
                  step = 0.05,
                  value = 1.0)
sliderInput("G_sig2",
                 "$$\\sigma_2:$$",
                  min = 0.001,
                  max = 3.0,
                  step = 0.05,
                  value = 1.0)
sliderInput("G_rho",
                 "$$\\rho:$$",
                  min = -1.0,
                  max = 1.0,
                  step = 0.01,
                  value = 0.9)
```

***Slices***

```{r,echo=FALSE}
sliderInput("G_x1",
                 "$$x_1:$$",
                  min = -5.0,
                  max = 5.0,
                  step = 0.05,
                  value = 0.0)
sliderInput("G_x2",
                 "$$x_2:$$",
                    min = -5.0,
                  max = 5.0,
                  step = 0.05,
                  value = 0.0)
```

***View***

```{r,echo=FALSE}
sliderInput("G_theta",
                 "$$\\theta:$$",
                  min = 1.0,
                  max = 90.0,
                  step = 1,
                  value = 45.0)
sliderInput("G_phi",
                 "$$\\phi:$$",
                    min = 1.0,
                  max = 90,
                  step = 1,
                  value = 45.0)
    
```

Column {.tabset}
-----------------------

### ***Motivation: A CLT***  

***A CLT, ...***

Motivation comes fromt he following observation:

Assume IID r.v. $\{X_i\}$ such that 
\[
X_i\sim \text{d}\pi(\cdot)
\] 
with $\mu=\text{E}(X_i)<\infty$ and $\Sigma = \text{Var}(X_i)>0$, for $i=1,...,n$.

Define 
\[
Z_n = \sqrt{n}(\bar{X}_n -\mu), \ \ \ \bar{X}_n=\frac{1}{n}\sum_{i=1}^{n}X_i
\] 

Then, for any $A\subset\mathbb{R}^d$  
\[
\lim_{n \rightarrow \infty} \Pr(Z_n\in A) 
=
\int_A (2 \pi)^{-\frac{d}{2}} \det(\Sigma)^{-\frac{1}{2}} \exp(-\frac{1}{2}t^{\text{T}}\Sigma^{-1}t) \text{d}t
\] 

where the results applies for non IID cases under certain conditions...

***More thinking...***

+ ooooh...  
++ look at that dude in the integral can be considered as a PDF of a quite useful distribution (due to the CLT)...
++ let's define the distribution properly, standardise and study furtehr this new distribution...


### ***Description ($d$-D case) *** 

***Introdunctions***  

Consider a r.v. $X\in\mathbb{R}^{d}$ follows a Normal (Gaussian) distribution
with location parameter $\mu\in\mathbb{R}^{d}$ and scale parameter
$\Sigma>0$ with dim $d\times d$

***Notation***  

\[
X\sim\text{N}_{d}(\mu,\Sigma)
\]

***Probability Density Function***

\begin{align*}
f_{X}(x|\mu,\Sigma)= & (2\pi)^{-\frac{d}{2}}\det(\Sigma)^{-\frac{1}{2}}\exp(-\frac{1}{2}(t-\mu)^{\text{T}}\Sigma^{-1}(t-\mu))
\end{align*}

***Cummulative Density Function***

\begin{align*}
F_{X}(x_{d},...x_{d}|\mu,\Sigma)=&\int_{-\infty}^{x_{1}}\cdots  \int_{-\infty}^{x_{2}}(2\pi)^{-\frac{d}{2}}\det(\Sigma)^{-\frac{1}{2}}\exp(-\frac{1}{2}(t-\mu)^{\text{T}}\Sigma^{-1}(t-\mu))\text{d}t_1\cdots\text{d}t_d
\end{align*}

***Mean, Mode, Median***

\[
\text{E(X)} = \mu
\]

well ... it is symentric, so compute the rest by yourself...

***Variance***

\[
\text{Var(X)} = \Sigma
\]

***Moment generating function***

\[
\text{E}(\exp(t^{\text{T}}X))=\exp(\mu^{\text{T}}t+\frac{1}{2}t^{\text{T}}\Sigma t),\,t\in\mathbb{R}^{d}
\]

***Characteristic function***

\[
\text{E}(\exp(it^{\text{T}}X))=\exp(i\mu^{\text{T}}t-\frac{1}{2}t^{\text{T}}\Sigma t),\,t\in\mathbb{R}^{d}
\]

***Partial distributions***

If $X=(X_{1},X_{2})$ with $X\in\mathbb{R}^{d}$, $X_{1}\in\mathbb{R}^{d_{1}}$,
$X_{2}\in\mathbb{R}^{d_{2}}$, and 
\[
X\sim\text{N}_{d}(\mu,\Sigma)
\]
 where 
\[
\mu=\begin{bmatrix}\mu_{1}\\
\mu_{2}
\end{bmatrix}\quad;\quad\Sigma=\begin{bmatrix}\Sigma_{11} & \Sigma_{21}^{\text{T}}\\
\Sigma_{21} & \Sigma_{22}
\end{bmatrix}
\]
then  

+ **Marginal distributions:**
\[
X_{1}\sim\text{N}_{d_{1}}(\mu_{1},\Sigma_{11})
\]
and 
\[
X_{2}\sim\text{N}_{d_{2}}(\mu_{2},\Sigma_{22})
\]  

+ **Conditional distributions:**
\[
X_{1}|X_{2}=x_{2}\sim\text{N}_{d_{1}}(\mu_{1|2},\Sigma_{1|2})
\]
 where $\Sigma_{1|2}=\Sigma_{11}-\Sigma_{21}^{\text{T}}\Sigma_{11}^{-1}\Sigma_{21}$,
and $\mu_{1|2}=\mu_{1}-\Sigma_{21}^{\text{T}}\Sigma_{22}^{-1}(\mu_{2}-x_{2})$   
and   
\[
X_{2}|X_{1}=x_{1}\sim\text{N}_{d_{2}}(\mu_{2|1},\Sigma_{2|1})
\]
 where $\Sigma_{2|1}=\Sigma_{22}-\Sigma_{21}\Sigma_{22}^{-1}\Sigma_{21}^{\text{T}}$,
and $\mu_{2|1}=\mu_{2}-\Sigma_{21}\Sigma_{11}^{-1}(\mu_{1}-x_{1})$




***Properties***

+ If $A\in\mathbb{R}^{q\times q}$, and $b\in\mathbb{R}^{q}$ are constants
then
\begin{align*}
X & \sim\text{N}_{d}(\mu,\Sigma)\;\Longleftrightarrow AX+b\sim\text{N}_{q}(A\mu+b,A\Sigma A^{\text{T}})
\end{align*}
+ If $L$ the Cholesky lower triangular of $\Sigma>0$ (i.e., $\Sigma=LL^{\text{T}}$
and constants) then
\[
X\sim\text{N}_{d}(\mu,\Sigma)\Longleftrightarrow L^{-1}(X-\mu)\sim\text{N}_{d}(0,I)
\]
and
\[
X\sim\text{N}_{d}(0,I)\Longleftrightarrow\mu+LX\sim\text{N}_{d}(\mu,\Sigma)
\]

***Related distributions***

+ **Student's t distribution:**  
if 
\[
X\sim\text{N}_{d}(0,\Sigma) \  ,\text{and} \  
u\sim\chi^2_{\nu},
\]
and 
$\mu\in\mathbb{R}^d$ 
then 
$Y=\mu+X(\frac{u}{\nu})^{-\frac{1}{2}}$ 
follows a 'Student's t distribution' with location parameter $\mu\in\mathbb{R}^d$, scale parameter $\Sigma$, and degrees of freedom $\nu$, i.e.   
\[
Y\sim\text{t}_d(\mu,\Sigma,\nu)
\]  
+ **Non-central $\chi^2$ distribution:**  
if 
\[
X\sim\text{N}_{d}(\mu,I)
\]
and then $Y=X^{\text{T}}QX$ where $Q$ is a symmetric projection of rank $r$ follows a 'Non-central $chi^2$ distribution' with non-centrality parameterr $\lambda=\mu^{\text{T}}Q\mu$ and degrees of freedom $\nu$, i.e. 
\[
Y\sim\chi^2_\nu(\lambda)
\]  


### ***Description ($2$-D case) *** 

***Introdunctions***  

Consider a r.v. $X\in\mathbb{R}^{2}$ follows a Normal (Gaussian) distribution
with location parameter $\mu$ and scale parameter
$\Sigma>0$ with dim $d\times d$

***Notation***  

\[
X\sim\text{N}_{2}(\mu,\Sigma)
\]

where

\[
\mu=\begin{bmatrix}\mu_{1}\\
\mu_{2}
\end{bmatrix}\quad;\quad\Sigma=\begin{bmatrix}\sigma_{11} & \sigma_{12}\\
\sigma_{12} & \sigma_{22}
\end{bmatrix}=\begin{bmatrix}\sigma_{1}^{2} & \rho\sigma_{1}\sigma_{2}\\
\rho\sigma_{1}\sigma_{2} & \sigma_{2}^{2}
\end{bmatrix}
\]

***Probability Density Function***

\begin{align*}
f_{X}(x|\mu,\Sigma)= & \frac{1}{2\pi\sigma_{1}\sigma_{2}\sqrt{1-\rho^{2}}}\exp(-\frac{1}{2(1-\rho^{2})}(\frac{(x_{1}-\mu_{1})^{2}}{\sigma_{1}^{2}}\\
 & \qquad\qquad+\frac{(x_{2}-\mu_{2})^{2}}{\sigma_{2}^{2}}-\frac{2\rho(x_{1}-\mu_{1})(x_{2}-\mu_{2})}{\sigma_{1}\sigma_{2}}))
\end{align*}

***Cummulative Density Function***

\begin{align*}
F_{X}(x_{d},...x_{d}|\mu,\Sigma)=&\int_{-\infty}^{x_{1}}  \int_{-\infty}^{x_{2}}\frac{1}{2\pi\sigma_{1}\sigma_{2}\sqrt{1-\rho^{2}}}\exp(-\frac{1}{2(1-\rho^{2})}(\frac{(x_{1}-\mu_{1})^{2}}{\sigma_{1}^{2}}\\
 & \qquad\qquad+\frac{(x_{2}-\mu_{2})^{2}}{\sigma_{2}^{2}}-\frac{2\rho(x_{1}-\mu_{1})(x_{2}-\mu_{2})}{\sigma_{1}\sigma_{2}}))\text{d}t_1\text{d}t_2
\end{align*}

***Mean, Mode, Median***

\[
\text{E(X)} = \begin{bmatrix}\mu_{1}\\
\mu_{2}
\end{bmatrix}
\]

well ... it is symentric, so compute the rest by yourself...

***Variance***

\[
\text{Var(X)} = \begin{bmatrix}\sigma_{1}^{2} & \rho\sigma_{1}\sigma_{2}\\
\rho\sigma_{1}\sigma_{2} & \sigma_{2}^{2}
\end{bmatrix}
\]

***Moment generating function***

\[
\text{E}(\exp(t^{\text{T}}X))=\exp(i(t_{1}\mu_{1}+t_{2}\mu_{2})+\frac{1}{2}(\sigma_{1}^{2}t_{1}^{2}+\sigma_{2}^{2}t_{2}^{2}+2\rho\sigma_{1}\sigma_{2}t_{1}t_{2}))\]

***Characteristic function***

\[
\text{E}(\exp(it^{\text{T}}X))=\exp(i(t_{1}\mu_{1}+t_{2}\mu_{2})-\frac{1}{2}(\sigma_{1}^{2}t_{1}^{2}+\sigma_{2}^{2}t_{2}^{2}+2\rho\sigma_{1}\sigma_{2}t_{1}t_{2}))
\]

***Partial distributions***

If $X=(X_{1},X_{2})$ with $X\in\mathbb{R}^{2}$, $X_{1}\in\mathbb{R}$,
$X_{2}\in\mathbb{R}$, and 
\[
X\sim\text{N}_{2}(\mu,\Sigma)
\]
 where 
\[
\mu=\begin{bmatrix}\mu_{1}\\
\mu_{2}
\end{bmatrix}\quad;\quad\Sigma=\begin{bmatrix}\sigma_{1}^2 & \rho\sigma_{1}\sigma_{2}\\
\rho\sigma_{1}\sigma_{2} & \sigma_{2}^2
\end{bmatrix}
\]
then  

+ **Marginal distributions:**
\[
X_{1}\sim\text{N}(\mu_{1},\sigma_{1}^2)
\]  
and 
\[
X_{2}\sim\text{N}(\mu_{2},\sigma_{2}^2)
\]  

+ **Conditional distributions:**
\[
X_{1}|X_{2}=x_{2}\sim\text{N}(\mu_{1}+\frac{\sigma_{1}}{\sigma_{2}}\rho(x_{2}-\mu_{2}),\,(1-\rho^{2})\sigma_{1}^{2})
\]
and
\[
X_{2}|X_{1}=x_{1}\sim\text{N}(\mu_{2}+\frac{\sigma_{2}}{\sigma_{1}}\rho(x_{1}-\mu_{1}),\,(1-\rho^{2})\sigma_{2}^{2})
\]












### ***Demo***

```{r,echo=FALSE}
renderPlot({
    G_plot(
      x1_sl = input$G_x1, 
            x2_sl = input$G_x2,
            m1 = input$G_m1, 
            m2 = input$G_m2, 
            s1 = input$G_sig1, 
            s2 = input$G_sig2, 
            r = input$G_rho, 
            theta = input$G_theta,
            phi = input$G_phi)
  })

```




# Student's t


```{r}

T_x1_sl_def <- 0.0
T_x2_sl_def <- 0.0

T_x1_min_def <- -10.0
T_x1_max_def <- 10.0
T_x2_min_def <- -10.0
T_x2_max_def <- 10.0

#T_x1_def <- 0.0
#T_x2_def <- 0.0
T_m1_def <- 0.0 
T_m2_def <-  0.0
T_s1_def <- 1.0
T_s2_def <- 1.0
T_r_def <- 0.0

T_theta_def <- 45
T_phi_def <- 45

T_nu_def <- 1


dbvt <-function (x, m, S, v) {
  
  d <-length( x )
  
  L <- t(chol(S))
  
  logdetSig <- sum(2*log(diag(L)))
                   
  term1 <- lgamma(0.5*(v+d)) -lgamma(0.5*v)-0.5*d*log(pi) -logdetSig
  
  Q <- solve(L,(x-m))
  Q <- sum(Q*Q)
  
  term2 <- -0.5*(v+d)*log(1+Q) 
  
  pdf <- exp( term1 + term2 )

  return(pdf)
}


T_PDF1 <- function(x,m,s,nu) {
  pdf <- dbvt(x,m,s^2,nu)
  #pdf <-dmvt( rep(x1, 1), delta = rep(0, 1), sigma = diag(1), df = 1)
  return (pdf)
}

T_PDF2 <- function(x,m,s,nu) {
  pdf <- dbvt(x,m,s^2,nu)
  #pdf <-dmvt( rep(x1, 1), delta = rep(0, 1), sigma = diag(1), df = 1)
  return (pdf)
}

T_PDF12 <- function(x1,x2,m1,m2,s1,s2,r,nu) {
  mu <- m1+s1/s2*r*(x2-m2)
  sig2 <- (1-r^2)*s1^2
  sig2 <- (nu+ (x2-m2)^2/s2^2  )/(nu+1) *sig2
  pdf <- dbvt(x1,mu,sig2,1)
  return (pdf)
}

T_PDF21 <- function(x2,x1,m1,m2,s1,s2,r,nu) {
  mu <- m2+s2/s1*r*(x1-m1)
  sig2 <- ((1-r^2)*s2^2)
  sig2 <- (nu +(x1-m1)^2/s1^2  )/(nu+1)* sig2
  pdf <- dbvt(x2,mu,sig2,1)
  return (pdf)
}





T_PDF <- function(x1,x2,m1,m2,s1,s2,r,nu) {

  S <- matrix( c(s1*s1,r*s1*s2,r*s1*s2,s2*s2) , nrow=2)
  
  pdf <-  dbvt(c(x1,x2),c(m1,m2),S,nu)
  #pdf <-dmvt( c(x1, x2), delta = c(m1,m2), sigma = S, df = nu)
  
    return( pdf )
}

T_plot<- function(x1_sl=T_x1_sl_def,
                  x2_sl=T_x2_sl_def,
                  m1=T_m1_def,
                  m2=T_m2_def,
                  s1=T_s1_def,
                  s2=T_s2_def,
                  r=T_r_def,
                  x1_min=T_x1_min_def, 
                  x1_max=T_x1_max_def,
                  x2_min=T_x2_min_def, 
                  x2_max=T_x2_max_def,
                  theta = T_theta_def,
                  phi = T_phi_def,
                  nu = T_nu_def) {


 # x1_sl=T_x1_sl_def
 # x2_sl=T_x2_sl_def
 # m1=T_m1_def
 # m2=T_m2_def
 # s1=T_s1_def
 # s2=T_s2_def
 # r=T_r_def
 # x1_min=T_x1_min_def
 # x1_max=T_x1_max_def
 # x2_min=T_x2_min_def
 # x2_max=T_x2_max_def
 # theta = T_theta_def
 # phi = T_phi_def
 #  nu = T_nu_def
  
  fntsz <- 1.5
  
  n_grid <- 50
  x1_grid <- seq(from = x1_min, to = x1_max, length.out=n_grid)
  x2_grid <- seq(from = x2_min, to = x2_max, length.out=n_grid)
  
  
  aa <-matrix (rep(5,times=2),nrow=1,ncol=2)
  bb <-matrix (rep(3,times=2),nrow=1,ncol=2)
  cc <-matrix (rep(2,times=4),nrow=2,ncol=2)
  dd <-matrix (rep(1,times=4),nrow=2,ncol=2)
  ee <-matrix (rep(6,times=2),nrow=2,ncol=1)
  ff <-matrix (rep(4,times=2),nrow=2,ncol=1)
  eeff <-cbind(ee,ff)
  AA <- rbind(aa,bb,cc)
  BB <- rbind(dd,eeff)
  AAA <- cbind(BB,AA)
  layout( AAA )
  rm(aa,bb,cc,ee,ff,eeff,AA,BB,AAA)
  

# PERSPECTIVE...
  
  PDF_local <- function(x1,x2,
                          m1_=m1,
                          m2_=m2,
                          s1_=s1,
                          s2_=s2,
                          r_=r,
                        nu_=nu) {
    T_PDF(x1,x2,m1_,m2_,s1_,s2_,r_,nu_)
  }
  
  #z_grid <- outer(x1_grid, x2_grid, PDF_local)
  
  z_grid <- matrix(rep(NaN,n_grid*n_grid), nrow=n_grid)
  for(i in 1:n_grid)
    for (j in 1:n_grid)
      z_grid[i,j] <- PDF_local(x1_grid[i],x2_grid[j])
  
  persp(x = x1_grid,
    y = x2_grid,
    z = z_grid, 
    xlim = c(x1_min,x1_max), 
    ylim = c(x2_min,x2_max), 
    xlab = expression('x1'),
    ylab = expression('x2'),
    zlab = "PDF",
    #main = "Gaussian",
    phi = phi, 
    theta = theta,
    cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz
  )
 
#COUNTORPLOT..
  
  contour(x = x1_grid,
    y = x2_grid,
    z = z_grid, 
    nlevels = 20,
    xlim = c(x1_min,x1_max), 
    ylim = c(x2_min,x2_max), 
    xlab = expression('x1'),
    ylab = expression('x2'),
    zlab = "PDF",
    #main = "Gaussian",
    cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz,
    method = "edge"
  )
  abline(v=x1_sl, col='blue')
  abline(h=x2_sl, col='green')
  legend('topright',
       bg="transparent",
       c('joint', 'x1', 'x2'),
       lty=c(1,1),
       lwd=c(2.5,2.5),
       col=c('black','blue','green')) 
  
#MARGINAL X1
  
    PDF_local <- function(x1,
                          m1_=m1,
                          s1_=s1,
                          nu1_=nu) {
          T_PDF1(x1,m1_,s1_,nu1_)
        }
  
  z_grid <- apply(matrix(x1_grid,ncol=1), 1,PDF_local)
    
  plot(x = x1_grid,
       y = z_grid,
       type='l',
       main="Marg. X1",
       xlab="",
       ylab='',
       xlim=c(x1_min,x1_max),
       cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz)
  
#MARGINAL X2
  
    PDF_local <- function(x2,
                          m2_=m2,
                          s2_=s2,
                          nu2_=nu) {
          T_PDF2(x2,m2_,s2_,nu2_)
        }
  
  z_grid <- apply(matrix(x2_grid,ncol=1), 1, PDF_local)
  
  plot(y = x2_grid,
       x = z_grid,
       type='l',
       main="Marg. X2",
       ylab="",
       xlab="",
       ylim=c(x2_min,x2_max),
       xlim=rev(range(z_grid)),
       cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz)
  
#CONDITIONAL X1|X2=x2
  
    PDF_local <- function(x1, 
                            x2_=x2_sl,
                          m1_=m1,
                          m2_=m2,
                          s1_=s1,
                          s2_=s2,
                          r_=r,
                          nu1_=nu) {
          return(T_PDF12(x1,x2_,m1_,m2_,s1_,s2_,r_,nu1_))
        }
  
  z_grid <- apply(matrix(x1_grid,ncol=1), 1, PDF_local)
  
  plot(x = x1_grid,
       y = z_grid,
       type='l',
       main="Cond. X1|X2=x2",
       xlab="",
       ylab='',
       xlim=c(x1_min,x1_max),
       col=c('green'),
       cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz)

  #CONDITIONAL X2|X1=x1
  
    PDF_local <- function(x2, 
                            x1_=x1_sl,
                          m1_=m1,
                          m2_=m2,
                          s1_=s1,
                          s2_=s2,
                          r_=r,
                          nu_=nu) {
          T_PDF21(x2,x1_,m1_,m2_,s1_,s2_,r_,nu_)
        }
  
  z_grid <- apply(matrix(x2_grid,ncol=1), 1, PDF_local)
  
  plot(y = x2_grid,
       x = z_grid,
       type='l',
       main="Cond. X2|X1=x1",
       ylab="",
       xlab='',
       ylim=c(x2_min,x2_max),
       xlim=rev(range(z_grid)),
       col=c('blue'),
       cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz)
    
}

```






Inputs {.sidebar} 
-----------------------


***The distribution***  

\begin{align*}
X\sim\text{t}_{d}(\mu,\Sigma,\nu)
\end{align*}




***Distr. parameters***

```{r,echo=FALSE}
sliderInput("T_m1",
                 "$$\\mu_1:$$",
                  min = -5.0,
                  max = 5.0,
                  step = 0.05,
                  value = 0.0)
sliderInput("T_m2",
                 "$$\\mu_2:$$",
                  min = -5.0,
                  max = 5.0,
                  step = 0.05,
                  value = 0.0)
sliderInput("T_sig1",
                 "$$\\sigma_1:$$",
                  min = 0.001,
                  max = 3.0,
                  step = 0.05,
                  value = 1.0)
sliderInput("T_sig2",
                 "$$\\sigma_2:$$",
                  min = 0.001,
                  max = 3.0,
                  step = 0.05,
                  value = 1.0)
sliderInput("T_rho",
                 "$$\\rho:$$",
                  min = -1.0,
                  max = 1.0,
                  step = 0.01,
                  value = 0.9)
sliderInput("T_nu",
                 "$$\\nu:$$",
                  min = 1,
                  max = 10,
                  step = 1,
                  value = 1)
```

***Slices***

```{r,echo=FALSE}
sliderInput("T_x1",
                 "$$x_1:$$",
                  min = -5.0,
                  max = 5.0,
                  step = 0.05,
                  value = 0.0)
sliderInput("T_x2",
                 "$$x_2:$$",
                    min = -5.0,
                  max = 5.0,
                  step = 0.05,
                  value = 0.0)
```

***View***

```{r,echo=FALSE}
sliderInput("T_theta",
                 "$$\\theta:$$",
                  min = 1.0,
                  max = 90.0,
                  step = 1,
                  value = 45.0)
sliderInput("T_phi",
                 "$$\\phi:$$",
                    min = 1.0,
                  max = 90,
                  step = 1,
                  value = 45.0)
    
```



Column {.tabset}
-----------------------

### ***Motivation: Regression?***  

***See the transformation***

Motivation comes fromt he following observation:

if 
\[
X\sim\text{N}_{d}(0,\Sigma) \  ,\text{and} \  
u\sim\chi^2_{\nu},
\]
and 
$\mu\in\mathbb{R}^d$ 
then 
$Y=\mu+X(\frac{u}{\nu})^{-\frac{1}{2}}$ 
follows a 'Student's t distribution' with location parameter $\mu\in\mathbb{R}^d$, scale parameter $\Sigma$, and degrees of freedom $\nu$, i.e.   
\[
Y\sim\text{t}_d(\mu,\Sigma,\nu)
\]  

***More thinking...***

+ ooooh...  
++ It can be used in Bayesian linear models to compute the predictive distribution  
++ let's define the distribution properly, standardise and study furtehr this new distribution...


### ***Description ($d$-D case) *** 

***Introdunctions***  

Consider a r.v. $X\in\mathbb{R}^{d}$ follows a Student's t distribution
with location parameter $\mu\in\mathbb{R}^{d}$, scale parameter
$\Sigma>0$ with dim $d\times d$, and degrees of freedom $\nu$.

***Notation***  

\[
X\sim\text{t}_{d}(\mu,\Sigma,\nu)
\]

***Probability Density Function***

\begin{align*}
f_{X}(x|\mu,\Sigma)= & \frac{\Gamma(\frac{\nu+d}{2})}{\Gamma(\frac{\nu}{2})\nu^{\frac{d}{2}}\pi^{\frac{d}{2}}\det(\Sigma)}(1+\frac{1}{2}(t-\mu)^{\text{T}}\Sigma^{-1}(t-\mu))^{-\frac{\nu+d}{2}}
\end{align*}

***Cummulative Density Function***

\begin{align*}
F_{X}(x_{d},...x_{d}|\mu,\Sigma,\nu)=\int_{-\infty}^{x_{1}}\cdots & \int_{-\infty}^{x_{2}}\frac{\Gamma(\frac{\nu+d}{2})}{\Gamma(\frac{\nu}{2})\nu^{\frac{d}{2}}\pi^{\frac{d}{2}}\det(\Sigma)}(1+\frac{1}{2}(t-\mu)^{\text{T}}\Sigma^{-1}(t-\mu))^{-\frac{\nu+d}{2}}\text{d}t
\end{align*}

***Mean, Mode, Median***

\[
\text{E(X)} = \mu
\]

well ... it is symentric, so compute the rest by yourself...

***Variance***

\[
\text{Var(X)} = \begin{cases}
\frac{\nu}{\nu-2}\Sigma & ,\text{ if }\nu>2\\
\text{undefined} & ,\text{ else }
\end{cases}
\]

***Moment generating function***

XXX

***Characteristic function***

XXX

***Partial distributions***

If $X=(X_{1},X_{2})$ with $X\in\mathbb{R}^{d}$, $X_{1}\in\mathbb{R}^{d_{1}}$,
$X_{2}\in\mathbb{R}^{d_{2}}$, and 
\[
X\sim\text{t}_{d}(\mu,\Sigma,\nu)
\]
 where 
\[
\mu=\begin{bmatrix}\mu_{1}\\
\mu_{2}
\end{bmatrix}\quad;\quad\Sigma=\begin{bmatrix}\Sigma_{11} & \Sigma_{21}^{\text{T}}\\
\Sigma_{21} & \Sigma_{22}
\end{bmatrix}
\]
then  

+ **Marginal distributions:**
\[
X_{1}\sim\text{t}_{d_{1}}(\mu_{1},\Sigma_{11},\nu)
\]
and 
\[
X_{2}\sim\text{t}_{d_{2}}(\mu_{2},\Sigma_{22},\nu)
\]  

+ **Conditional distributions:**
\[
X_{1}|X_{2}=x_{2}\sim\text{t}_{d_{1}}(\mu_{1|2},\Sigma_{1|2},\nu_{1|2})
\]
with
\begin{align*}
\mu_{1|2}= & \mu_{1}-\Sigma_{21}\Sigma_{22}^{-1}(\mu_{2}-x_{2})\\
\Sigma_{1|2}= & \frac{\nu+(\mu_{2}-x_{2})^{\text{T}}\Sigma_{22}^{-1}(\mu_{2}-x_{2})}{\nu+d_{2}}\Sigma_{11|2}\\
\Sigma_{11|2}= & \Sigma_{11}-\Sigma_{21}\Sigma_{11}^{-1}\Sigma_{21}^{\text{T}}\\
\nu_{1|2}= & \nu+d_{2}
\end{align*}
and
\[
X_{2}|X_{1}=x_{1}\sim\text{t}_{d_{2}}(\mu_{2|1},\Sigma_{2|1},\nu_{2|1})
\]
with
\begin{align*}
\mu_{2|1}= & \mu_{2}-\Sigma_{21}\Sigma_{11}^{-1}(\mu_{1}-x_{1})\\
\Sigma_{2|1}= & \frac{\nu+(\mu_{1}-x_{1})\Sigma_{11}^{-1}(\mu_{1}-x_{1})}{\nu+d_{1}}\Sigma_{22|1}\\
\Sigma_{22|1}= & \Sigma_{22}-\Sigma_{21}^{\text{T}}\Sigma_{22}^{-1}\Sigma_{21}\\
\nu_{1|2}= & \nu+d_{1}
\end{align*}



***Properties***

+ If $A\in\mathbb{R}^{q\times q}$, and $b\in\mathbb{R}^{q}$ are constants
then
\begin{align*}
X & \sim\text{t}_{d}(\mu,\Sigma,\nu)\;\Longleftrightarrow AX+b\sim\text{t}_{d}(A\mu+b,A\Sigma A^{\text{T}},\nu)
\end{align*}
+ If $L$ the Cholesky lower triangular of $\Sigma>0$ (i.e., $\Sigma=LL^{\text{T}}$
and constants) then
\[
X\sim\text{t}_{d}(\mu,\Sigma,\nu)\Longleftrightarrow L^{-1}(X-\mu)\sim\text{t}_{d}(0,I,\nu)
\]
and
\[
X\sim\text{t}_{d}(0,I,\nu)\Longleftrightarrow\mu+LX\sim\text{t}_{d}(\mu,\Sigma,\nu)
\]

### Demo

```{r,echo=FALSE}
renderPlot({
    T_plot(
      x1_sl = input$T_x1, 
            x2_sl = input$T_x2,
            m1 = input$T_m1, 
            m2 = input$T_m2, 
            s1 = input$T_sig1, 
            s2 = input$T_sig2, 
            r = input$T_rho, 
            theta = input$T_theta,
            phi = input$T_phi,
      nu = input$T_nu)
  })

```
