<!-- -------------------------------------------------------------------------------- -->

<!-- Copyright 2019 Georgios Karagiannis -->

<!-- georgios.karagiannis@durham.ac.uk -->
<!-- Assistant Professor -->
<!-- Department of Mathematical Sciences, Durham University, Durham,  UK  -->

<!-- This file is part of Bayesian_Statistics (MATH3341/4031 Bayesian Statistics III/IV) -->
<!-- which is the material of the course (MATH3341/4031 Bayesian Statistics III/IV) -->
<!-- taught by Georgios P. Katagiannis in the Department of Mathematical Sciences   -->
<!-- in the University of Durham  in Michaelmas term in 2019 -->

<!-- Bayesian_Statistics is free software: you can redistribute it and/or modify -->
<!-- it under the terms of the GNU General Public License as published by -->
<!-- the Free Software Foundation version 3 of the License. -->

<!-- Bayesian_Statistics is distributed in the hope that it will be useful, -->
<!-- but WITHOUT ANY WARRANTY; without even the implied warranty of -->
<!-- MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the -->
<!-- GNU General Public License for more details. -->

<!-- You should have received a copy of the GNU General Public License -->
<!-- along with Bayesian_Statistics  If not, see <http://www.gnu.org/licenses/>. -->

<!-- -------------------------------------------------------------------------------- -->

---
title: "Bernoulli model with conjugate priors"
subtitle: "Case study: Space shuttle Challenger disaster"
author: "Georgios P. Karagiannis @ MATH3341/4031 Bayesian statistics III/IV (practical implementation)"
output:
  html_notebook: 
    number_sections: true
  word_document: default
  html_document:
    df_print: paged
    number_sections: true
  pdf_document: default
---
[Back to README](https://github.com/georgios-stats/Bayesian_Statistics/tree/master/ComputerPracticals#aim)

```{r, results="hide"}
rm(list=ls())
```


---

***Aim***

Students will become able to:  

+ produce Monte Carlo approximations of posterior quantities required for Bayesian analysis with the RJAGS R package  

+ implement Bayesian posterior analysis in R with RJAGS package  

Students are not required to learn by heart any of the concepts discussed

---

***Reference list***

*The material about RJAGS package is not examinable material, but it is provided for the interested student. It contains references that students can follow if they want to further explore the concepts introdced.*

+ Lecture notes:  
    + the examples and exercises related to the Bernoulli model with conjugate prior  
    
+ Application (optional):
    + [Dalal, S. R., Fowlkes, E. B., & Hoadley, B. (1989). Risk analysis of the space shuttle: Pre-Challenger prediction of failure. Journal of the American Statistical Association, 84(408), 945-957.](https://www.tandfonline.com/doi/abs/10.1080/01621459.1989.10478858)    

+ References for *RJAGS*:  
    + [JAGS homepage](http://mcmc-jags.sourceforge.net)  
    + [JAGS R CRAN Repository](https://cran.r-project.org/web/packages/rjags/index.html)  
    + [JAGS Reference Manual](https://cran.r-project.org/web/packages/rjags/rjags.pdf)  
    + [JAGS user manual](https://sourceforge.net/projects/mcmc-jags/files/Manuals/4.x/jags_user_manual.pdf/download) 

+ Reference for *R*:  
    + [Cheat sheet with basic commands](https://www.rstudio.com/wp-content/uploads/2016/10/r-cheat-sheet-3.pdf)   

+ Reference of *rmarkdown* (optional):  
    + [R Markdown cheatsheet](https://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf)  
    + [R Markdown Reference Guide](http://442r58kc8ke1y38f62ssb208-wpengine.netdna-ssl.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)  
    + [knitr options](https://yihui.name/knitr/options)

+ Reference for *Latex* (optional):  
    + [Latex Cheat Sheet](https://wch.github.io/latexsheet/latexsheet-a4.pdf)  

---

***New software***   

+ R package `rjags` functions:    
    + `jags.model{rjags}`  
    + `jags.samples{rjags}` 
    + `update{rjags}`

---


# Application: Challenger O-ring {-}

On January 28, 1986, a routine launch was anticipated for the Challenger space shuttle. Seventy-three seconds into the flight, disaster happened: the shuttle broke apart, killing all seven crew members on board. [Here is the video](https://youtu.be/fSTrmJtHLFU?t=99). The Rogers commission concluded that the Challenger accident was caused by gas leak through the 6 O-ring joints of the shuttle. Essentially the presence of distressed O-ring joints was the crucial factor that caused the explosion. 

[Dalal, Fowlkes and Hoadley (1989)](https://www.tandfonline.com/doi/abs/10.1080/01621459.1989.10478858) analysed a dataset that contains the presence of   distressed O-rings, the temperature in the platform, and the leak check pressure  for 23 previous shuttle flights. The the data-set is provided below, where in column *Defective.O.rings*, (1) stands for presence of distressed O-rings, and (0) for absence, while the rest of the columns are self explained. 


```{r, results="hide"}
# Load R package for printing
library(knitr)
library(kableExtra)
```

```{r}
# load the data
#mydata <- read.csv("./challenger_data.csv")
mydata <- read.csv("https://raw.githubusercontent.com/georgios-stats/Bayesian_Statistics/master/ComputerPracticals/Bernoulli_model_with_conjugate_priors/challenger_data.csv")
# print data 
## (that's a sophisticated command with fancy output, feel free to ignore it)
kable(mydata)%>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

Here,  

+ we will use only the observations in variable *Defective.O.rings* (so ignore the temperature and pressure measurements) from *04/12/1981* to *01/12/1986* ($23$ flights), with purpose to:  

+ learn the (limiting relative) frequency of having defective O-rings? 

+ predict the outcome in the $24$th flight on	*1/28/86*, given the information from the previous $23$ flights considered.

---

# The model: Bernoulli model {-}

Let $y_{i}$ denote the presence of a defective O-ring in the $i$th flight ($0$ for absence, and $1$ for presence).

Regarding the statistical model, we assume that $y_{i}$ can be modeled as observations generated independently from  a Bernoulli distribution with with common parameter $p$. Here, $p$ denotes the relative frequency of defective O-rings at any flight.  

Regarding the prior model, we assign a Beta prior distribution with fixed hyper-parameters $a_0=1.0$, $b_0=1.0$ on $p$ to account for its uncertainty. 

The Bayesian hierarchical model under consideration is:  
\[
\begin{cases}
y_{i}|p\sim & \text{Bernoulli}(p),\quad\text{for, }i=1,...,n\\
p\sim & \text{Beta}(a_0,b_0) ,
\end{cases}
\]
with hyper-parameter values $a_0=1.0$, $b_0=1.0$.  

# Task

We write a RJAGS program implementing the hierarchical model above, in order to generate a sample of size $n=100000$ from the posterior distribution  $$p^{(j)}\sim\pi(p|y_{1:n})\  ,\  \  j=1,...,n.$$

### ... your answer {-}

For now I am doing it for you.

***step 1***

Load the library

```{r, results="hide"}
# Load rjags
library("rjags") 
```

***step 2***

Create an input script, for rjags, containing the Bayesian hierarchical model  

```{r}
# Input parameters  : n, y, a_0, b_0
# output parameters : p
hierarhicalmodel <- "

  model {

  # this is related tot he sampling distribution
    
    for ( i in 1 : n ) {
      y[ i ] ~ dbern( p )
    }
    
   # this is related to the prior distributions
    
    p ~ dbeta( a_0 , b_0 )
    
  }

"
```

***step 3***

Create an input list, for jags, containing the data and fixed  parameters of the model  

```{r}

y_obs <- mydata[ -nrow(mydata) , 4 ] # exclude the last row, and use only the 4th column

y_obs <- as.numeric( y_obs == 1 )    # make it numeric 

n_obs <- length( y_obs )

a_0 <- 1.0

b_0 <- 1.0

data.bayes <- list(y = y_obs,
                    n = n_obs,
                    a_0 = a_0,
                    b_0 = b_0)
```

***step 4***

Create an input list, for jags, containing the data and fixed  parameters of the model 

```{r, results="hide"}
model.smpl <- jags.model( file = textConnection(hierarhicalmodel),
                          data = data.bayes)

```

For further reading: 

Alternatively we could have used the routine `coda.samples{rjags}` returning the same information but with an object of type `mcmc.list` that can be analysed by the tools provided in the R packages `coda` and `boa`. We do not discuss about these packages as the are not in stable release yet, and they may contain bugs. Pls keep an eye on them.



***step 5***

Initialize the sampler with $N_{\text{adapt}}=1000$ iterations.   

+ This is a warming-up procedure (used as a black-box), where the sampler is automatically tuned and calibated before it starts generatign your samples.   

+ Regarding $N_{\text{adapt}}=1000$, the larger the better.

```{r}
adapt( object = model.smpl,
       n.iter = 1000 )
```

***step 6***

Generate a posterior sample of size $N=10000$.

Use  

+ `jags.samples{rjags}`  

<!-- Function jags.samples initially generates a sample sequence of `n.iter` values, but it returns as a sample only every  `thin` sample value from the initial sample sequence. Hence the sample that it returns has size $N = $ `n.iter` $\times$ `thin`.  -->

We need to pay attention on two flaqs:  

+ the `n.iter`: the total size of the total sample sequence.   

<!-- + the `thin`: the thining interval -->

+ the `variable.names`: it specifies the names of the random variables corresponding to the posterior samples I am interested in generating   



```{r, results="hide"}
N = 10000      # the size of the sample we ll gonna get
output = jags.samples( model= model.smpl,          # the model
                       variable.names= c("p"),    # names of variables to be sampled
                       n.iter = N                # size of sample
                       )
save.image(file="BernoulliModel.RData") 
```

Check the names of the variables sampled  

+ use `names {base}`

```{r}
names(output)
```

Check the dimensions of each of the variables sampled  

+ use `dim {base}`

```{r}
dim( output$p )
# the first dimension is the numbers of columns of the variable
# the second dimention is the size of the sample drawn
# the third dimension is the number of the sub-samples drawn (in our case it is just 1)
```

Copy the sample of each variable in a vector with a more friendly name...

```{r}
pr.smpl <-output$p
```


---

# Task 

Print the trace plot of the sample drawn from the posterior distribution $$p^{(j)}\sim\pi(p|y_{1:n})\  \  j=1,...,N$$. 

Use functions:  
    
+ `plot {graphics}`.

A glimpse forward into Term 2 of Bayesian stats:  

+ A good quality sample for the purposes of Monte Carlo integration is the one whose trace plot looks completely uncorrelated.  
    + e.g. like the independence assumption of the residuals in the simple linear regression in SC2.  

+ To improve the quality of the sample, you can go back to the sampling stem and play with the flaq values of `n.iter` and `thin` in `jags.samples{rjags}`.  

### ... your answer {-}

Below try to address the task by creating R chunks, and running them, and adding comments to the document.


```{r}
# write your R code


```

---



# Task

+ Compute and plot the MC approximation for the posterior distribution cumulative function (CDF) of $p$, by computing the empirical CDF as 
\begin{align*}
F_{\pi}(p\le c|y_{1:n}) 
&=   E_{\pi}(\text{1}(p\le c)|y_{1:n}) \\
&\approx \frac{1}{N}\sum_{i=1}^{N}\text{1}(p^{(i)}\le c)
\end{align*}
for $c\in (0,1)$


+ Compute and plot the exact posterior CDF  of $p$,  which is the CDF of the distribution 
\[
p|y_{1:n}\sim\text{Beta}(a_{n},b_{n})
\]
where  
\begin{align*}
a_{n}= & a_{0}+n\bar{y}\\
b_{n}= & b_{0}+n-n\bar{y}.
\end{align*} 



Print them on the same plot.  


Use functions:  

+ `lines {graphics}`  

### ... your answer {-}

Below try to address the task by creating R chunks, and running them, and adding comments to the document.


```{r}
# write your R code


```


---


# Task

+ Compute and plot the MC approximation for the posterior distribution density of $p|y_{1:n}$ as a histogram 
    + use the function `hist {graphics}` provided from R.
    
    + Just for your information, the mathematical formula of a histogram estimator is
$$
\pi(p|y_{1:n}) \approx \frac{1}{2\epsilon}\frac{1}{N}\sum_{j=1}^{N} \text{1}\left(p^{(j)}\in (p-\epsilon,p+\epsilon]\right) \  \text{for small }\epsilon
$$ 
for $p\in (0,1)$,   however, you do not need to code it. Just use  the function `hist {graphics}`


+ Compute and plot the exact the posterior distribution density of $p|y_{1:n}$ which is the PDF of 
\[
p|y_{1:n}\sim\text{Beta}(a_{n},b_{n})
\]
where  
\begin{align*}
a_{n}= & a_{0}+n\bar{y}\\
b_{n}= & b_{0}+n-n\bar{y}.
\end{align*}


Use functions:  

+ `hist {graphics}` 

### ... your answer {-}

Below try to address the task by creating R chunks, and running them, and adding comments to the document.

```{r}
# write your R code


```


---

# Task  

+ Compute the MC approximate of the   posterior probability that the frequence parameter $p$ is greater than or equal to $0.5$ as     
\begin{align}
\text{Pr}_\pi (p\ge 0.5|y_{1:n}) =& 1-\text{Pr}_\pi (p < 0.5|y_{1:n})\\
=& 1-\text{E}_\pi (\text{1}(p < 0.5)|y_{1:n})\\
\approx & \frac{1}{N}\sum_{i=1}^{N} (\text{1}(p^{(i)} < 0.5))
\end{align}  

+ Compute its exact value of as   
\begin{align}
\text{Pr}_\pi (p\ge 0.5|y_{1:n}) =& 1-\text{Pr}_\pi (p< 0.5|y_{1:n}) \\
& 1-\text{Pr}_{\text{Beta}(a_{n},b_{n})} (p < 0.5|y_{1:n}) \\
& 1-\int_{-\infty}^{0.5} \text{Beta}(p|a_{n},b_{n}) \text{d}p
\end{align} 
where
\begin{align*}
a_{n}= & a_{0}+n\bar{y}\\
b_{n}= & b_{0}+n-n\bar{y}
\end{align*}
 


### ... your answer {-}

Below try to address the task by creating R chunks, and running them, and adding comments to the document.


```{r}
# write your R code


```


---

# Task  

+ compute the MC approximate of the  $95$% posterior equal tail credible interval of the frequency parameter $p$ is   
$$
[
Q_{0.025}(p|y_{1:n})\  ,\  Q_{0.975}(p|y_{1:n})
]
$$
where $Q_{\alpha}(p|y_{1:n})$  is the $\alpha$-th quantile of the posterior distribution of $p$  

+ compute the exact $95$% posterior equal tail credible interval of the frequency parameter $p$ is   
$$
[
Q_{0.025}(p|y_{1:n})\  ,\  Q_{0.975}(p|y_{1:n})
]
$$
where $Q_{\alpha}(p|y_{1:n})$ is the $\alpha$-th quantile of the posterior distribution of $p$ which is   
$$
p\sim \text{Beta}(a_{n},b_{n})
$$
where
\begin{align*}
a_{n}= & a_{0}+n\bar{y}\\
b_{n}= & b_{0}+n-n\bar{y}
\end{align*}




Use:  

+ `quantile{stats}`  
+ `qbeta{stats}`  


### ... your answer {-}

Below try to address the task by creating R chunks, and running them, and adding comments to the document.


```{r}
# write your R code


```


---

# Task 


+ Compute the MC approximate of the posterior expected value of $p$,  $\text{E}_{\pi}(p|y_{1:n})$, as
$$
\text{E}_{\pi}(p|y_{1:n}) \approx \frac{1}{N}\sum_{i=1}^{N}p^{(i)}
$$

+ Compute exact value of $\text{E}(p|y_{1:n})$  which is
\[
\text{E}_{\pi}(p|y_{1:n})=\frac{a_{n}}{a_{n}+b_{n}}
\]
where
\begin{align*}
a_{n}= & a_{0}+n\bar{y}\\
b_{n}= & b_{0}+n-n\bar{y}
\end{align*}

### ... your answer {-}

Below try to address the task by creating R chunks, and running them, and adding comments to the document.

```{r}
# write your R code


```




---

# Task  

+ Compute the MC approximate of the posterior expected value of the odds parameter $\theta=\frac{p}{1-p}$ which is denoted as $\text{E}_{\pi}(\theta|y_{1:n})$.  
$$
\text{E}_{\pi}(\theta|y_{1:n})=\text{E}_{\pi}(\frac{p}{1-p}|y_{1:n})\approx \frac{1}{N}\sum_{i=1}^{N}\frac{p^{(i)}}{1-p^{(i)}}
$$

+ Compute the MC approximation of $\text{E}_{\pi}(\theta|y_{1:n})$ with its exact value   which is 
\[
\text{E}_{\pi}(\theta|y_{1:n})=\text{E}_{\pi}(\frac{p}{1-p}|y_{1:n})=\frac{a_{n}}{b_{n}-1}
\]
where
\begin{align*}
a_{n}= & a_{0}+n\bar{y}\\
b_{n}= & b_{0}+n-n\bar{y}
\end{align*}



### ... your answer {-}

Below try to address the task by creating R chunks, and running them, and adding comments to the document.


```{r}
# write your R code


```





---


# Task

+ Compute the MC approximation of the predictive distribution mass function of $y_{n+1}|y_{1:n}$, as
\begin{align*}
f_{\pi}(y_{n+1}=c|y_{1:n}) 
=& \int_{0}^{1} f(y_{n+1}=c|p) \pi(p|y_{1:n}) \text{d}p, \  \    c\in \{0,1\}   \\
&\\
=& \int_{0}^{1} p^{c} (1-p)^{1-c} \pi(p|y_{1:n}) \text{d}p, \  \  c\in \{0,1\}  \\
&\\
=& 
\begin{cases}
\int_{0}^{1} (1-p)         \pi(p|y_{1:n}) \text{d}p & \  \text{, c}=0 \\
&\\
\int_{0}^{1}p    \pi(p|y_{1:n}) \text{d}p & \  \text{, c}=1
\end{cases} \\
=& 
\begin{cases}
1-\text{E}(p|y_{1:n}) & \  \text{, c}=0 \\
&\\
\text{E}(p|y_{1:n}) \text{d}p & \  \text{, c}=1
\end{cases} \\
&\\
\approx& 
\begin{cases}
1-  \frac{1}{N}\sum_{j=1}^{N} p^{(j)} & \  \text{, c}=0 \\
  &\\
\frac{1}{N}\sum_{j=1}^{N} p^{(j)} & \  \text{, c}=1
\end{cases}
\end{align*}
by using a barplot. 

+ Compute the  exact predictive distribution mass function of $y_{n+1}|y_{1:n}$, as
\begin{align*}
f(y_{n+1}=c|y_{1:n})
=&
\frac{B(a_{n}+c,b_{n}+1-c)}{B(a_{n},b_{n})}\text{1}(c\in \{0,1\}) \\
=& 
\begin{cases}
\frac{B(a_{n},b_{n}+1)}{B(a_{n},b_{n})} & \text{, c}=0 \\
  &\\
\frac{B(a_{n}+1,b_{n})}{B(a_{n},b_{n})} & \text{, c}=1
\end{cases}
\end{align*}

where  

\begin{align*}
a_{n}= & a_{0}+n\bar{y}\\
b_{n}= & b_{0}+n-n\bar{y}
\end{align*}

by using a barplot. 



Use functions:  

+ `barplot {graphics}`,  

### ... your answer {-}

Below try to address the task by creating R chunks, and running them, and adding comments to the document.

```{r}
# write your R code


```


# Discussion

 
 Now, it is January 27, 1986, and you take part in the 3-hour teleconference with people from Morton Thiokol, Marshall space flight center, and Kennedy space center.


The forcast says that tommorrow the temperature too frosty. This is a bit unusual in the area and NASA people wonder if temperature can cause problems. 

You perform the aforesaid statistical analysis and you find that the predictive probability in the next flight that there will be at least a defective O ring is less than $0.5$. 

You report your findinds, and the next day, January 28, 1986 explodes due to the presence of a defective O-ring component.

What did it go wrong with your aforesaid analysis?

How could you possibly do your analysis so that it can succesfully sugest you that the next flight (the one on January 28, 1986) will be a disaster if it is preformed?


If you have finished this handout, you can proceed tot he next [Bernoulli regression model].









    
