<!-- -------------------------------------------------------------------------------- -->

<!-- Copyright 2019 Georgios Karagiannis -->

<!-- georgios.karagiannis@durham.ac.uk -->
<!-- Assistant Professor -->
<!-- Department of Mathematical Sciences, Durham University, Durham,  UK  -->

<!-- This file is part of Bayesian_Statistics (MATH3341/4031 Bayesian Statistics III/IV) -->
<!-- which is the material of the course (MATH3341/4031 Bayesian Statistics III/IV) -->
<!-- taught by Georgios P. Katagiannis in the Department of Mathematical Sciences   -->
<!-- in the University of Durham  in Michaelmas term in 2019 -->

<!-- Bayesian_Statistics is free software: you can redistribute it and/or modify -->
<!-- it under the terms of the GNU General Public License as published by -->
<!-- the Free Software Foundation version 3 of the License. -->

<!-- Bayesian_Statistics is distributed in the hope that it will be useful, -->
<!-- but WITHOUT ANY WARRANTY; without even the implied warranty of -->
<!-- MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the -->
<!-- GNU General Public License for more details. -->

<!-- You should have received a copy of the GNU General Public License -->
<!-- along with Bayesian_Statistics  If not, see <http://www.gnu.org/licenses/>. -->

<!-- -------------------------------------------------------------------------------- -->

---
title: "Credible sets"
output: flexdashboard::flex_dashboard
runtime: shiny
---

<!-- Conjugate prior Applet -->
<!-- Copyright (C) 2018  Georgios Karagiannis -->
<!-- georgios.karagiannis@durham.ac.uk -->

<!-- This program is free software: you can redistribute it and/or modify -->
<!-- it under the terms of the GNU General Public License as published by -->
<!-- the Free Software Foundation, either version 3 of the License, or -->
<!-- (at your option) any later version. -->

<!-- This program is distributed in the hope that it will be useful, -->
<!-- but WITHOUT ANY WARRANTY; without even the implied warranty of -->
<!-- MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the -->
<!-- GNU General Public License for more details. -->

<!-- You should have received a copy of the GNU General Public License -->
<!-- along with this program. If not, see <http://www.gnu.org/licenses/>. -->

```{r}
library('rootSolve')
```


# Unimodal case

```{r}

a_def_uni <- 2.0
b_def_uni <- 2.0

x_min_def_uni <- 0.0
x_max_def_uni <- 1.0

k_cut_def_uni <- 0.0018
alpha_def_uni <- 0.05

fun_pdf_uni <- function ( x, 
                          a = a_def_uni,
                          b = b_def_uni) {
  
  fx <- dbeta(x, a, b, ncp = 0, log = FALSE)
  
  return (fx)
}



# get the roots

fun_cost_uni <- function( x , k_cut, a, b ) { 
  return( fun_pdf_uni(x, a, b)-k_cut ) 
}

comp_crint_uni <-function(k_cut, x_min, x_max, a, b ) {
  
  interval <- c( x_min , x_max )
  
  fun <- function (x) fun_cost_uni( x , k_cut, a, b )
  
  rootpoints <- uniroot.all(fun, interval)  
  
  cr_n <- (0.5*length(rootpoints))
  
  cr <- matrix(rootpoints,cr_n,byrow = TRUE)
  
  return ( cr)
}


comp_mass_uni <- function (crint, a, b) {
  
  fun <- function (x) fun_pdf_uni( x , a, b )
  
  n <- dim(crint)[1]
  
  mass <- 0.0
  for(i in 1:n)
    mass <- mass + integrate( fun, crint[i,1], crint[i,2] )$value
  
  return(mass)
}



check_alpha_uni <- function( alpha, crint , a, b ) {
  al_t <- comp_mass_uni(crint , a, b)-1
  if ( 1-al_t >= 1-alpha ) 
    return (1)
  else
    return (0)
}



plot_cr_uni <- function( 
  alpha = alpha_def_uni,
  k_cut = k_cut_def_uni,
  x_min = x_min_def_uni, 
  x_max = x_max_def_uni,
  a = a_def_uni, 
  b = b_def_uni) {
  
  fntsz <- 1.5
  
  
  # plot the PDF
  
  x_grid <- seq(from = x_min, to = x_max, length.out=1000)
  
  fun_pdf <- function (x) fun_pdf_uni( x , a, b )
  
  fun_pdf_grid <- apply(matrix(x_grid,1), 1, fun_pdf)
  
  plot(x_grid, fun_pdf_grid, 
       type='l',
       main = paste('Find the ', toString(format(1-alpha,digits=4)), 'HPD interval'),
       ylab = 'Density',
       xlab = 'values',
          cex.lab=fntsz, 
          cex.axis=fntsz, 
          cex.main=fntsz, 
          cex.sub=fntsz)
  
  # plot the ruller
  
  abline(h = k_cut, 
         col='red',
         lwd = 3,
          cex.lab=fntsz, 
          cex.axis=fntsz, 
          cex.main=fntsz, 
          cex.sub=fntsz)
  
  # find the roots
  
  crint <- comp_crint_uni(k_cut, x_min , x_max, a, b )
  
  # plot arroes to show the boundaries
  
  n <- dim(crint)[1]
  m <- dim(crint)[2]
  for(i in 1:n)
    for(j in 1:m) {
      xx <-  crint[i,j]
      fxx <- fun_pdf_uni( xx, a, b )
      arrows(  xx, fxx, xx, 0.0, col=rgb(0, 1, 0, 1), length = 0.15  )
    }
  
  # compute the CI mass
  
  # check if it is a CI
  alpha_cr <- 1-comp_mass_uni(crint, a, b)
  if ( 1-alpha_cr>=1-alpha   )
    text( 0.5, 0.5, 'BINGO!!!'  ,
        cex=fntsz)
  else
    text( 0.5, 0.5, 
          'NOT BINGO!!!'  ,
        cex=fntsz)
  
    text( x_min+0.1, k_cut+0.05, 
        paste('k = ',toString(format(k_cut,digits=4) )),
        cex=fntsz)
    text( x_min+0.3, k_cut+0.05, 
        paste('mass=',toString(format(1-alpha_cr,digits=4) )),
        cex=fntsz)
  
  # plot shaded area of the CI
  
  n <- dim(crint)[1]
  for (i in 1:n) {
    
    xtick <- seq( from = crint[i,1], to = crint[i,2], length=100)  
    ytick <- apply(matrix(xtick,1), 1, fun_pdf)
    
    xtick <- c(crint[i,1], xtick, crint[i,2])
    ytick <- c(0.0, ytick, 0.0)
    
    polygon( xtick , ytick , col=rgb(0, 0, 1, 0.3))
  }
  
  # plot the legend
  legend('topright',
         bg="transparent",
         c( 
           paste('Interval mass..: ', toString(format(1-alpha_cr,digits=4)), sep=''), 
           'Bounbaries', 
           paste( 'k .....................:',toString(format(k_cut,digits=4) ) ) 
         ),
         lty=c(1),
         lwd=c(2.5),
         col=c('blue', 'green', 'red'),
        cex=fntsz
  ) 
  
}



```



Inputs {.sidebar} 
-----------------------



***The distribution***  

\begin{align*}
\omega|x_{1:n} \sim \text{Be}(g, h) 
\end{align*}

... a Beta distribution  

***Credible interval parameters***

```{r,echo=FALSE}
sliderInput("Uni_alpha",
                 "$$1-\\alpha:$$",
                  min = 0.0001,
                  max = 0.9999,
                  step = 0.05,
                  value = 0.8)

sliderInput("Uni_ka",
                 "$$k_{\\alpha}$$",
                  min   = 0.001,
                  max   = 1.99,
                  step  = 0.01,
                  value = 0.0018)
```

***Distr. parameters***

```{r,echo=FALSE}
sliderInput("Uni_a",
                 "$$g:$$",
                  min = 1.1,
                  max = 5.1,
                  step = 0.1,
                  value = 2.0)

sliderInput("Uni_b",
                 "$$h:$$",
                  min = 1.1,
                  max = 5.1,
                  step  = 0.1,
                  value = 2.0)
```



Column {.tabset}
-----------------------


### ***Description (posterior) *** 

We present the construction of the credible sets for parametric inference. 

***Definition: Posterior Credible Set***  

Any set $C_{a}\subseteq\Theta$ such that 
\[
\pi(\theta\in C_{a}|x_{1:n})=\int_{C_{a}}\pi(\text{d}\theta|x_{1:n})\ge1-a
\]
 is called `$100(1-a)\%$' posterior credible set for $\theta$, with
respect to the posterior distribution $\pi(\text{d}\theta|x_{1:n})$.


***Definition: Posterior highest probability density (HPD) Set***  

The $100(1-a)\%$ HPD set for $\theta\in\Theta$ with respect to the posterior distribution $\pi(\text{d}\theta|x_{1:n})$ is the subset $C_{a}$ of $\Theta$ of the form 
\[
C_{a}=\{\theta\in\Theta:\pi(\theta|x_{1:n})\ge k_{a}\}
\]
 where $k_{a}$ is the largest constant such that 
\[
\pi(\theta\in C_{a}|x_{1:n})\ge1-a
\]

***Theorem: 1D Case of HPD***  

Let $\theta$ be a random variable that admits measure $\pi(\text{d}\theta|x_{1:n})$ with unimodal density $\pi(\theta|x_{1:n})$. If the interval $C_{a}=[L,U]$ satisfies 

1. $\int_{L}^{U}\pi(\theta|x_{1:n})\text{d}\theta=1-a$,

2. $\pi(U|x_{1:n})=\pi(L|x_{1:n})>0$, and 

3. $\theta_{\text{mode}}\in(L,U)$, where $\theta_{\text{mode}}$ is the mode of $\pi(\theta|x_{1:n})$,

then interval $C_{a}=[L,U]$ is the HPD interval of $\theta$ with respect to $\pi(\text{d}\theta|x_{1:n})$.

***Remark***  

HPD sets are the credible sets with the smallest `size'


### ***Description (predictive) *** 

We present the construction of the credible sets for predictive inference. 

***Definition: Predictive Credible Set***  

Any set $C_{a}\subseteq\mathcal{X}$ such that 
\[
p(y\in C_{a}|x_{1:n})=\int_{C_{a}}p(\text{d} y|x_{1:n})\ge1-a
\]
 is called `$100(1-a)\%$' predictive credible set for $y=x_{n+1}$, with
respect to the predictive distribution $p(\text{d} y|x_{1:n})$.


***Definition: Predictive highest probability density (HPD) Set***  

The $100(1-a)\%$ HPD set for $y\in\mathcal{X}$ with respect to the posterior distribution $p(\text{d}y|x_{1:n})$ is the subset $C_{a}$ of $\mathcal{X}$ of the form 
\[
C_{a}=\{y\in\mathcal{X}:p(\text{d} y|x_{1:n})\ge k_{a}\}
\]
 where $k_{a}$ is the largest constant such that 
\[
p(y\in C_{a}|x_{1:n})\ge1-a
\]

***Theorem: 1D Case of HPD***  

Let $y$ be a random variable that admits measure $p(\text{d}y|x_{1:n})$ with unimodal density $p(y|x_{1:n})$. If the interval $C_{a}=[L,U]$ satisfies 

1. $\int_{L}^{U}p(y|x_{1:n})\text{d}y=1-a$,

2. $p(U|x_{1:n})=p(L|x_{1:n})>0$, and 

3. $y_{\text{mode}}\in(L,U)$, where $y_{\text{mode}}$ is the mode of $p(y|x_{1:n})$,

then interval $C_{a}=[L,U]$ is the HPD interval of $y$ with respect to $\pi(\text{d}y|x_{1:n})$.

***Remark***  

HPD sets are the credible sets with the smallest `size'

### ***Toy example*** 

***Interest lies in*** constructing the  $100(1-a)\%$ HPD interval of $\omega\in\Omega$ of $\omega\sim\pi(\text{d}\omega|x_{1:n})$ with PDF:
\[
\pi(\omega|x_{1:n}) = \text{Be}(\omega|g,h)
\]

... a Beta distribution


***This can be***  
+ Posterior distr. for \[\omega=\theta, \text{ and } \Omega=\Theta, \text{ and }, \pi(\text{d}\cdot|x_{1:n})\]  
+ Predictive distr. for \[\omega=y,     \text{ and } \Omega=\mathcal{X}, \text{ and }, p(\text{d}\cdot|x_{1:n})\] 


### ***Parametric inferece***

```{r,echo=FALSE}
renderPlot({
   plot_cr_uni(alpha=1-input$Uni_alpha, 
                  k_cut=input$Uni_ka, 
                  a=input$Uni_a, 
                  b=input$Uni_b
                  )
  })
```



# Multimodal case


```{r}
w1_def_mult <- 0.6 
mu1_def_mult <- -43.8
sig1_def_mult <- 8 

w2_def_mult <- 0.3
mu2_def_mult <- 0.0
sig2_def_mult <- 10 

mu3_def_mult <- 40.0
sig3_def_mult <- 10

x_min_def_mult <- -70
x_max_def_mult <- 63

k_cut_def_mult <- 0.0018

alpha_def_mult <- 0.05


fun_pdf_mult <- function ( x, w1=w1_def_mult, w2=w2_def_mult, 
                            mu1=mu1_def_mult, sig1=sig1_def_mult, 
                            mu2=mu2_def_mult, sig2=sig2_def_mult, 
                            mu3=mu3_def_mult, sig3=sig3_def_mult ) {
  
  w3 <- 1-(w1+w2)
  fx <-     w1*dnorm(x, mean = mu1, sd = sig1, log = FALSE)
  fx <- fx +w2*dnorm(x, mean = mu2, sd = sig2, log = FALSE)
  fx <- fx +w3*dnorm(x, mean = mu3, sd = sig3, log = FALSE)
  
  return (fx)
}


# get the roots

fun_cost_mult <- function( x , k_cut ) { 
  return( fun_pdf_mult(x)-k_cut ) 
}

comp_crint_mult <-function(k_cut, x_min, x_max ) {
  
  interval <- c( x_min , x_max )
  
  fun <- function (x) fun_cost_mult( x , k_cut )
  
  rootpoints <- uniroot.all(fun, interval)  
  
  cr_n <- (0.5*length(rootpoints))
  
  cr <- matrix(rootpoints,cr_n,byrow = TRUE)
  
  return ( cr)
}



comp_mass_mult <- function (crint) {
  
  n <- dim(crint)[1]

  mass <- 0.0
  for(i in 1:n)
    mass <- mass + integrate( fun_pdf_mult, crint[i,1], crint[i,2] )$value
  
  return(mass)
}


check_alpha_mult <- function( alpha, crint ) {
  al_t <- comp_mass_mult(crint)-1
  if ( 1-al_t >= 1-alpha ) 
    return (1)
  else
    return (0)
}

plot_cr_mult <- function( 
                    alpha = alpha_def_mult,
                    k_cut = k_cut_def_mult,
                    x_min = x_min_def_mult, 
                    x_max = x_max_def_mult ) {
  
  fntsz <- 1.5
  
  # plot the PDF
  
  x_grid <- seq(from = x_min, to = x_max, length.out=1000)
  
  fun_pdf_grid <- apply(matrix(x_grid,1), 1, fun_pdf_mult)
  
  plot(x_grid, fun_pdf_grid, 
       type='l',
       main = paste('Find the ', toString(format(1-alpha,digits=4)), 'HPD interval'),
       ylab = 'Density',
       xlab = 'values',
          cex.lab=fntsz, 
          cex.axis=fntsz, 
          cex.main=fntsz, 
          cex.sub=fntsz)
  
  # plot the ruller
  
  abline(h = k_cut, 
         col='red',
         lwd = 3,
          cex.lab=fntsz, 
          cex.axis=fntsz, 
          cex.main=fntsz, 
          cex.sub=fntsz)
  
  # find the roots
  
  crint <- comp_crint_mult(k_cut, x_min , x_max )
  
  # plot arroes to show the boundaries
  
  n <- dim(crint)[1]
  m <- dim(crint)[2]
  for(i in 1:n)
    for(j in 1:m) {
      xx <-  crint[i,j]
      fxx <- fun_pdf_mult( xx )
      arrows(  xx, fxx, xx, 0.0, col=rgb(0, 1, 0, 1), length = 0.15  )
    }
  
  # compute the CI mass
  
  # check if it is a CI
  alpha_cr <- 1-comp_mass_mult(crint)
  if ( 1-alpha_cr>=1-alpha   )
    text( 0, 0.02, 'BINGO!!!' , cex=fntsz )
  else
    text( 0, 0.02, 'NOT BINGO!!!' , cex=fntsz )
    
      text( x_min+10, k_cut+0.001, 
        paste('k = ',toString(format(k_cut,digits=4) )),
        cex=fntsz)
    text( x_min+30, k_cut+0.001, 
        paste('mass=',toString(format(1-alpha_cr,digits=4) )),
        cex=fntsz)
  
  # plot shaded area of the CI
 
  n <- dim(crint)[1]
  for (i in 1:n) {
    
    xtick <- seq( from = crint[i,1], to = crint[i,2], length=100)  
    ytick <- apply(matrix(xtick,1), 1, fun_pdf_mult)
    
    xtick <- c(crint[i,1], xtick, crint[i,2])
    ytick <- c(0.0, ytick, 0.0)
    
    polygon( xtick , ytick , col=rgb(0, 0, 1, 0.3))
  }
  
  # plot the legend
  legend('topright',
         bg="transparent",
         c( 
           paste('Interval mass..: ', toString(format(1-alpha_cr,digits=4)), sep=''), 
           'Bounbaries', 
           paste( 'k .....................:',toString(format(k_cut,digits=4) ) ) 
         ),
         lty=c(1),
         lwd=c(2.5),
         col=c('blue', 'green', 'red'),
        cex=fntsz
  ) 
  
}

```



Inputs {.sidebar} 
-----------------------

***The distribution***  

\begin{align*}
\omega|x_{1:n} \sim \  & 0.6\ \text{N}(-43.8,8^2) \\
&+0.3\ \text{N}(0,10^2)\\
&+0.1\ \text{N}(40,10^2)
\end{align*}

... a mixture of Gaussian distributions  

***Credible interval parameters***

```{r,echo=FALSE}
sliderInput("Mult_alpha",
                 "$$1-\\alpha:$$",
                  min = 0.0001,
                  max = 0.9999,
                  step = 0.05,
                  value = 0.8)

sliderInput("Mult_ka",
                 "$$k_{\\alpha}$$",
                  min   = 0.001,
                  max   = 0.029,
                  step  = 0.0001,
                  value = 0.0018)
```

***Distr. parameters***



Column {.tabset}
-----------------------


### ***Description (posterior) *** 

We present the construction of the credible sets for parametric inference. 

***Definition: Posterior Credible Set***  

Any set $C_{a}\subseteq\Theta$ such that 
\[
\pi(\theta\in C_{a}|x_{1:n})=\int_{C_{a}}\pi(\text{d}\theta|x_{1:n})\ge1-a
\]
 is called `$100(1-a)\%$' posterior credible set for $\theta$, with
respect to the posterior distribution $\pi(\text{d}\theta|x_{1:n})$.


***Definition: Posterior highest probability density (HPD) Set***  

The $100(1-a)\%$ HPD set for $\theta\in\Theta$ with respect to the posterior distribution $\pi(\text{d}\theta|x_{1:n})$ is the subset $C_{a}$ of $\Theta$ of the form 
\[
C_{a}=\{\theta\in\Theta:\pi(\theta|x_{1:n})\ge k_{a}\}
\]
 where $k_{a}$ is the largest constant such that 
\[
\pi(\theta\in C_{a}|x_{1:n})\ge1-a
\]

***Theorem: 1D Case of HPD***  

Let $\theta$ be a random variable that admits measure $\pi(\text{d}\theta|x_{1:n})$ with unimodal density $\pi(\theta|x_{1:n})$. If the interval $C_{a}=[L,U]$ satisfies 

1. $\int_{L}^{U}\pi(\theta|x_{1:n})\text{d}\theta=1-a$,

2. $\pi(U|x_{1:n})=\pi(L|x_{1:n})>0$, and 

3. $\theta_{\text{mode}}\in(L,U)$, where $\theta_{\text{mode}}$ is the mode of $\pi(\theta|x_{1:n})$,

then interval $C_{a}=[L,U]$ is the HPD interval of $\theta$ with respect to $\pi(\text{d}\theta|x_{1:n})$.

***Remark***  

HPD sets are the credible sets with the smallest `size'


### ***Description (predictive) *** 

We present the construction of the credible sets for predictive inference. 

***Definition: Predictive Credible Set***  

Any set $C_{a}\subseteq\mathcal{X}$ such that 
\[
p(y\in C_{a}|x_{1:n})\ge1-a
\]
 is called `$100(1-a)\%$' predictive credible set for $y=x_{n+1}$, with
respect to the predictive distribution $p(\text{d} y|x_{1:n})$.


***Definition: Predictive highest probability density (HPD) Set***  

The $100(1-a)\%$ HPD set for $y\in\mathcal{X}$ with respect to the posterior distribution $p(\text{d}y|x_{1:n})$ is the subset $C_{a}$ of $\mathcal{X}$ of the form 
\[
C_{a}=\{y\in\mathcal{X}:p(\text{d} y|x_{1:n})\ge k_{a}\}
\]
 where $k_{a}$ is the largest constant such that 
\[
p(y\in C_{a}|x_{1:n})=\int_{C_{a}}p(\text{d} y|x_{1:n})\ge1-a
\]

***Theorem: 1D Case of HPD***  

Let $y$ be a random variable that admits measure $p(\text{d}y|x_{1:n})$ with unimodal density $p(y|x_{1:n})$. If the interval $C_{a}=[L,U]$ satisfies 

1. $\int_{L}^{U}p(y|x_{1:n})\text{d}y=1-a$,

2. $p(U|x_{1:n})=p(L|x_{1:n})>0$, and 

3. $y_{\text{mode}}\in(L,U)$, where $y_{\text{mode}}$ is the mode of $p(y|x_{1:n})$,

then interval $C_{a}=[L,U]$ is the HPD interval of $y$ with respect to $\pi(\text{d}y|x_{1:n})$.

***Remark***  

HPD sets are the credible sets with the smallest `size'

### ***Toy example*** 

***Interest lies in*** constructing the  $100(1-a)\%$ HPD interval of $\omega\in\Omega$ of $\omega\sim\pi(\text{d}\omega|x_{1:n})$ with PDF:
\[
\pi(\omega|x_{1:n}) = 0.6\text{N}(\omega|-43.8,8^2)
+0.3\text{N}(\omega|0,10^2)
+0.1\text{N}(\omega|40,10^2)
\]

... a mixture of Gaussian distributions  

***This can be***  
+ Posterior distr. for \[\omega=\theta, \text{ and } \Omega=\Theta, \text{ and }, \pi(\text{d}\cdot|x_{1:n})\]  
+ Predictive distr. for \[\omega=y,     \text{ and } \Omega=\mathcal{X}, \text{ and }, p(\text{d}\cdot|x_{1:n})\] 


### ***Parametric inferece***

```{r,echo=FALSE}
renderPlot({
   plot_cr_mult(alpha=1-input$Mult_alpha, 
                  k_cut=input$Mult_ka
                  )
  })
```









