<!-- -------------------------------------------------------------------------------- -->

<!-- Copyright 2019 Georgios Karagiannis -->

<!-- georgios.karagiannis@durham.ac.uk -->
<!-- Assistant Professor -->
<!-- Department of Mathematical Sciences, Durham University, Durham,  UK  -->

<!-- This file is part of Bayesian_Statistics (MATH3341/4031 Bayesian Statistics III/IV) -->
<!-- which is the material of the course (MATH3341/4031 Bayesian Statistics III/IV) -->
<!-- taught by Georgios P. Katagiannis in the Department of Mathematical Sciences   -->
<!-- in the University of Durham  in Michaelmas term in 2019 -->

<!-- Bayesian_Statistics is free software: you can redistribute it and/or modify -->
<!-- it under the terms of the GNU General Public License as published by -->
<!-- the Free Software Foundation version 3 of the License. -->

<!-- Bayesian_Statistics is distributed in the hope that it will be useful, -->
<!-- but WITHOUT ANY WARRANTY; without even the implied warranty of -->
<!-- MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the -->
<!-- GNU General Public License for more details. -->

<!-- You should have received a copy of the GNU General Public License -->
<!-- along with Bayesian_Statistics  If not, see <http://www.gnu.org/licenses/>. -->

<!-- -------------------------------------------------------------------------------- -->

---
title: "Compare priors/posteriors"
output: flexdashboard::flex_dashboard
runtime: shiny
---

<!-- Compare priors Applet -->
<!-- Copyright (C) 2017  Georgios Karagiannis -->
<!-- georgios.karagiannis@durham.ac.uk -->

<!-- This program is free software: you can redistribute it and/or modify -->
<!-- it under the terms of the GNU General Public License as published by -->
<!-- the Free Software Foundation, either version 3 of the License, or -->
<!-- (at your option) any later version. -->

<!-- This program is distributed in the hope that it will be useful, -->
<!-- but WITHOUT ANY WARRANTY; without even the implied warranty of -->
<!-- MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the -->
<!-- GNU General Public License for more details. -->

<!-- You should have received a copy of the GNU General Public License -->
<!-- along with this program. If not, see <http://www.gnu.org/licenses/>. -->

```{r}
get_randnumb <-function(n) {
  u <- seq(0.001,0.999,length.out = n)
#  u <- runif(0,1,n = n)
}
```

```{r}
#library(rmutil)
dbetabinom <- function (y, size, m, s, log = FALSE) {
  # from rmutil package
    if (any(y < 0)) 
        stop("y must contain non-negative values")
    if (any(size < 0)) 
        stop("size must contain non-negative values")
    if (any(m <= 0) || any(m >= 1)) 
        stop("m must lie between 0 and 1")
    if (any(s <= 0)) 
        stop("s must be positive")
    ly <- max(length(y), length(m), length(s), length(size))
    if (length(y) != ly) {
        if (length(y) == 1) 
            y <- rep(y, ly)
        else stop("length of y incorrect")
    }
    if (length(size) != ly) {
        if (length(size) == 1) 
            size <- rep(size, ly)
        else stop("size must be the same length as y")
    }
    if (any(y > size)) 
        stop("y must be <= size")
    if (length(m) != ly) {
        if (length(m) == 1) 
            m <- rep(m, ly)
        else stop("m and y must have the same length")
    }
    if (length(s) != ly) {
        if (length(s) == 1) 
            s <- rep(s, ly)
        else stop("s and y must have the same length")
    }
    t <- s * m
    u <- s * (1 - m)
    tmp <- lbeta(y + t, size - y + u) - lbeta(t, u) + lchoose(size, 
        y)
    if (!log) 
        tmp <- exp(tmp)
    tmp
}

BinomialBeta_pdf<-function(x,a,b,n) {
  pdf <- dbetabinom(y=x, size=n, m=a/(a+b), s=a+b) ;
  return (pdf) 
}
```


```{r}
StudentT_pdf<-function(x,mu,lam,al) {
  pdf <- 0.5*log(lam) -0.5*log(al) -0.5*log(pi) ;
  pdf <- pdf +lgamma(0.5*(al+1)) -lgamma(0.5*al) ;
  pdf <- pdf -0.5*(al+1)*log(1+lam/al*(x-mu)^2) ;
  pdf <- exp(pdf) ;
  return (pdf)  
}
```


```{r}
dinvgamma <- function(x, a, b){
  return ( exp( a*log(b)-lgamma(a) -(a+1)*log(x) -b/x ) ) ;
}
```



# Bernoulli model


```{r,echo=FALSE}
BernoulliModel_SufStat<-function(x,n) {
 r <- sum(x[1:n]) ; 
}

# Conjugate

BernoulliModel_PDF_posterior<-function(theta,r,n,a,b) {
  pdf <- dbeta(theta, a+r, b+n-r) ;
   return (pdf)  
}

BernoulliModel_PDF_prior<-function(theta,a,b) {
  pdf <- dbeta(theta, a, b) ;
   return (pdf)  
}

BernoulliModel_PDF_pred<-function(y,r,n,a,b) {
  pdf <- BinomialBeta_pdf(y, a+r, b+n-r, 1) ;
   return (pdf)  
}

# Jeffreys

BernoulliModel_PDF_Jeffreys_prior<-function(theta) {
  pdf <- dbeta(theta, 0.5, 0.5) ;
   return (pdf)  
}

BernoulliModel_PDF_Jeffreys_posterior<-function(theta,r,n) {
  pdf <- dbeta(theta, 0.5+r, 0.5+n-r) ;
   return (pdf)  
}

BernoulliModel_PDF_Jeffreys_pred<-function(y,r,n) {
  pdf <- BinomialBeta_pdf(y, 0.5+r, 0.5+n-r, 1) ;
   return (pdf)  
}

# Laplace

BernoulliModel_PDF_Laplace_prior<-function(theta) {
  pdf <- dbeta(theta, 1.0, 1.0) ;
   return (pdf)  
}

BernoulliModel_PDF_Laplace_posterior<-function(theta,r,n) {
  pdf <- dbeta(theta, 1.0+r, 1.0+n-r) ;
   return (pdf)  
}

BernoulliModel_PDF_Laplace_pred<-function(y,r,n) {
  pdf <- BinomialBeta_pdf(y, 1.0+r, 1.0+n-r, 1) ;
   return (pdf)  
}

plot_Bernoulli_parampred <- function(theta=0.5, n=10, a=1, b=1) {
  
  # theta=0.5; n=10; a=1; b=1
  
  fntsz <- 1.5;
  op <- par(cex = fntsz)
  
  obs <-get_randnumb(n) ;
  obs <- 1*(obs<theta) ;
  
  # Conjugate
  
  r <- BernoulliModel_SufStat(obs,n) ;

  x_theta <- seq(from = 0, to = 1, length.out = 50) ;
  theta_prior_pdf <- BernoulliModel_PDF_prior(x_theta, a=a, b=b) ;
  theta_post_pdf <- BernoulliModel_PDF_posterior(x_theta, r=r,  n=n, a=a, b=b) ;
  
  x_pred <- c(0,1) ;
  xnew_pred_pdf <- c(BernoulliModel_PDF_pred(0, r=r,  n=n, a=a, b=b),
                     BernoulliModel_PDF_pred(1, r=r,  n=n, a=a, b=b));
  
  # Jeffreys
  
  theta_Jeffreys_prior_pdf <- BernoulliModel_PDF_Jeffreys_prior(x_theta) ;
  theta_Jeffreys_post_pdf <- BernoulliModel_PDF_Jeffreys_posterior(x_theta, r=r,  n=n) ;
  
  xnew_Jeffreys_pred_pdf <- c(BernoulliModel_PDF_Jeffreys_pred(0, r=r,  n=n),
                              BernoulliModel_PDF_Jeffreys_pred(1, r=r,  n=n));
  
  # Laplace
  
  theta_Laplace_prior_pdf <- BernoulliModel_PDF_Laplace_prior(x_theta) ;
  theta_Laplace_post_pdf <- BernoulliModel_PDF_Laplace_posterior(x_theta, r=r,  n=n) ;
  
  xnew_Laplace_pred_pdf <- c(BernoulliModel_PDF_Laplace_pred(0, r=r,  n=n),
                              BernoulliModel_PDF_Laplace_pred(1, r=r,  n=n));

  #layout( cbind( matrix(1,4,4), matrix(2,4,4), matrix(3,4,4) , matrix(4,4,4) ) )
  
  par(mfrow=c(2,2))
  
  # posterior
  
  y_max <- max(theta_post_pdf,theta_Jeffreys_post_pdf,theta_Laplace_post_pdf)
  
  plot(x_theta, 
       theta_post_pdf, 
       type = "l",  
       main='Posterior PDFs', 
       xlab = expression(theta),
       ylab = expression(paste(pi(theta,"|",x[1:n]))), 
       col="blue" ,
      xlim=c(0,1),
      ylim=c(0,y_max+1),
      cex.lab=fntsz, 
      cex.axis=fntsz, 
      cex.main=fntsz, 
      cex.sub=fntsz) ;
  
  lines(x_theta, 
        theta_Jeffreys_post_pdf,
        col= "black",
        cex.lab=fntsz, 
        cex.axis=fntsz, 
        cex.main=fntsz, 
        cex.sub=fntsz) ;
  
  lines(x_theta, 
        theta_Laplace_post_pdf,
        col= "green",
        cex.lab=fntsz, 
        cex.axis=fntsz, 
        cex.main=fntsz, 
        cex.sub=fntsz) ;
  
  abline(v=theta, 
         col="purple")
    
  legend('topright',
        c('Conjugate','Jeffrey','Laplace'),
        bg="transparent",
        lty=c(1,1),
        lwd=c(2.5,2.5),
        col=c('blue','black','green'),
        cex=fntsz
      )
  
  # predictive

  bardata <- data.frame(
                    thezero=c(xnew_pred_pdf[1],
                              xnew_Jeffreys_pred_pdf[1],
                              xnew_Laplace_pred_pdf[1]),
                    theone=c(xnew_pred_pdf[2],
                             xnew_Jeffreys_pred_pdf[2],
                             xnew_Laplace_pred_pdf[2])
                    )
  
  names(bardata) <-c("0","1")
  
  #attach(bardata)
  bardata <- as.matrix(bardata) ;
  
  barplot(bardata,
          main="Predictive PMF", 
          ylab = "x", 
          beside=TRUE, 
          ylim=c(0,1),
          col=c("blue", "black", "green"),
          cex.lab=fntsz, 
          cex.axis=fntsz, 
          cex.main=fntsz, 
          cex.sub=fntsz) ;
  
  legend("topright", 
         c('Conjugate','Jeffrey','Laplace'),
        bg="transparent",
        lwd=c(2.5,2.5),
        col=c('blue','black','green'),
        cex=fntsz,
        bty="n", 
        fill=c("blue", "black", "green")
        )

  # prior
    
  y_max <- c(theta_prior_pdf,theta_Jeffreys_prior_pdf,theta_Laplace_prior_pdf) ;
  y_max <- y_max[is.finite(y_max)]
    y_max <- max(y_max)
  
  plot(x_theta, 
       theta_prior_pdf, 
       type = "l",  
       main='Prior PDFs', 
       xlab = expression(theta),
       ylab = expression(paste(pi(theta,"|",x[1:n]))), 
       col="blue" ,
      xlim=c(0,1),
      ylim=c(0,y_max+1),
      cex.lab=fntsz, 
      cex.axis=fntsz, 
      cex.main=fntsz, 
      cex.sub=fntsz) ;
  
  lines(x_theta, 
        theta_Jeffreys_prior_pdf,
        col= "black",
        cex.lab=fntsz, 
        cex.axis=fntsz, 
        cex.main=fntsz, 
        cex.sub=fntsz) ;
  
  lines(x_theta, 
        theta_Laplace_prior_pdf,
        col= "green",
        cex.lab=fntsz, 
        cex.axis=fntsz, 
        cex.main=fntsz, 
        cex.sub=fntsz) ;
  
  abline(v=theta, 
         col="purple")
    
  legend('topright',
        c('Conjugate','Jeffrey','Laplace'),
        bg="transparent",
        lty=c(1,1),
        lwd=c(2.5,2.5),
        col=c('blue','black','green'),
        cex=fntsz
      )
  
  
    # plot(x_theta, 
    #    theta_prior_pdf, 
    #    type = "l",  
    #    main='Prior PDFs', 
    #    xlab = expression(theta),
    #    ylab = expression(paste(pi(theta,"|",x[1:n]))), 
    #    col="blue" ,
    #   xlim=c(0,1),
    #   cex.lab=fntsz, 
    #   cex.axis=fntsz, 
    #   cex.main=fntsz, 
    #   cex.sub=fntsz) ;
  
  
}




```


Inputs {.sidebar} 
-----------------------

***Parametric model parameters***

```{r,echo=FALSE}
sliderInput("Bern_theta",
                 "$$\\theta:$$",
                  min = 0.0001,
                  max = 0.9999,
                  step = 0.05,
                  value = 0.5)

sliderInput("Bern_n",
                 "$$n:$$",
                  min = 1,
                  max = 80,
                  step = 1,
                  value = 10)
```

***Prior hyper-parameters***

```{r,echo=FALSE}
sliderInput("Bern_a",
                 "$$a:$$",
                  min = 0.05,
                  max = 10.0,
                  step = 0.05,
                  value = 1)

sliderInput("Bern_b",
                 "$$b:$$",
                  min = 0.05,
                  max = 10.0,
                  step = 0.05,
                  value = 1)
```


Column {.tabset}
-----------------------


### ***Description*** 

We consider the Bayesian model

\[
\begin{cases}
x_{i}|\theta & \overset{\text{iid}}{\sim}\text{Br}(\theta),\,\,\forall i=1:n\\
\theta & \sim \pi(\text{d}\theta )
\end{cases}
\]

***Conjugate priors***

The priors are $\pi(\text{d}\theta)$ specified are:

\[
 \theta \sim \text{Be}(a,b)
\]

The posterior distribution for $\theta|x_{1:n}$ is: 
\begin{align*}
\theta|x_{1:n}
&\sim
\text{Be}(\sum_{i=1}^{n}x_{i}+a,n-\sum_{i=1}^{n}x_{i}+b)
\end{align*}

The predictive distribution for a future observation $x_{\text{new}}|x_{1:n}$ is

\begin{align*}
x_{\text{new}}|x_{1:n} 
 &\sim
 \text{Bb}(\sum_{i=1}^{n}x_{i}+a,n-\sum_{i=1}^{n}x_{i}+b,1) 
\end{align*}

***Jeffrey's prior's***

The priors are $\pi(\text{d}\theta)$ specified are:

\[
 \theta \sim \text{Be}(0.5,0.5)
\]

The posterior distribution for $\theta|x_{1:n}$ is: 
\begin{align*}
\theta|x_{1:n}
&\sim
\text{Be}(\sum_{i=1}^{n}x_{i}+0.5, n-\sum_{i=1}^{n}x_{i}+0.5)
\end{align*}

The predictive distribution for a future observation $x_{\text{new}}|x_{1:n}$ is

\begin{align*}
x_{\text{new}}|x_{1:n} 
 &\sim
 \text{Bb}(\sum_{i=1}^{n}x_{i}+0.5, n-\sum_{i=1}^{n}x_{i}+0.5,1) 
\end{align*}

***Laplace's prior's***

The priors are $\pi(\text{d}\theta)$ specified are:

\[
 \theta \sim \text{Un}(0,1) \text{  or equiv.  } \text{Be}(1,1)
\]

The posterior distribution for $\theta|x_{1:n}$ is: 
\begin{align*}
\theta|x_{1:n}
&\sim
\text{Be}(\sum_{i=1}^{n}x_{i}+1, n-\sum_{i=1}^{n}x_{i}+1)
\end{align*}

The predictive distribution for a future observation $x_{\text{new}}|x_{1:n}$ is

\begin{align*}
x_{\text{new}}|x_{1:n} 
 &\sim
 \text{Bb}(\sum_{i=1}^{n}x_{i}+1, n-\sum_{i=1}^{n}x_{i}+1,1) 
\end{align*}


### ***Parametric and Predictive inferece***

```{r,echo=FALSE}
renderPlot({
   plot_Bernoulli_parampred(theta=input$Bern_theta, 
                  n=input$Bern_n, 
                  a=input$Bern_a, 
                  b=input$Bern_b
                  )
  })
```



