<!-- -------------------------------------------------------------------------------- -->

<!-- Copyright 2019 Georgios Karagiannis -->

<!-- georgios.karagiannis@durham.ac.uk -->
<!-- Assistant Professor -->
<!-- Department of Mathematical Sciences, Durham University, Durham,  UK  -->

<!-- This file is part of Bayesian_Statistics (MATH3341/4031 Bayesian Statistics III/IV) -->
<!-- which is the material of the course (MATH3341/4031 Bayesian Statistics III/IV) -->
<!-- taught by Georgios P. Katagiannis in the Department of Mathematical Sciences   -->
<!-- in the University of Durham  in Michaelmas term in 2019 -->

<!-- Bayesian_Statistics is free software: you can redistribute it and/or modify -->
<!-- it under the terms of the GNU General Public License as published by -->
<!-- the Free Software Foundation version 3 of the License. -->

<!-- Bayesian_Statistics is distributed in the hope that it will be useful, -->
<!-- but WITHOUT ANY WARRANTY; without even the implied warranty of -->
<!-- MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the -->
<!-- GNU General Public License for more details. -->

<!-- You should have received a copy of the GNU General Public License -->
<!-- along with Bayesian_Statistics  If not, see <http://www.gnu.org/licenses/>. -->

<!-- -------------------------------------------------------------------------------- -->

---
title: "Distributions"
output: flexdashboard::flex_dashboard
runtime: shiny
---

<!-- Distributions Applet -->
<!-- Copyright (C) 2017  Georgios Karagiannis -->
<!-- georgios.karagiannis@durham.ac.uk -->

<!-- This program is free software: you can redistribute it and/or modify -->
<!-- it under the terms of the GNU General Public License as published by -->
<!-- the Free Software Foundation, either version 3 of the License, or -->
<!-- (at your option) any later version. -->

<!-- This program is distributed in the hope that it will be useful, -->
<!-- but WITHOUT ANY WARRANTY; without even the implied warranty of -->
<!-- MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the -->
<!-- GNU General Public License for more details. -->

<!-- You should have received a copy of the GNU General Public License -->
<!-- along with this program. If not, see <http://www.gnu.org/licenses/>. -->

# Bernoulli 

```{r,echo=FALSE}
plot_Bernoulli <- function(p=0.2) {
  
  fntsz <- 1.3 ;
  op <- par(cex = fntsz)
  
  x <- c(0,1) ;
  f_x <- c(1-p,p) ;
  F_x <- c(1-p,1) ;
  # F_x <- stepfun(x,c(F_x,1.0))
  
  par(mfrow=c(1,2))
  
  barplot( f_x,  names = x, xlab = "x", main = "Probability", ylab = "Pr(X=x)", ylim=c(0,1),
     cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz)
  
  # plot(F_x ,   main = "Cumulative probability", ylab = "Pr(X<=x)" , ylim=c(0,1), xlim=c(0.0,1.5))
  
  barplot( F_x,  names = x, xlab = "x", main = "Cumulative probability", ylab = "Pr(X<=x)", ylim=c(0,1),
     cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz)
  
}
```

Column {.sidebar}
-----------------------------------------------------------------------

***Bernoulli distribution***

$$X\sim\text{Bernoulli}(p)$$  $$p\in(0,1)$$

***Distribution parameters***

```{r,echo=FALSE}
sliderInput("p_Bern",
                 "$$p:$$",
                  min = 0.0,
                  max = 1,
                  value = 0.05)
```

Column
-----------------------------------------------------------------------


### ***Description*** {data-height=400}

Bernoulli bistribution describes a random experiment with exactly two possible outcomes, "success" (coded as 1) and "failure" (coded as 0), in which the probability of success $p$ is the same every time the experiment is conducted

***Probability:*** 
$$
\Pr(X=x|p)=
\begin{cases}
p^{x}(1-p)^{1-x}   & \text{, if } x\in \{0,1\} \\
0 & \text{, otherwise}
\end{cases}
\  \  
\text{  ,or  } 
\  \  
\Pr(X=x|p)=
\begin{cases}
1-p & \text{, if } x=0\\
p   & \text{, if } x=1 
\end{cases}
$$
***Expectation:*** 
$$
E(X) = p
$$
***Variance:*** 
$$
Var(X) = p(1-p)
$$

### ***Probability and Cummulative probability function plots*** {data-height=600}

```{r,echo=FALSE}
renderPlot({
    plot_Bernoulli(p=input$p_Bern)
  })
```





# Binomial 


```{r,echo=FALSE}
plot_binomial_withoutNorm <- function(n=10, p=0.2, Qcorrect=FALSE) {
  
  fntsz <- 1.4 ;
op <- par(cex = fntsz)

  
  xaxis_min <- 0.0 ;
  xaxis_max <- n ;
  
  x <- seq(xaxis_min,xaxis_max ,1) ;
  f_x <- dbinom(x, size=n, prob=p) ;
  F_x_val <- pbinom(x, size=n, prob=p) ;
  F_x <- stepfun(x,c(0.0,F_x_val))
    
  xaxis_norm_min = xaxis_min ;
  xaxis_norm_max = xaxis_max ;
  
  
  assum <- (n*p>10 && n*(1-p)>10)
  
  par(mfrow=c(1,2))
  
  plot(x, f_x, type = "n",  main = "Probability", ylab = "Pr(X=x)", xaxt = "n", ylim=c(0,1.1*max(f_x)) ,
cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz)
  boxplot(as.data.frame(rbind(0,f_x)), at = x, names = x,
          boxwex = 1, medlty = 0, add = TRUE, col = "lightgrey" ,
cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz)
  
  legend('topright',
         bg="transparent",
         c('Binomial'),
         lty=c(1),
         lwd=c(2.5),
         col=c('lightgrey'))
  
  plot(F_x,  main = "Cumulative probability",   ylab = "Pr(X<=x)" , ylim=c(0,1), xlim=c(0,n+1),col.ver = "bisque", xaxt='n',
cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz,cex=fntsz,
pch = 20)
  
  y2 <- rep(c(F_x_val), each=2)
  y2 <- y2[-length(y2)]
  x2 <- rep(x, each=2)[-1]
  x3 <- c(min(x2), x2, max(x2))
  x3[ length(x3)] <- n+1
  y3 <- c(0, y2, 1.0)
  
  polygon(c(x3,n+1), c(y3,0.0), border=NA, col="lightgrey",
cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz)
  
  plot(F_x,  main = "Cumulative probability",  ylab = "Pr(X<=x)" , ylim=c(0,1), xlim=c(0,n+1),col.ver = "bisque", add=TRUE,xaxt='n',
cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz,cex=fntsz,
pch = 20)
axis(side=1,at=x,cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz,cex=fntsz)
  
  x_norm <- seq(
            xaxis_min,
            xaxis_max,
            length.out=100) ;

  
  
}
```
```{r,echo=FALSE}
plot_binomial_withNorm <- function(n=10, p=0.2, Qcorrect=FALSE) {
  
  fntsz <- 1.4 ;
op <- par(cex = fntsz)
  
  xaxis_min <- 0.0 ;
  xaxis_max <- n ;
  
  x <- seq(xaxis_min,xaxis_max ,1) ;
  f_x <- dbinom(x, size=n, prob=p) ;
  F_x_val <- pbinom(x, size=n, prob=p) ;
  F_x <- stepfun(x,c(0.0,F_x_val))
  
  
 x_norm <- seq(
            xaxis_min,
            xaxis_max,
            length.out=100) ;
  
    
  xaxis_norm_min = xaxis_min ;
  xaxis_norm_max = xaxis_max ;
  
  
  assum <- (n*p>10 && n*(1-p)>10)
  
  par(mfrow=c(1,2))
  
  plot(x, f_x, type = "n",  main = "Probability", ylab = "Pr(X=x)", xaxt = "n", ylim=c(0,1.1*max(f_x)) ,
cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz)
  if (assum) 
    boxplot(as.data.frame(rbind(0,f_x)), at = x, names = x,
            boxwex = 1, medlty = 0, add = TRUE, col = "lightskyblue3",
cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz )
  else
    boxplot(as.data.frame(rbind(0,f_x)), at = x, names = x,
          boxwex = 1, medlty = 0, add = TRUE, col = "lightgrey" ,
cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz)
  
  curve(dnorm(x,mean=n*p,sd=sqrt(n*p*(1-p))), from = xaxis_min, to = xaxis_max, n = 101, add = TRUE, lwd=4, col='red',
cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz)
  
  if (assum) 
    legend('topright',
         bg="transparent",
         c('Binomial','Normal'),
         lty=c(1,1),
         lwd=c(2.5,2.5),
         col=c('lightskyblue3','red'))
    else
      legend('topright',
         bg="transparent",
             c('Binomial','Normal'),
             lty=c(1,1),
             lwd=c(2.5,2.5),
             col=c('grey','red'))
  
  plot(F_x,  main = "Cumulative probability",   ylab = "Pr(X<=x)" , ylim=c(0,1), xlim=c(0,n+1),col.ver = "bisque", xaxt='n',
cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz,cex=fntsz,
pch = 20)
  
  y2 <- rep(c(F_x_val), each=2)
  y2 <- y2[-length(y2)]
  x2 <- rep(x, each=2)[-1]
  x3 <- c(min(x2), x2, max(x2))
  x3[ length(x3)] <- n+1
  y3 <- c(0, y2, 1.0)
  
  if (assum) {
    polygon(c(x3,n+1), c(y3,0.0), border=NA, col="lightskyblue3",
cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz)
  }else {
    polygon(c(x3,n+1), c(y3,0.0), border=NA, col="lightgrey",
cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz)
  }
  plot(F_x,  main = "Cumulative probability",  ylab = "Pr(X<=x)" , ylim=c(0,1), xlim=c(0,n+1),col.ver = "bisque", add=TRUE, xaxt='n',
cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz, cex=fntsz,
pch = 20)
  
axis(side=1,at=x,cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz,cex=fntsz)

  x_norm <- seq(
            xaxis_min,
            xaxis_max,
            length.out=100) ;

  if (Qcorrect==FALSE) {
    F_x_norm <-  pnorm(x_norm, mean=n*p, sd=sqrt(n*p*(1-p)) )
  } else if (Qcorrect==TRUE){
    F_x_norm <-  pnorm(x_norm+0.5, mean=n*p, sd=sqrt(n*p*(1-p)) )
  }
  lines(x_norm ,F_x_norm, lwd=4,  col='red', ylim=c(0,1), xlim=c(0,n+1),
cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz)
  
  
}
```
```{r,echo=FALSE}
plot_binomial <- function(n=10, p=0.2, Qnorm=FALSE, Qcorrect=FALSE) {
  if(Qnorm) {
    plot_binomial_withNorm(n, p, Qcorrect) 
  } else {
    plot_binomial_withoutNorm(n, p, Qcorrect) 
  }
}
```



Column {.sidebar}
-----------------------------------------------------------------------

***Binomial distribution***

$$X\sim\text{bin}(n,p)$$ 
$$n\in\mathbb{N}_{+} , \ p\in [0,1]$$

***Distribution parameters***

```{r,echo=FALSE}
      sliderInput("n_bin",
                  "n: ",
                  min = 0.0,
                  max = 100.0,
                  value = 10.0,
                  step = 1.0 )
      sliderInput("p_bin",
                  "p :",
                  min = 0.0,
                  max = 1,
                  value = 0.05)
```

***Check Normal approximation***

```{r,echo=FALSE}
      checkboxInput("Qnorm_bin", 
                    label = "add Normal curve", 
                    value = FALSE)
```
                    

If the box is checked: 

  * the bars turn blue if the rull-of-thump assumptions $$np>10$$ and $$np(1-p)>10$$ for the normatilty approximation are satisfied.
  
***Apply correction in Normal approx.***

```{r,echo=FALSE}
      checkboxInput("Qcorrect_bin", 
                    label = "apply correction", 
                    value = FALSE)
```
                    

If the box is checked: 

  * It applies the correction in the Normal approximation CDF curve, such as: 
  
  $P_{\text{bin}}(X\le x)=$ $\   \   P_{\text{Normal}}(X\le x+0.5)$
  
  


Column
-----------------------------------------------------------------------

### ***Description*** {data-height=400}

  + The binomial distribution models the number of successes after reapeating a Bernulli experiment $n$ times. 
  + $X\sim\text{bin}(n,p)$ is the distribution of $X=\sum_{i=1}^{n}Y_i$, when $\{Y_i\}_{i=1}^n$ are IID random variables from  $Y_i \sim Bernoulli(p)$.

***Probability:*** 
$$
\Pr(X=x|n,p)=
\begin{cases}
\binom{n}{x}p^{x}(1-p)^{n-x}& \text{, if } x\in \{0,1,...,n\} \\
0& \text{, otherwise}  \\
\end{cases}
$$
***Expectation:*** 
$$
E(X) = np
$$
***Variance:*** 
$$
Var(X) = np(1-p)
$$

### ***Probability and Cummulative probability function plots***{data-height=600}

```{r,echo=FALSE}
renderPlot({
     plot_binomial(n=input$n_bin,
                   p=input$p_bin,
                   Qnorm=input$Qnorm_bin,
                   Qcorrect=input$Qcorrect_bin)
  })
```



# Uniform 

```{r,echo=FALSE}
plot_unif <- function(alpha= 0.0, beta= 1.0, axis_min=-2, axis_max=2) {
  
  fntsz <- 1.1 ;
op <- par(cex = fntsz)

  x <- seq(alpha, beta, length=100) ;
  
  f_x <- rep( 1/(beta-alpha), length=100) ;
  
  F_x <- punif(x, min = alpha, max = beta) ;
  
  par(mfrow=c(1,2))
  
  plot(c(axis_min,alpha,x,beta,axis_max), c(0.0,0.0,f_x,0.0,0.0), 
       type = "l",
       main="Probability density function",
       xlab="x",
       ylab=expression('f'['X']*'(x)'),
       xlim=c(axis_min,axis_max), 
       ylim=c(0.0, 2.0),
cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz
  )
  abline(a = 0, b = 0,col='red')
  
  plot(c(axis_min,alpha,x,beta,axis_max), c(0.0,0.0,F_x,1.0,1.0), 
       type = "l",
       main="Cumulative density function",
       xlab="x",
       ylab=expression('F'['X']*'(x)   =   Pr(X<=x)'),
       xlim=c(c(axis_min,axis_max)), 
       ylim=c(0.0, 1.0),
cex.lab=fntsz, cex.axis=fntsz, cex.main=fntsz, cex.sub=fntsz
  )
  abline(a = 0, b = 0,col='red')
  abline(a = 1, b = 0,col='red')
  
}
```

Column {.sidebar}
-----------------------------------------------------------------------

***Uniform distribution***

$$x\sim\text{U}(\alpha,\beta)$$ 
$$\alpha\in\mathbb{R},\  \beta\in\mathbb{R},\  \alpha<\beta$$

***Distribution parameters***

```{r,echo=FALSE}
      sliderInput("ab_unif",
                  "$$\\  \\  [\\  \\alpha  \\   , \\  \\beta \\  ]$$ ",
                  min = -2.0,
                  max = 2.0,
                  value = c(-0.0,1.0),
                  step = 0.05 )
```

<!-- ***Zoom in/out control*** -->

<!-- ```{r,echo=FALSE} -->
<!--       sliderInput("axis_max_exp", -->
<!--                   "Zoom axis x :", -->
<!--                   min = 10^(-10), -->
<!--                   max = 10.0, -->
<!--                   value = 5.0) -->
<!-- ``` -->

Column
-----------------------------------------------------------------------

### ***Description*** {data-height=400}

When somebody asks: "Give me a (continuous) random number from $\alpha$ to $\beta$", he means "Draw/Sample a value from $\text{U}(\alpha,\beta)$"

***Density function:*** 
$$
f_X(x|\alpha, \beta)=
\begin{cases}
\frac{1}{\beta-\alpha} & \text{, } x \in [\alpha,\beta] \\
0 & \text{, otherwise}
\end{cases}
$$
***Expectation:*** 
$$
E(X) = \frac{\beta+\alpha}{2}
$$
***Variance:*** 
$$
Var(X) = \frac{(\beta-\alpha)^2}{12}
$$


### ***Density and Cummulative probability function plots***{data-height=600}

```{r,echo=FALSE}
renderPlot({
    plot_unif(alpha=input$ab_unif[1],
              beta=input$ab_unif[2])
  })
```




# Normal 

```{r,echo=FALSE}
plot_normal <- function(mu= 0.0, sig2=1.0, xaxis_min=-10.0, xaxis_max=10.0, yaxis_max=1.0) {
  
  
  x <- seq(xaxis_min,xaxis_max,length=100)
  
  f_x <- dnorm(x, mean=mu, sd=sqrt(sig2)) ;
  
  F_x <- pnorm(x, mean=mu, sd=sqrt(sig2)) ;
  
  mydata <- data.frame(x, f_x, F_x) ;
  
  par(mfrow=c(1,2))
  
  plot(x, f_x, 
       type = "l",
       main="Probability density function",
       xlab="x",
       ylab=expression('f'['X']*'(x)'),
       xlim=c(xaxis_min,xaxis_max), 
       ylim=c(0.0, yaxis_max)
  )
  abline(a = 0, b = 0,col='red')
  
  plot(x, F_x, 
       type = "l",
       main="Cumulative density function",
       xlab="x",
       ylab=expression('F'['X']*'(x)   =   Pr(X<=x)'),
       xlim=c(xaxis_min,xaxis_max), 
       ylim=c(0.0, 1.0)
  )
  abline(a = 0, b = 0,col='red')
  abline(a = 1, b = 0,col='red')
  
}
```

Column {.sidebar}
-----------------------------------------------------------------------

***Normal distribution***

$$X\sim\text{N}(\mu,\sigma^2)$$ 
$$\mu\in\mathbb{R}, \ \sigma^2>0$$

***Distribution parameters***

```{r,echo=FALSE}
      sliderInput("mu_norm",
                  "$$\\mu:$$",
                  min = -6.0,
                  max = 6.0,
                  value = 0.0,
                  step = 0.05 )
      sliderInput("sig2_norm",
                  "$$\\sigma^2:$$",
                  min = 0.1,
                  max = 9.0,
                  value = 1.0,
                  step = 0.05) 
```

<!-- ***Zoom in/out control*** -->

<!-- ```{r,echo=FALSE} -->
<!--       sliderInput("xaxis_min_norm", -->
<!--                   "Zoom axis X (left) :", -->
<!--                   min = -10.0, -->
<!--                   max = 10.0, -->
<!--                   value = -4.0) -->
<!--       sliderInput("xaxis_max_norm", -->
<!--                   "Zoom axis X (right) :", -->
<!--                   min = -10.0, -->
<!--                   max = 10.0, -->
<!--                   value = 4.0)  -->
<!--       sliderInput("yaxis_max_norm", -->
<!--                   "Zoom axis Y :", -->
<!--                   min = 10^(-10), -->
<!--                   max = 1.2, -->
<!--                   value = 1.0, -->
<!--                   step = ((1.2)-(10^(-10)))/100 ) -->
<!-- ``` -->

Column
-----------------------------------------------------------------------


### ***Description*** {data-height=400}

***Density function:*** 
$$
f_X(x|\mu,\sigma^2)=
{\frac {1}{\sigma {\sqrt {2\pi }}}}\,e^{-{\frac {(x-\mu )^{2}}{2\sigma ^{2}}}}, \  \  x\in\mathbb{R}
$$
***Expectation:*** 
$$
E(X) = \mu
$$
***Variance:*** 
$$
Var(X) = \sigma^2
$$
***Properties:***

  * $X\sim\text{N}(\mu,\sigma^{2})$ $\Longleftrightarrow$ $\frac{X-\mu}{\sigma}\sim\text{N}(0,1)$
  * If $X\sim\text{N}(\mu,\sigma^{2})$ , $a,b\in\mathbb{R}$ and $Y=a+bX$, then $Y\sim\text{N}(a+b\mu,b^{2}\sigma^{2})$
  * If $X_{i}\overset{IID}{\sim}\text{N}(\mu_{i},\sigma_{i}^{2})$ $a\in\mathbb{R}$, $b_{i}\in\mathbb{R}$ for $i=1,...,N$, and $Y=a+\sum_{i=1}^{N}b_{i}X_{i}$, then $Y\sim\text{N}(a+\sum_{i=1}^{N}b_{i}\mu_{i},\sum_{i=1}^{N}b_{i}^{2}\sigma_{i}^{2})$.

***Central Limit Theorem (CLT):***

Suppose $X_{1},X_{2},...,X_{n}$ are IID random variables (with some distribution) with $E(X_{i})=\mu$ and $Var(X_{i})=\sigma^{2}<\infty$. Let $\bar{X}=\frac{1}{n}\sum_{i=1}^{n}X_{i}$. Then, as $n$ approaches infinity:
$$
\sqrt{n}\frac{\bar{X}-\mu}{\sigma}\overset{\text{approx.}}{\sim}\text{N}(0,1).
$$
***Implementation to statistics (the quantities):***

  * If random sample $X_{1},X_{2},...,X_{n}\overset{IID}{\sim}\text{N}(\mu,\sigma^{2})$,
then 
$$
 \sqrt{n} \frac{\bar{X}-\mu}{ \sigma }  \sim \text{N}(0,1)
$$
 where $\bar{X}=\frac{1}{n}\sum_{i=1}^{n}X_{i}$.
 
  * If random sample $X_{1},X_{2},...,X_{n}\overset{IID}{\sim}\text{N}(\mu,\sigma^{2})$,
then  
$$
 (n-1)\frac{S^{2}}{\sigma^{2}}\sim\chi_{n-1}^{2}
$$
 where $S^{2}=\frac{1}{n-1}\sum_{i=1}^{n}(X_{i}-\bar{X})^{2}$.
 
  * If random sample $X_{1},X_{2},...,X_{n}\overset{IID}{\sim}\text{N}(\mu,\sigma^{2})$,
then  
$$
 \frac{\bar{X}-\mu}{\sqrt{S^{2}/n}}\sim t_{n-1}
$$
 where $S^{2}=\frac{1}{n-1}\sum_{i=1}^{n}(X_{i}-\bar{X})^{2}$ and $\bar{X}=\frac{1}{n}\sum_{i=1}^{n}X_{i}$.
 
 * Consider a random sample $X_{1},X_{2},...,X_{n}\overset{IID}{\sim}\text{N}(\mu_X,\sigma^{2})$, and a random sample $Y_{1},Y_{2},...,Y_{m}\overset{IID}{\sim}\text{N}(\mu_Y,\sigma^{2})$. The two random samples are independent each other. 
Then  
$$
 \frac{(\bar{X}-\bar{Y})-(\mu_X-\mu_Y)}{S_p\sqrt{\frac{1}{n}+\frac{1}{m}}}\sim t_{n+m-2}
$$
 where $S_p^2=\frac{(n-1)S_X^{2}+(m-1)S_Y^{2}}{n+m-2}$, $S_X^{2}=\frac{1}{n-1}\sum_{i=1}^{n}(X_{i}-\bar{X})^{2}$, $\bar{X}=\frac{1}{n}\sum_{i=1}^{n}X_{i}$, $S_Y^{2}=\frac{1}{m-1}\sum_{i=1}^{m}(Y_{i}-\bar{Y})^{2}$ and $\bar{Y}=\frac{1}{m}\sum_{i=1}^{m}Y_{i}$.
 
 * If $X_{1},X_{2},...,X_{n}\overset{IID}{\sim}\text{N}(\mu,\sigma^{2})$, then 
$$
 (n-1)\frac{S^{2}}{\sigma^{2}} \sim\chi_{n-1}^{2} 
$$
where $S^{2}=\frac{1}{n-1}\sum_{i=1}^{n}(X_{i}-\bar{X})^{2}$.  
[by Cochran's theorem] 
 
   * Consider a random sample $X_{1},X_{2},...,X_{n}\overset{IID}{\sim}\text{N}(\mu_X,\sigma^{2})$, and 
   a random sample random sample $Y_{1},Y_{2},...,Y_{M}\overset{IID}{\sim}\text{N}(\mu_Y,\sigma^{2})$. The two random samples are independent each other. Then it is: 
$$
 (n+m-2)\frac{S_p^{2}}{\sigma^{2}} \sim\chi^{2}_{n+m-2} 
$$ 
where $S_p^2$ is the pool sample variance defined as $S_p^2=\frac{(n-1)S_X^{2}+(m-1)S_Y^{2}}{n+m-2}$, with $S_X^{2}=\frac{1}{n-1}\sum_{i=1}^{n}(X_{i}-\bar{X})^{2}$, and $S_Y^{2}=\frac{1}{m-1}\sum_{i=1}^{m}(Y_{i}-\bar{Y})^{2}$.  
    
    + This result can be generalized to consider more than 2 random samples. (Applications to ANOVA)
 
  * If random samples $X_{1},X_{2},...,X_{n_{X}}\overset{IID}{\sim}\text{N}(\mu_{X},\sigma_{X}^{2})$, and $Y_{1},Y_{2},...,Y_{n_{Y}}\overset{IID}{\sim}\text{N}(\mu_{Y},\sigma_{Y}^{2})$, and if  
$$
F=\frac{S_{Y}^{2}/\sigma_{Y}^{2}}{S_{X}^{2}/\sigma_{X}^{2}} 
$$
where $S_{X}^{2}=\frac{1}{n_{X}-1}\sum_{i=1}^{n_{X}}(X_{i}-\bar{X})^{2}$, $\bar{X}=\frac{1}{n_{X}}\sum_{i=1}^{n_{X}}X_{i}$,  $S_{Y}^{2}=\frac{1}{n_{Y}-1}\sum_{i=1}^{n_{Y}}(Y_{i}-\bar{Y})^{2}$ and $\bar{Y}=\frac{1}{n_{Y}}\sum_{i=1}^{n_{Y}}Y_{i}$,  
then 
$$
F \sim F_{n_{Y}-1,n_{X}-1}
$$

### ***Density and Cummulative probability function plots*** {data-height=600}

```{r,echo=FALSE}
renderPlot({
    plot_normal(mu=input$mu_norm,
                sig2=input$sig2_norm)
  })
```



# Student t 

```{r,echo=FALSE}
plot_student_t_withoutNorm <- function(df=2.0, xaxis_max=12, yaxis_max=0.45) {
  
  xaxis_min <- -xaxis_max ;
  
  x <- seq(xaxis_min,xaxis_max,length=100)
  
  f_x <- dt(x, df=df) ;
  
  F_x <- pt(x, df=df) ;
  
  par(mfrow=c(1,2))
  
  plot(x, f_x, 
       type = "l",
       main="Probability density function",
       xlab="x",
       ylab=expression('f'['X']*'(x)'),
       xlim=c(xaxis_min,xaxis_max), 
       ylim=c(0.0, yaxis_max)
  )
  
  legend('topright',
         bg="transparent",
         c('t'),
         lty=c(1,1),
         lwd=c(2.5,2.5),
         col=c('black')) 
  
  abline(a = 0, b = 0,col='red')
  plot(x, F_x, 
       type = "l",
       main="Cumulative density function",
       xlab="x",
       ylab=expression('F'['X']*'(x)   =   Pr(X<=x)'),
       xlim=c(xaxis_min,xaxis_max), 
       ylim=c(0.0, 1.0)
  )
  abline(a = 0, b = 0,col='red')
  abline(a = 1, b = 0,col='red')
}
```

```{r,echo=FALSE}
plot_student_t_withNorm <- function(df=2.0, xaxis_max=12, yaxis_max=0.45) {
  
  xaxis_min <- -xaxis_max ;
  
  x <- seq(xaxis_min,xaxis_max,length=100)
  
  f_x <- dt(x, df=df) ;
  
  F_x <- pt(x, df=df) ;
  
  f_x_norm <- dnorm(x,0,1) ;
  
  F_x_norm <- pnorm(x,0,1) ;
  
  par(mfrow=c(1,2))
  
  plot(x, f_x, 
       type = "l",
       main="Probability density function",
       xlab="x",
       ylab="f(x)",
       xlim=c(xaxis_min,xaxis_max), 
       ylim=c(0.0, yaxis_max)
  )
  
    lines(x,f_x_norm,col="blue")
    legend('topright',
         bg="transparent",
           c('t','N(0,1)'),
           lty=c(1,1),
           lwd=c(2.5,2.5),
           col=c('black','blue')) 

  
  abline(a = 0, b = 0,col='red')
  plot(x, F_x, 
       type = "l",
       main="Cumulative distribution function",
       xlab="x",
       ylab="F(x)=Pr(X<=x)",
       xlim=c(xaxis_min,xaxis_max), 
       ylim=c(0.0, 1.0)
  )
  lines(x,F_x_norm,col="blue")
  abline(a = 0, b = 0,col='red')
  abline(a = 1, b = 0,col='red')
}
```

```{r,echo=FALSE}
plot_student_t <- function(df=2.0, xaxis_max=5, yaxis_max=0.45, Qnorm=FALSE) {
  if(Qnorm) {
    plot_student_t_withNorm(df, xaxis_max, yaxis_max) 
    } else {
    plot_student_t_withoutNorm(df, xaxis_max, yaxis_max) 
  }
}
```

Column {.sidebar}
-----------------------------------------------------------------------

***Student t distribution***

$$X\sim\text{t}_{d}$$ 
$$d>0$$

***Distribution parameters***

```{r,echo=FALSE}
      sliderInput("d_t",
                  "$$d: $$",
                  min = 1.0,
                  max = 50.0,
                  value = 1.0,
                  step = 0.1 )
```

***Compare with $\text{N}(0,1)$***

```{r,echo=FALSE}
      checkboxInput("Qnorm_t", 
                    label = "add Normal curve", 
                    value = FALSE)
```

<!-- ***Zoom in/out control*** -->

<!-- ```{r,echo=FALSE} -->
<!--       sliderInput("xaxis_max", -->
<!--                   "Zoom axis X :", -->
<!--                   min = 0.0, -->
<!--                   max = 10.0, -->
<!--                   value = 4.0) -->
<!--       sliderInput("yaxis_max", -->
<!--                   "Zoom axis Y :", -->
<!--                   min = 10^(-10), -->
<!--                   max = 0.5, -->
<!--                   value = 1.0, -->
<!--                   step = ((0.5)-(10^(-10)))/100 ) -->
<!-- ``` -->


Column
-----------------------------------------------------------------------

### ***Description*** {data-height=400}

***Density function:*** 
$$
f_X(x|d)=
\textstyle\frac{\Gamma \left(\frac{d+1}{2} \right)} {\sqrt{d\pi}\,\Gamma \left(\frac{d}{2} \right)} \left(1+\frac{x^2}{d} \right)^{-\frac{d+1}{2}} , \  \  x\in\mathbb{R}
$$
***Expectation:*** 
$$
E(X) = 0
$$
***Variance:*** 
$$
Var(X) = 
\begin{cases}
\infty & , 1<d\le  2  \\
\frac{d}{d-2} &, d>2  \\
\text{undefined} &\text{, otherwise } 
\end{cases}
$$
***Comments:***

   * Distribution $t_{d}$ can be derived as the distribution of $X:=\frac{Z}{\sqrt{Y/d}}$ when $Z\sim\text{N}(0,1)$, and $Y\sim\chi_{d}^{2}$.


***Properties:***

   * PDF of $X\sim\text{t}_d$ is simetric around $0$; 
      + hence $P(X<-x)=P(X>x)$
   * PDF & CDF of $\text{t}_d$ have a similar shape to a standard normal $\text{N}(0,1)$ 
   * PDF of $\text{t}_d$ has "fatter tails"" than  $\text{N}(0,1)$    
      + more probability is assigned above areas farther from the centre (on the tails)
      + it has greater spread than a standard normal 
      + the tails getting fatter when $d$ gets smaller 
   * In the limit, it is:  $\lim_{d\rightarrow\infty}  \text{t}_d =\text{N}(0,1)$ 
   * We use tables (or a computer) to find probabilities for $\text{t}_d$.

***Implementation to statistics (the quantities):***

   * If $X_{1}, \ X_{2}, ..., \ X_{n}\overset{IID}{\sim}\text{N}(\mu,\sigma^{2})$, then   
$$
 \sqrt{n}\frac{\bar{X}-\mu}{\sqrt{S^{2}}} \sim t_{n-1}
$$ 
where $S^{2}=\frac{1}{n-1}\sum_{i=1}^{n}(X_{i}-\bar{X})^{2}$ and $\bar{X}=\frac{1}{n}\sum_{i=1}^{n}X_{i}$.

   * Consider a random sample $X_{1},X_{2},...,X_{n}\overset{IID}{\sim}\text{N}(\mu_X,\sigma^{2})$, and a random sample $Y_{1},Y_{2},...,Y_{m}\overset{IID}{\sim}\text{N}(\mu_Y,\sigma^{2})$. The two random samples are independent each other. Then  
$$
 \frac{(\bar{X}-\bar{Y})-(\mu_X-\mu_Y)}{S_p\sqrt{\frac{1}{n}+\frac{1}{m}}}\sim t_{n+m-2}
$$
 where $S_p^2=\frac{(n-1)S_X^{2}+(m-1)S_Y^{2}}{n+m-2}$, $S_X^{2}=\frac{1}{n-1}\sum_{i=1}^{n}(X_{i}-\bar{X})^{2}$, $\bar{X}=\frac{1}{n}\sum_{i=1}^{n}X_{i}$, $S_Y^{2}=\frac{1}{m-1}\sum_{i=1}^{m}(Y_{i}-\bar{Y})^{2}$ and $\bar{Y}=\frac{1}{m}\sum_{i=1}^{m}Y_{i}$.


### ***Density and Cummulative probability function plots*** {data-height=600}

```{r,echo=FALSE}
renderPlot({
    plot_student_t(df=input$d_t,
                   Qnorm=input$Qnorm_t)
  })
```







# Exponential 

```{r,echo=FALSE}
plot_exp <- function(ell= 0.5, axis_max=11) {
  
  
  x <- seq(10^(-10),axis_max,length=100)
  
  f_x <- dexp(x, rate = ell) ;
  
  F_x <- pexp(x, rate = ell) ;
  
  par(mfrow=c(1,2))
  
  plot(x, f_x, 
       type = "l",
       main="Probability density function",
       xlab="x",
       ylab=expression('f'['X']*'(x)'),
       xlim=c(0.0,axis_max), 
       ylim=c(0.0, 2.0)
  )
  abline(a = 0, b = 0,col='red')
  
  plot(x, F_x, 
       type = "l",
       main="Cumulative distribution function",
       xlab="x",
       ylab=expression('F'['X']*'(x)   =   Pr(X<=x)'),
       xlim=c(0.0,axis_max), 
       ylim=c(0.0, 1.0)
  )
  abline(a = 0, b = 0,col='red')
  abline(a = 1, b = 0,col='red')
  
}
```

Column {.sidebar}
-----------------------------------------------------------------------

***Exponential distribution***

$$x\sim\mathcal{E}(\lambda)$$ 
$$\lambda>0$$

***Distribution parameters***

```{r,echo=FALSE}
      sliderInput("ell_exp",
                  "$$\\lambda :$$ ",
                  min = 0.1,
                  max = 3.0,
                  value = 0.5,
                  step = (3-0)/100 )
```

<!-- ***Zoom in/out control*** -->

<!-- ```{r,echo=FALSE} -->
<!--       sliderInput("axis_max_exp", -->
<!--                   "Zoom axis x :", -->
<!--                   min = 10^(-10), -->
<!--                   max = 10.0, -->
<!--                   value = 5.0) -->
<!-- ``` -->

Column
-----------------------------------------------------------------------

### ***Description*** {data-height=400}

***Density function:*** 
$$
f_X(x|\lambda)=
\begin{cases}
\lambda e^{-\lambda x} & \text{, } x>0 \\
0 & \text{, otherwise}
\end{cases}
$$
***Expectation:*** 
$$
E(X) = \frac{1}{\lambda}
$$
***Variance:*** 
$$
Var(X) = \frac{1}{\lambda^2}
$$


### ***Density and Cummulative probability function plots***{data-height=600}

```{r,echo=FALSE}
renderPlot({
    plot_exp(ell=input$ell_exp)
  })
```


# Chi squared 

```{r,echo=FALSE}
plot_chi_sq <- function(df=1.0, xaxis_max=70.0, yaxis_max=0.36) {
  
  xaxis_min <- 10^(-5) ;
  
  x <- seq(xaxis_min,xaxis_max,length=100)
  
  f_x <- dchisq(x, df=df) ;
  
  F_x <- pchisq(x, df=df) ;
  
  par(mfrow=c(1,2))
  
  plot(x, f_x, 
       type = "l",
       main="Probability density function",
       xlab="x",
       ylab=expression('f'['X']*'(x)'),
       xlim=c(0.0,xaxis_max), 
       ylim=c(0.0, yaxis_max)
  )
  abline(a = 0, b = 0,col='red')
  
  plot(x, F_x, 
       type = "l",
       main="Cumulative distribution function",
       xlab="x",
       ylab=expression('F'['X']*'(x)   =   Pr(X<=x)'),
       xlim=c(0.0,xaxis_max), 
       ylim=c(0.0, 1.0)
  )
  abline(a = 0, b = 0,col='red')
  abline(a = 1, b = 0,col='red')
}
```

Column {.sidebar}
-----------------------------------------------------------------------

***Chi squared***

$$X\sim\chi^2_d$$
$$d>0 $$

***Distribution parameters***

```{r,echo=FALSE}
      sliderInput("df_chisq",
                  "$$d: $$",
                  min = 1.0,
                  max = 50.0,
                  value = 1.0,
                  step = 0.1 ) 
```

Column
-----------------------------------------------------------------------

### ***Description*** {data-height=400}

***Density function:*** 
$$
f_X(x|d)=
\begin{cases}
{\frac {1}{2^{\frac {d}{2}}\Gamma \left({\frac {d}{2}}\right)}}\;x^{{\frac {d}{2}}-1}e^{-{\frac {x}{2}}}\,& \text{, } x>0 \\
0 & \text{, otherwise}
\end{cases}
$$
***Expectation:*** 
$$
E(X) = d
$$
***Variance:*** 
$$
Var(X) = 2d
$$
***Comments:***

   * Distribution $\chi_{d}^{2}$ can be derived as the distribution of $X:=\sum_{i=1}^{d}Z_{i}^{2}$ when $Z_{i}\overset{IID}{\sim}\text{N}(0,1)$, for $i=1,...,d$.  
   
   * Consider independent random variables $Y_{i}\sim\chi_{d_i}^2$ for $i=1,...,n$. Let random variable $X:=\sum_{i=1}^{n}Y_{i}$. Then $X$ follows distribution $$X\sim\chi_d^2,$$ where $d=\sum_{i=1}^{n}d_i$.

***Implementation to statistics (the quantities):***

   * If $X_{1},X_{2},...,X_{n}\overset{IID}{\sim}\text{N}(\mu,\sigma^{2})$, then 
$$
 (n-1)\frac{S^{2}}{\sigma^{2}} \sim\chi_{n-1}^{2} 
$$
where $S^{2}=\frac{1}{n-1}\sum_{i=1}^{n}(X_{i}-\bar{X})^{2}$.  
[by Cochran's theorem]  

   * Consider a random sample $X_{1},X_{2},...,X_{n}\overset{IID}{\sim}\text{N}(\mu_X,\sigma^{2})$, and 
   a random sample random sample $Y_{1},Y_{2},...,Y_{M}\overset{IID}{\sim}\text{N}(\mu_Y,\sigma^{2})$. The two random samples are independent each other. Then it is: 
$$
 (n+m-2)\frac{S_p^{2}}{\sigma^{2}} \sim\chi^{2}_{n+m-2} 
$$ 
where $S_p^2$ is the pool sample variance defined as $S_p^2=\frac{(n-1)S_X^{2}+(m-1)S_Y^{2}}{n+m-2}$, with $S_X^{2}=\frac{1}{n-1}\sum_{i=1}^{n}(X_{i}-\bar{X})^{2}$, and $S_Y^{2}=\frac{1}{m-1}\sum_{i=1}^{m}(Y_{i}-\bar{Y})^{2}$.  
    
    + This result can be generalized to consider more than 2 random samples. (Applications to ANOVA)

### ***Density and Cummulative probability function plots*** {data-height=600}

```{r,echo=FALSE}
renderPlot({
    plot_chi_sq(df=input$df_chisq)
  })
```



# F

```{r,echo=FALSE}
plot_F <- function(df1=1.0, df2=1.0, xaxis_max=10.0, yaxis_max=1.0) {
  
  xaxis_min <- 10^(-5) ;
  
  x <- seq(xaxis_min,xaxis_max,length=100)
  
  f_x <- df(x, df1=df1, df2=df2) ;
  
  x_cdf <- seq(xaxis_min,5*xaxis_max,length=100)
  
  F_x <- pf(x_cdf, df1=df1, df2=df2) ;
  
  par(mfrow=c(1,2))
  
  plot(x, f_x, 
       type = "l",
       main="Probability density function",
       xlab="x",
       ylab=expression('f'['X']*'(x)'),
       xlim=c(0.0,xaxis_max), 
       ylim=c(0.0, yaxis_max)
  )
  abline(a = 0, b = 0,col='red')
  
  plot(x_cdf, F_x, 
       type = "l",
       main="Cumulative distribution function",
       xlab="x",
       ylab=expression('F'['X']*'(x)   =   Pr(X<=x)'),
       xlim=c(0.0,5*xaxis_max), 
       ylim=c(0.0, 1.0)
  )
  abline(a = 0, b = 0,col='red')
  abline(a = 1, b = 0,col='red')
  
}
```

Column {.sidebar}
-----------------------------------------------------------------------

***F distribution***

$$x\sim F_{d_1,d_2}$$
$$d_1>0, \ d_2>0$$

***Distribution parameters***

```{r,echo=FALSE}
      sliderInput("df1_F",
                  "$$d_1:$$ ",
                  min = 0.5,
                  max = 50,
                  value = 1.0,
                  step = 0.1 ) 
      sliderInput("df2_F",
                  "$$d_2:$$ ",
                  min = 0.5,
                  max = 50,
                  value = 1.0,
                  step = 0.1 )
```

Column
-----------------------------------------------------------------------

### ***Description*** {data-height=400}

***Density function:*** 
$$
f_X(x|d_1,d_2)=
\begin{cases}
{\frac  {1}{x\,{\mathrm  {B}}\!\left({\frac  {d_{1}}{2}},{\frac  {d_{2}}{2}}\right)}} {\sqrt  {{\frac  {(d_{1}\,x)^{{d_{1}}}\,\,d_{2}^{{d_{2}}}}{(d_{1}\,x+d_{2})^{{d_{1}+d_{2}}}}}}}& \text{, } x>0 \\
0 & \text{, otherwise}
\end{cases}
$$
***Expectation:*** 
$$
E(X) = \frac{d_2}{d_2-2}, d_2>2
$$
***Variance:*** 
$$
Var(X) = {\frac  {2\,d_{2}^{2}\,(d_{1}+d_{2}-2)}{d_{1}(d_{2}-2)^{2}(d_{2}-4)}},\ \ d_2>4
$$
***Comments:***  

   * $F_{d_{1},d_{2}}$ can be derived as the distribution of $X:=\frac{Y_{1}/d_{1}}{Y_{2}/d_{2}}$, when $Y_{1}\sim\chi_{d_{1}}^{2}$, $Y_{2}\sim\chi_{d_{2}}^{2}$.  

   * If $X\sim F_{n,m}$ then $1/X\sim F_{m,n}$  

   * About the quantiles: $$F_{n,m,a}=\frac{1}{F_{m,n,1-a}}$$  
   
***Implementation to statistics (the quantities):***

  * If random samples $X_{1},X_{2},...,X_{n_X}\overset{IID}{\sim}\text{N}(\mu_{X},\sigma_X^{2})$, and $Y_{1},Y_{2},...,Y_{n_Y}\overset{IID}{\sim}\text{N}(\mu_{Y},\sigma_Y^{2})$, then 
$$
 \frac{S_{Y}^{2}/\sigma_Y^{2}}{S_{X}^{2}/\sigma_X^{2}} \sim F_{n_{Y}-1,n_{X}-1}
$$
where $S_{X}^{2}=\frac{1}{n_X-1}\sum_{i=1}^{n_X}(X_{i}-\bar{X})^{2}$, $\bar{X}=\frac{1}{n_X}\sum_{i=1}^{n_X}X_{i}$, $S_{Y}^{2}=\frac{1}{n_Y-1}\sum_{i=1}^{n_Y}(Y_{i}-\bar{Y})^{2}$ and $\bar{Y}=\frac{1}{n_Y}\sum_{i=1}^{n_Y}Y_{i}$.

### ***Density and Cummulative probability function plots***{data-height=600}

```{r,echo=FALSE}
renderPlot({
    plot_F(df1=input$df1_F,
           df2=input$df2_F)
  })

```