<!-- -------------------------------------------------------------------------------- -->

<!-- Copyright 2019 Georgios Karagiannis -->

<!-- georgios.karagiannis@durham.ac.uk -->
<!-- Assistant Professor -->
<!-- Department of Mathematical Sciences, Durham University, Durham,  UK  -->

<!-- This file is part of Bayesian_Statistics (MATH3341/4031 Bayesian Statistics III/IV) -->
<!-- which is the material of the course (MATH3341/4031 Bayesian Statistics III/IV) -->
<!-- taught by Georgios P. Katagiannis in the Department of Mathematical Sciences   -->
<!-- in the University of Durham  in Michaelmas term in 2019 -->

<!-- Bayesian_Statistics is free software: you can redistribute it and/or modify -->
<!-- it under the terms of the GNU General Public License as published by -->
<!-- the Free Software Foundation version 3 of the License. -->

<!-- Bayesian_Statistics is distributed in the hope that it will be useful, -->
<!-- but WITHOUT ANY WARRANTY; without even the implied warranty of -->
<!-- MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the -->
<!-- GNU General Public License for more details. -->

<!-- You should have received a copy of the GNU General Public License -->
<!-- along with Bayesian_Statistics  If not, see <http://www.gnu.org/licenses/>. -->

<!-- -------------------------------------------------------------------------------- -->


---
title: "Mixture priors/posteriors"
output: flexdashboard::flex_dashboard
runtime: shiny
---

<!-- Mixture priors Applet -->
<!-- Copyright (C) 2017  Georgios Karagiannis -->
<!-- georgios.karagiannis@durham.ac.uk -->

<!-- This program is free software: you can redistribute it and/or modify -->
<!-- it under the terms of the GNU General Public License as published by -->
<!-- the Free Software Foundation, either version 3 of the License, or -->
<!-- (at your option) any later version. -->

<!-- This program is distributed in the hope that it will be useful, -->
<!-- but WITHOUT ANY WARRANTY; without even the implied warranty of -->
<!-- MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the -->
<!-- GNU General Public License for more details. -->

<!-- You should have received a copy of the GNU General Public License -->
<!-- along with this program. If not, see <http://www.gnu.org/licenses/>. -->

```{r}
get_randnumb <-function(n) {
  u <- seq(0.001,0.999,length.out = n)
#  u <- runif(0,1,n = n)
}
```

```{r}
#library(rmutil)
dbetabinom <- function (y, size, m, s, log = FALSE) {
  # from rmutil package
    if (any(y < 0)) 
        stop("y must contain non-negative values")
    if (any(size < 0)) 
        stop("size must contain non-negative values")
    if (any(m <= 0) || any(m >= 1)) 
        stop("m must lie between 0 and 1")
    if (any(s <= 0)) 
        stop("s must be positive")
    ly <- max(length(y), length(m), length(s), length(size))
    if (length(y) != ly) {
        if (length(y) == 1) 
            y <- rep(y, ly)
        else stop("length of y incorrect")
    }
    if (length(size) != ly) {
        if (length(size) == 1) 
            size <- rep(size, ly)
        else stop("size must be the same length as y")
    }
    if (any(y > size)) 
        stop("y must be <= size")
    if (length(m) != ly) {
        if (length(m) == 1) 
            m <- rep(m, ly)
        else stop("m and y must have the same length")
    }
    if (length(s) != ly) {
        if (length(s) == 1) 
            s <- rep(s, ly)
        else stop("s and y must have the same length")
    }
    t <- s * m
    u <- s * (1 - m)
    tmp <- lbeta(y + t, size - y + u) - lbeta(t, u) + lchoose(size, 
        y)
    if (!log) 
        tmp <- exp(tmp)
    tmp
}

BinomialBeta_pdf<-function(x,a,b,n) {
  pdf <- dbetabinom(y=x, size=n, m=a/(a+b), s=a+b) ;
  return (pdf) 
}
```


```{r}
StudentT_pdf<-function(x,mu,lam,al) {
  pdf <- 0.5*log(lam) -0.5*log(al) -0.5*log(pi) ;
  pdf <- pdf +lgamma(0.5*(al+1)) -lgamma(0.5*al) ;
  pdf <- pdf -0.5*(al+1)*log(1+lam/al*(x-mu)^2) ;
  pdf <- exp(pdf) ;
  return (pdf)  
}
```


```{r}
dinvgamma <- function(x, a, b){
  return ( exp( a*log(b)-lgamma(a) -(a+1)*log(x) -b/x ) ) ;
}
```



# Bernoulli model


```{r,echo=FALSE}
BernoulliModel_SufStat<-function(x,n) {
 r <- sum(x[1:n]) ; 
}

BernoulliModel_PDF_prior<-function(theta,a,b) {
  pdf <- dbeta(theta, a, b) ;
   return (pdf)  
}

BernoulliModel_PDF_posterior<-function(theta,r,n,a,b) {
  pdf <- dbeta(theta, a+r, b+n-r) ;
   return (pdf)  
}

BernoulliModel_PDF_pred<-function(y,r,n,a,b) {
  pdf <- BinomialBeta_pdf(y, a+r, b+n-r, 1) ;
   return (pdf)  
}

BernoulliModel_PDF_mixprior<-function(theta,Kmax,w,a,b) {
  pdf <- w[1]*dbeta(theta, a[1], b[1])
   for (k in 2:Kmax)
      pdf <- pdf + w[k]*dbeta(theta, a[k], b[k]) ;
   return (pdf)  
}

BernoulliModel_PDF_mixposterior<-function(theta,r,n,Kmax,w,a,b) {
  a_star = a+r ;
  b_star = b+n-r ;
  w_star = w * beta(a_star,b_star) / beta(a,b) ;
  w_star = w_star/sum(w_star) ;
  pdf <- w_star[1]*dbeta(theta, a_star[1], b_star[1])
  for (k in 2:Kmax)
    pdf <- pdf + w_star[k]*dbeta(theta, a_star[k], b_star[k]) ;
   return (pdf)  
}

BernoulliModel_PDF_mixpred<-function(y,r,n,Kmax,w,a,b) {
  a_star = a+r ;
  b_star = b+n-r ;
  w_star = w * beta(a_star,b_star) / beta(a,b) ;
  w_star = w_star/sum(w_star) ;
    pdf <- w_star[1]*BinomialBeta_pdf(y, a_star[1], b_star[1], 1)
  for (k in 2:Kmax)
    pdf <- pdf + w_star[k]*BinomialBeta_pdf(y, a_star[k], b_star[k], 1) ;
  return (pdf)  
}

plot_Bernoulli_param <- function(theta=0.5, n=10, 
                                 Kmax=2, 
                                 w1=0.5, 
                                 a1=10, a2=15, 
                                 b1=20, b2=15) {
  
# theta=0.5
# n=10
# Kmax=2
# w1=0.5
# w2=0.2
# a1=10
# a2=20
# b1=20
# b2=10
  
  w2 = 1-w1 ;
  
  w = c(w1,w2)
  a = c(a1,a2)
  b = c(b1,b2)
  
  fntsz <- 1.5;
  op <- par(cex = fntsz)
  
  obs <-get_randnumb(n) ;
  obs <- 1*(obs<theta) ;
  
  r <- BernoulliModel_SufStat(obs,n) ;

  x_theta <- seq(from = 0, to = 1, length.out = 50) ;
  theta_prior_pdf <- BernoulliModel_PDF_mixprior(x_theta, Kmax=Kmax, w=w, a=a, b=b) ;
  theta_post_pdf <- BernoulliModel_PDF_mixposterior(x_theta, r=r,  n=n, Kmax=Kmax, w=w, a=a, b=b);
  
  x_pred <- c(0,1) ;
  xnew_pred_pdf <- c(BernoulliModel_PDF_mixpred(0, r=r,  n=n, Kmax=Kmax, w=w, a=a, b=b),
                     BernoulliModel_PDF_mixpred(1, r=r,  n=n, Kmax=Kmax, w=w, a=a, b=b));
  

  #layout( cbind( matrix(1,4,4), matrix(2,4,4), matrix(3,4,4) , matrix(4,4,4) ) )
  
    y_max <- max(theta_post_pdf, theta_prior_pdf)
  
  par(mfrow=c(1,1))
  
  plot(x_theta, 
       theta_post_pdf, 
       type = "l",  
       main='Posterior vs Prior PDF', 
       xlab = expression(theta),
       ylab = expression(paste(pi(theta,"|",x[1:n]), " and " ,pi(theta))), 
       col="blue" ,
      xlim=c(0,1),
      ylim=c(0,y_max+1),
      cex.lab=fntsz, 
      cex.axis=fntsz, 
      cex.main=fntsz, 
      cex.sub=fntsz) ;
  
  lines(x_theta, 
        theta_prior_pdf,
        col= "green",
        cex.lab=fntsz, 
        cex.axis=fntsz, 
        cex.main=fntsz, 
        cex.sub=fntsz) ;
  
  abline(v=theta, 
         col="purple")
    
    legend('topright',
        c('Posterior','Prior',expression(paste(theta, " real value"))),
        bg="transparent",
        lty=c(1,1),
        lwd=c(2.5,2.5),
        col=c('blue','green','purple'),
        cex=fntsz
    )
}


plot_Bernoulli_pred <- function(theta=0.5, n=10, 
                                 Kmax=2, 
                                 w1=0.5,  
                                 a1=10, a2=15, 
                                 b1=20, b2=15) {
  
# theta=0.5
# n=10
# Kmax=2
# w1=0.5
# w2=0.2
# a1=10
# a2=20
# b1=20
# b2=10
  
  w2 = 1-w1 ;
  
  w = c(w1,w2)
  a = c(a1,a2)
  b = c(b1,b2)
  
  fntsz <- 1.5;
  op <- par(cex = fntsz)
  
  obs <-get_randnumb(n) ;
  obs <- 1*(obs<theta) ;
  
  r <- BernoulliModel_SufStat(obs,n) ;

  x_theta <- seq(from = 0, to = 1, length.out = 50) ;
  
  x_pred <- c(0,1) ;
  xnew_pred_pdf <- c(BernoulliModel_PDF_mixpred(0, r=r,  n=n, Kmax=Kmax, w=w, a=a, b=b),
                     BernoulliModel_PDF_mixpred(1, r=r,  n=n, Kmax=Kmax, w=w, a=a, b=b));
  

  #layout( cbind( matrix(1,4,4), matrix(2,4,4), matrix(3,4,4) , matrix(4,4,4) ) )
  
  par(mfrow=c(1,1))
  
  barplot(xnew_pred_pdf,  
          names = x_pred, 
          xlab = "x", 
          main = "Predictive PMF", 
          ylab = expression(paste("p(",x[n+1], "|", x["1:n"] , ")")), 
          ylim=c(0,1.1),
          cex.lab=fntsz, 
          cex.axis=fntsz, 
          cex.main=fntsz, 
          cex.sub=fntsz) ;
}  
  
  


```


Inputs {.sidebar} 
-----------------------

***Parametric model parameters***

```{r,echo=FALSE}
sliderInput("Bern_theta",
                 "$$\\theta:$$",
                  min = 0.0001,
                  max = 0.9999,
                  step = 0.05,
                  value = 0.5)

sliderInput("Bern_n",
                 "$$n:$$",
                  min = 1,
                  max = 200,
                  step = 1,
                  value = 10)
```

***Prior hyper-parameters***

```{r,echo=FALSE}
sliderInput("Bern_w1",
                 "$$\\varpi_1:$$",
                  min = 0.0,
                  max = 1.0,
                  step = 0.01,
                  value = 0.25)
sliderInput("Bern_a1",
                 "$$a_1:$$",
                  min = 0.05,
                  max = 35.0,
                  step = 0.05,
                  value = 18)
sliderInput("Bern_b1",
                 "$$b_1:$$",
                  min = 0.05,
                  max = 35.0,
                  step = 0.05,
                  value = 11)
sliderInput("Bern_a2",
                 "$$a_2:$$",
                  min = 0.05,
                  max = 35.0,
                  step = 0.05,
                  value = 10)
sliderInput("Bern_b2",
                 "$$b_2:$$",
                  min = 0.05,
                  max = 35.0,
                  step = 0.05,
                  value = 18)
```


Column {.tabset}
-----------------------


### ***Description*** 

We consider the Bayesian model

\[
\begin{cases}
x_{i}|\theta & \overset{\text{iid}}{\sim}\text{Br}(\theta),\,\,\forall i=1:n\\
\theta & \sim \pi(\text{d}\theta)
\end{cases}
\]

- The Mixture prior with components conjugate priors with $m=2$ components is a Beta (finite) mixture  
\[
\pi(\text{d}\theta)	=\sum_{l=1}^{m}\varpi_{l}\text{Be}(\text{d}\theta|a_{l},b_{l})
\]

- The posterior distribution for $\theta|x_{1:n}$ is a Beta (finite) mixture model:
such that
\[
\theta|x_{1:n}\sim\pi(\text{d}\theta|x_{1:n})
\]
such that
\[
\pi(\text{d}\theta|x_{1:n})	=\sum_{l=1}^{m}\varpi_{l}^{*}\text{Be}(\text{d}\theta|a^{*}_{l},b^{*}_{l})
\]
and 
\begin{align*}
 a_{l}^{*} 
 &=
 \sum_{i=1}^{n}x_{i}+a_{l} \\
 b_{l}^{*}
 &=
 n-\sum_{i=1}^{n}x_{i}+b_{l} \\
 \varpi_{l}^{*}
 &=
 \frac{\varpi_{l}\frac{\Gamma(a_{l}+b_{l})}{\Gamma(a_{l})\Gamma(b_{l})}\frac{\Gamma(a_{l}^{*})\Gamma(b_{l}^{*})}{\Gamma(a_{l}^{*}+b_{l}^{*})}}{\sum_{l=1}^{m}\varpi_{l}\frac{\Gamma(a_{l}+b_{l})}{\Gamma(a_{l})\Gamma(b_{l})}\frac{\Gamma(a_{l}^{*})\Gamma(b_{l}^{*})}{\Gamma(a_{l}^{*}+b_{l}^{*})}}
\end{align*}

- The predictive distribution for a future observation $x_{\text{new}}|x_{1:n}$ is a fixture of Beta-binomial distributions with $m=2$ components
\begin{align*}
x_{\text{new}}|x_{1:n} 
 &\sim
 p(\text{d}x_{\text{new}}|x_{1:n} )
\end{align*}
Where
\[
 p(\text{d}x_{\text{new}}|x_{1:n} )
=
\sum_{l=1}^{m}\varpi_{l}^{*}\text{Bb}(a_{l}^{*},b_{l}^{*},1) 
\]


### ***Parametric inferece***

```{r,echo=FALSE}
renderPlot({
   plot_Bernoulli_param(theta=input$Bern_theta, 
                  n=input$Bern_n, 
                  w1=input$Bern_w1, 
                  a1=input$Bern_a1, 
                  a2=input$Bern_a2, 
                  b1=input$Bern_b1, 
                  b2=input$Bern_b2
                  )
  })
```

### ***Predictive inference***

```{r,echo=FALSE}
renderPlot({
   plot_Bernoulli_pred(theta=input$Bern_theta, 
                  n=input$Bern_n, 
                  w1=input$Bern_w1, 
                  a1=input$Bern_a1, 
                  a2=input$Bern_a2, 
                  b1=input$Bern_b1, 
                  b2=input$Bern_b2
                  )
  })
```


# Bernoulli model (parametrisation 2)


```{r,echo=FALSE}

plot_Bernoulli2_param <- function(theta=0.5, n=10, 
                                 Kmax=2, 
                                 w1=0.5, 
                                 tau_10=18, tau_11=9, 
                                 tau_20=28, tau_21=14) {
  
# theta=0.5
# n=10
# Kmax=2
# w1=0.5
# w2=0.2
# a1=10
# a2=20
# b1=20
# b2=10
  
  
a1 = tau_11+1
b1 = tau_10-tau_11+1
a2 = tau_21+1
b2 = tau_20-tau_21+1
  
  
  w2 = 1-w1 ;
  
  w = c(w1,w2)
  a = c(a1,a2)
  b = c(b1,b2)
  
  fntsz <- 1.5;
  op <- par(cex = fntsz)
  
  obs <-get_randnumb(n) ;
  obs <- 1*(obs<theta) ;
  
  r <- BernoulliModel_SufStat(obs,n) ;

  x_theta <- seq(from = 0, to = 1, length.out = 50) ;
  theta_prior_pdf <- BernoulliModel_PDF_mixprior(x_theta, Kmax=Kmax, w=w, a=a, b=b) ;
  theta_post_pdf <- BernoulliModel_PDF_mixposterior(x_theta, r=r,  n=n, Kmax=Kmax, w=w, a=a, b=b);
  
  x_pred <- c(0,1) ;
  xnew_pred_pdf <- c(BernoulliModel_PDF_mixpred(0, r=r,  n=n, Kmax=Kmax, w=w, a=a, b=b),
                     BernoulliModel_PDF_mixpred(1, r=r,  n=n, Kmax=Kmax, w=w, a=a, b=b));
  

  #layout( cbind( matrix(1,4,4), matrix(2,4,4), matrix(3,4,4) , matrix(4,4,4) ) )
  
    y_max <- max(theta_post_pdf, theta_prior_pdf)
  
  par(mfrow=c(1,1))
  
  plot(x_theta, 
       theta_post_pdf, 
       type = "l",  
       main='Posterior vs Prior PDF', 
       xlab = expression(theta),
       ylab = expression(paste(pi(theta,"|",x[1:n]), " and " ,pi(theta))), 
       col="blue" ,
      xlim=c(0,1),
      ylim=c(0,y_max+1),
      cex.lab=fntsz, 
      cex.axis=fntsz, 
      cex.main=fntsz, 
      cex.sub=fntsz) ;
  
  lines(x_theta, 
        theta_prior_pdf,
        col= "green",
        cex.lab=fntsz, 
        cex.axis=fntsz, 
        cex.main=fntsz, 
        cex.sub=fntsz) ;
  
  abline(v=theta, 
         col="purple")
    
    legend('topright',
        c('Posterior','Prior',expression(paste(theta, " real value"))),
        bg="transparent",
        lty=c(1,1),
        lwd=c(2.5,2.5),
        col=c('blue','green','purple'),
        cex=fntsz
    )
}


plot_Bernoulli2_pred <- function(theta=0.5, n=10, 
                                 Kmax=2, 
                                 w1=0.5, 
                                 tau_10=18, tau_11=9, 
                                 tau_20=28, tau_21=14) {
  
# theta=0.5
# n=10
# Kmax=2
# w1=0.5
# w2=0.2
# a1=10
# a2=20
# b1=20
# b2=10
  
  
a1 = tau_11+1
b1 = tau_10-tau_11+1
a2 = tau_21+1
b2 = tau_20-tau_21+1
  
  w2 = 1-w1 ;
  
  w = c(w1,w2)
  a = c(a1,a2)
  b = c(b1,b2)
  
  fntsz <- 1.5;
  op <- par(cex = fntsz)
  
  obs <-get_randnumb(n) ;
  obs <- 1*(obs<theta) ;
  
  r <- BernoulliModel_SufStat(obs,n) ;

  x_theta <- seq(from = 0, to = 1, length.out = 50) ;
  
  x_pred <- c(0,1) ;
  xnew_pred_pdf <- c(BernoulliModel_PDF_mixpred(0, r=r,  n=n, Kmax=Kmax, w=w, a=a, b=b),
                     BernoulliModel_PDF_mixpred(1, r=r,  n=n, Kmax=Kmax, w=w, a=a, b=b));
  

  #layout( cbind( matrix(1,4,4), matrix(2,4,4), matrix(3,4,4) , matrix(4,4,4) ) )
  
  par(mfrow=c(1,1))
  
  barplot(xnew_pred_pdf,  
          names = x_pred, 
          xlab = "x", 
          main = "Predictive PMF", 
          ylab = expression(paste("p(",x[n+1], "|", x["1:n"] , ")")), 
          ylim=c(0,1.1),
          cex.lab=fntsz, 
          cex.axis=fntsz, 
          cex.main=fntsz, 
          cex.sub=fntsz) ;
}  
  
  
```


Inputs {.sidebar} 
-----------------------

***Parametric model parameters***

```{r,echo=FALSE}
sliderInput("Bern2_theta",
                 "$$\\theta:$$",
                  min = 0.0001,
                  max = 0.9999,
                  step = 0.05,
                  value = 0.5)

sliderInput("Bern2_n",
                 "$$n:$$",
                  min = 1,
                  max = 200,
                  step = 1,
                  value = 10)
```

***Prior hyper-parameters***

```{r,echo=FALSE}
sliderInput("Bern2_w1",
                 "$$\\varpi_1:$$",
                  min = 0.0,
                  max = 1.0,
                  step = 0.01,
                  value = 0.25)
sliderInput("Bern2_tau1",
                 "$$\\tau_{1}=(\\tau_{1,1},\\tau_{1,0}):$$",
                  min = 0.00,
                  max = 70,
                  step = 0.05,
                  value = c(17,27))
sliderInput("Bern2_tau2",
                 "$$\\tau_{2}=(\\tau_{2,1},\\tau_{2,0}):$$",
                  min = 0.00,
                  max = 70,
                  step = 0.05,
                  value = c(9,26))
```


Column {.tabset}
-----------------------


### ***Description*** 

We consider the Bayesian model

\[
\begin{cases}
x_{i}|\theta & \overset{\text{iid}}{\sim}\text{Br}(\theta),\,\,\forall i=1:n\\
\theta & \sim \pi(\text{d}\theta)
\end{cases}
\]

- The Mixture prior with components conjugate priors with $m=2$ components is a Beta (finite) mixture  
\[
\pi(\text{d}\theta)	=\sum_{l=1}^{m}\varpi_{l}\text{Be}(\text{d}\theta|\tau_{l,1}+1,\tau_{l,0}-\tau_{l,1}+1)
\]
where 
    * $\tau_{l,0}$ is the prior hyperperameter  reflecting the effective number of observations(as replacing the sufficient statistic part $n$ of the likelihood), and   
    * $\tau_{l,1}$ is the  prior hyperparameter reflecting the total amount that these (prior) pseudo-observations contribute to the sufficint statistice (as replacing the sufficient statistic part $\sum_{i=1}^{n}x_{i}$ of the likelihood).

- The posterior distribution for $\theta|x_{1:n}$ is a Beta (finite) mixture model:
such that
\[
\theta|x_{1:n}\sim\pi(\text{d}\theta|x_{1:n})
\]
such that
\[
\pi(\text{d}\theta|x_{1:n})	=\sum_{l=1}^{m}\varpi_{l}^{*}\text{Be}(\text{d}\theta|\tau^{*}_{l,1}+1,\tau^{*}_{l,0}-\tau^{*}_{l,1}+1)
\]
and 
\begin{align*}
 \tau^{*}_{l,0} 
 &=
 n+\tau_{l,0} \\
 \tau^{*}_{l,1}
 &=
 \sum_{i=1}^{n}x_{i}+\tau_{l,1} \\
 \varpi_{l}^{*}
 &=
 \frac
 {
 \varpi_{l}
 \frac
 {\Gamma(\tau_{l,0}+2)}
 {\Gamma(\tau_{l,1}+1)\Gamma(\tau_{l,0}-\tau^{*}_{l,1}+1)}
 \frac{
 \Gamma(\tau^{*}_{l,1}+1)\Gamma(\tau^{*}_{l,0}-\tau^{*}_{l,1}+1)
 }
 {
 \Gamma(\tau^{*}_{l,0})}
 }
 {\sum_{l=1}^{m} \varpi_{l}
 \frac
 {\Gamma(\tau_{l,0}+2)}
 {\Gamma(\tau_{l,1}+1)\Gamma(\tau_{l,0}-\tau^{*}_{l,1}+1)}
 \frac{
 \Gamma(\tau^{*}_{l,1}+1)\Gamma(\tau^{*}_{l,0}-\tau^{*}_{l,1}+1)
 }
 {
 \Gamma(\tau^{*}_{l,0})}}
\end{align*}

- The predictive distribution for a future observation $x_{\text{new}}|x_{1:n}$ is a fixture of Beta-binomial distributions with $m=2$ components
\begin{align*}
x_{\text{new}}|x_{1:n} 
 &\sim
 p(\text{d}x_{\text{new}}|x_{1:n} )
\end{align*}
Where
\[
 p(\text{d}x_{\text{new}}|x_{1:n} )
=
\sum_{l=1}^{m}\varpi_{l}^{*}\text{Bb}(\tau^{*}_{l,1}+1,\tau^{*}_{l,0}-\tau^{*}_{l,1}+1,1) 
\]


### ***Parametric inferece***

```{r,echo=FALSE}
renderPlot({
   plot_Bernoulli2_param(theta=input$Bern2_theta, 
                  n=input$Bern2_n, 
                  w1=input$Bern2_w1, 
                  tau_10=input$Bern2_tau1[2], 
                  tau_11=input$Bern2_tau1[1], 
                  tau_20=input$Bern2_tau2[2], 
                  tau_21=input$Bern2_tau2[1]
                  )
  })
```

### ***Predictive inference***

```{r,echo=FALSE}
renderPlot({
   plot_Bernoulli2_pred(theta=input$Bern2_theta, 
                  n=input$Bern2_n, 
                  w1=input$Bern2_w1, 
                  tau_10=input$Bern2_tau1[2], 
                  tau_11=input$Bern2_tau1[1], 
                  tau_20=input$Bern2_tau2[2], 
                  tau_21=input$Bern2_tau2[1]
                  )
  })
```






