<!-- -------------------------------------------------------------------------------- -->

<!-- Copyright 2019 Georgios Karagiannis -->

<!-- georgios.karagiannis@durham.ac.uk -->
<!-- Assistant Professor -->
<!-- Department of Mathematical Sciences, Durham University, Durham,  UK  -->

<!-- This file is part of Bayesian_Statistics (MATH3341/4031 Bayesian Statistics III/IV) -->
<!-- which is the material of the course (MATH3341/4031 Bayesian Statistics III/IV) -->
<!-- taught by Georgios P. Katagiannis in the Department of Mathematical Sciences   -->
<!-- in the University of Durham  in Michaelmas term in 2019 -->

<!-- Bayesian_Statistics is free software: you can redistribute it and/or modify -->
<!-- it under the terms of the GNU General Public License as published by -->
<!-- the Free Software Foundation version 3 of the License. -->

<!-- Bayesian_Statistics is distributed in the hope that it will be useful, -->
<!-- but WITHOUT ANY WARRANTY; without even the implied warranty of -->
<!-- MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the -->
<!-- GNU General Public License for more details. -->

<!-- You should have received a copy of the GNU General Public License -->
<!-- along with Bayesian_Statistics  If not, see <http://www.gnu.org/licenses/>. -->

<!-- -------------------------------------------------------------------------------- -->

---
title: "Weak Law of Large Numbers"
output: flexdashboard::flex_dashboard
runtime: shiny
---

<!-- Weak law of large numbers Applet -->
<!-- Copyright (C) 2017  Georgios Karagiannis -->
<!-- georgios.karagiannis@durham.ac.uk -->

<!-- This program is free software: you can redistribute it and/or modify -->
<!-- it under the terms of the GNU General Public License as published by -->
<!-- the Free Software Foundation, either version 3 of the License, or -->
<!-- (at your option) any later version. -->

<!-- This program is distributed in the hope that it will be useful, -->
<!-- but WITHOUT ANY WARRANTY; without even the implied warranty of -->
<!-- MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the -->
<!-- GNU General Public License for more details. -->

<!-- You should have received a copy of the GNU General Public License -->
<!-- along with this program. If not, see <http://www.gnu.org/licenses/>. -->

# Normal distribution  
```{r,echo=FALSE}
plot_wlln_normal <- function(mu= 0.0, sig2=1.0, N=1000, fresh=0) {
  
  K = 100 ;
  
  n <- seq(1,N) ;
  
  # write( rnorm(10^5,0,1) , file = "./z_rnorm.dat")
  
  if (fresh==1) {
    z <- matrix(rnorm(N*K, mean = 0, sd = 1),N,K) ;
    # write(z,file = "./z_rnorm.dat") ;
  }
  else {
    z <- scan(file = "./z_rnorm.dat",nmax=N*K) ;
    z <- matrix(z ,N,K) ; 
  }
  
  x <- mu+sqrt(sig2)*z[1:N,] ;
  x_bar <- matrix(rep(NaN,N*K),N,K) ;
  
  for (j in 1:K) x_bar[,j] <- cumsum(x[,j])/n ;
  CI_x_bar <- t(t(apply(x_bar, 1, sort)) [ ,c( ceiling(2.5/100*K), ceiling(97.5/100*K))] )
  
  x_min_exp <- mu -3.5*sqrt(sig2/N) ;
  x_max_exp <- mu +3.5*sqrt(sig2/N) ;
  
  x_min <-  0 -2.5*sqrt(1000) ;
  x_max <- 0 +2.5*sqrt(1000) ;
  
  n_min = 1 ;
  n_max = 400
  
  x_eval <- seq(x_min,x_max,length=100) ;
  Pr_x <- dnorm(x_eval,mu,sqrt(sig2)) ;
  
  layout( matrix( c(1,1,2,2,2,2,2,2),nrow=2,ncol=4) )
  
  plot(Pr_x, x_eval, 
       ylim=c(x_min,x_max), 
       xlim=rev(range(Pr_x)), 
       type='l',
       ylab='X',
       xlab='PDF f(X)')
  abline(a=mu, b=0, col='red')
  
  plot(n, x_bar[,1], 
       type='l',
       main="Trace plot of the sample mean",
       xlab="Sample size",
       ylab= 'Sample average of X',
       xlim=c(n_min,n_max), 
       ylim=c(x_min,x_max),
       col='lightgrey'
  )
  #for (j in 2:min(10,K))
  for (j in 2:K)
    lines(n, x_bar[,j], 
         type='l',
         col='lightgrey')
  lines(n, CI_x_bar[1,], 
        type='l',
        col='blue', 
        lwd=2
        )
  lines(n, CI_x_bar[2,], 
        type='l',
        col='blue', 
        lwd=2
  )
  abline(a=mu, b=0, col='red', 
         lwd=2
         )
  
  legend('topright',
         bg="transparent",
         c('Sample mean','95% CI of sample mean','Population mean'),
         lty=c(1,1),
         lwd=c(2.5,2.5),
         col=c('lightgrey','blue','red')) 
  
}
```

Column {.sidebar}
-----------------------------------------------------------------------

***Sample***

$$X_i\sim\text{N}(\mu,\sigma^2)$$
$$\text{E}(X_i) = \mu$$
$$\text{Var}(X_i) = \sigma^2$$
$$ i=1,...,N$$

***Parameters***
```{r,echo=FALSE}
      sliderInput("mu_norm",
                  "$$\\mu$$",
                  min = -10.0,
                  max = 10.0,
                  value = 0.0,
                  step = ((3)-(-3))/100 )
      sliderInput("sig2_norm",
                  "$$\\text{Var}(X)=\\sigma^2$$",
                  min = 10^(-10),
                  max = 1000,
                  value = 100.0,
                  step = 1)
      sliderInput("N_norm",
                  "N :",
                  min = 2,
                  max = 400,
                  value = 200)  
      selectInput("fresh_norm", "Fresh:",
                  c("Yes" = 1,
                    "No" = 0))
```

Column
-----------------------------------------------------------------------

### ***Theorems*** {data-height=400}

***Chebyshev's inequality:*** 
Let $X$ be a r.v. with mean $\text{E}(X)=\mu$ and variance
$\text{Var}(X)=\sigma^{2}<\infty$. Then, 
$$
\Pr(|X-\mu|\ge \epsilon)\le\frac{\sigma^{2}}{\epsilon^{2}}, \text{ for any } \epsilon>0.
$$

***Chebyshev's inequality (For the average):*** 
Let $X_{1},...,X_{N}$ be i.i.d. r.v., with
 mean $\text{E}(X_{i})=\mu$ and variance $\text{Var}(X_{i})=\sigma^{2}<\infty$.
Let $\bar{X}=\frac{1}{N}\sum_{i=1}^{N}X_{i}$. Then, 
$$
\Pr(|\bar{X}-\mu|\ge \epsilon)\le\frac{\sigma^{2}/N}{\epsilon^{2}}, \text{ for any } \epsilon>0.
$$

***Weak Law of Large Numbers (WLLN):*** 
Let $X_{1},...,X_{N}$ be a sequence of i.i.d. r.v., with
 mean $\text{E}(X_{i})=\mu$ and variance $\text{Var}(X_{i})=\sigma^{2}<\infty$.
Let $\bar{X}=\frac{1}{N}\sum_{i=1}^{N}X_{i}$. Then as $N\rightarrow\infty$,  
$$
\lim_{n\rightarrow\infty}\Pr(|\bar{X}-\mu|\ge\epsilon)=0, \text{ for any } \epsilon>0.
$$

### ***Simulation plots*** {data-height=600}

```{r,echo=FALSE}
renderPlot({
    plot_wlln_normal(mu=input$mu_norm,
                     sig2=input$sig2_norm,
                     N=input$N_norm,
                     fresh=input$fresh_norm)
  })
```









# Bernoulli distribution 
```{r,echo=FALSE}
plot_wlln_Bernoulli <- function(sig2=0.1, N=1000, fresh=0) {
  
  pr = 0.2*(1-sqrt(1-4*sig2)) ;
  # pr = 0.2*(1-sqrt(1+4*sig2)) ;
  
  K = 100 ;
  
  n <- seq(1,N) ;
  
  # write( rnorm(10^5,0,1) , file = "./z_rnorm.dat")
  
  if (fresh==1) {
    z <- runif(N*K);
    # write(z,file = "./z_runif.dat") ;
  }
  else {
    z <- scan(file = "./z_runif.dat",nmax=N*K) ;
  }
  x <- matrix(as.integer(z<pr),N,K) ;
  
  x_bar <- matrix(rep(NaN,N*K),N,K) ;
  for (j in 1:K) x_bar[,j] <- cumsum(x[,j])/n ;
  CI_x_bar <- t(t(apply(x_bar, 1, sort)) [ ,c( ceiling(2.5/100*K), ceiling(97.5/100*K))] )
  
  x_min <-  0 ;
  x_max <- 1 ;

  n_min = 1 ;
  n_max = 400
  
  x_eval <- c(0.0,1.0) ;
  Pr_x <- c(1-pr,pr) ;
  
  layout( matrix( c(1,1,2,2,2,2,2,2),nrow=2,ncol=4) )
  
  barplot(Pr_x, names=c(0,1), 
          horiz=TRUE, 
          xlim = rev(c(0,1)),
          xlab="Pr(X)",
          ylab= 'X',
          main = 'Prob. of X'
          )
  
  plot(n, x_bar[,1], 
       type='l',
       main="Trace plot of the sample mean",
       xlab="Sample size",
       ylab= 'Sample average of X',
       xlim=c(n_min,n_max), 
       ylim=c(x_min,x_max),
       col='lightgrey'
  )
  #for (j in 2:min(10,K))
  for (j in 2:K)
    lines(n, x_bar[,j], 
          type='l',
          col='lightgrey')
  lines(n, CI_x_bar[1,], 
        type='l',
        col='blue', 
        lwd=2
  )
  lines(n, CI_x_bar[2,], 
        type='l',
        col='blue', 
        lwd=2
  )
  abline(a=pr, b=0, col='red', 
         lwd=2
  )
  
  legend('topright',
         bg="transparent",
         c('Sample mean','95% CI of sample mean','Population mean'),
         lty=c(1,1),
         lwd=c(2.5,2.5),
         col=c('lightgrey','blue','red')) 
  
}
```

Column {.sidebar}
-----------------------------------------------------------------------

***Sample***

$$X_i\sim\text{Bernoulli}(p)$$
$$\text{E}(X_i) = p$$
$$\text{Var}(X_i) = p(1-p)$$
$$ i=1,...,N$$

***Parameters***
```{r,echo=FALSE}
      sliderInput("sig2_Bern",
                  "$$\\text{Var}(X): $$",
                  min = 0,
                  max = 0.25,
                  value = 0.1,
                  step = 0.001)
      sliderInput("N_Bern",
                  "$$N :$$",
                  min = 2,
                  max = 400,
                  value = 200) 
      selectInput("fresh_Bern", "Fresh:",
                  c("Yes" = 1,
                    "No" = 0))
```

Column
-----------------------------------------------------------------------

### ***Theorems*** {data-height=400}

***Chebyshev's inequality:*** 
Let $X$ be a r.v. with mean $\text{E}(X)=\mu$ and variance
$\text{Var}(X)=\sigma^{2}<\infty$. Then, 
$$
\Pr(|X-\mu|\ge \epsilon)\le\frac{\sigma^{2}}{\epsilon^{2}}, \text{ for any } \epsilon>0.
$$

***Chebyshev's inequality (For the average):*** 
Let $X_{1},...,X_{N}$ be i.i.d. r.v., with
 mean $\text{E}(X_{i})=\mu$ and variance $\text{Var}(X_{i})=\sigma^{2}<\infty$.
Let $\bar{X}=\frac{1}{N}\sum_{i=1}^{N}X_{i}$. Then, 
$$
\Pr(|\bar{X}-\mu|\ge \epsilon)\le\frac{\sigma^{2}/N}{\epsilon^{2}}, \text{ for any } \epsilon>0.
$$

***Weak Law of Large Numbers (WLLN):*** 
Let $X_{1},...,X_{N}$ be a sequence of i.i.d. r.v., with
 mean $\text{E}(X_{i})=\mu$ and variance $\text{Var}(X_{i})=\sigma^{2}<\infty$.
Let $\bar{X}=\frac{1}{N}\sum_{i=1}^{N}X_{i}$. Then as $N\rightarrow\infty$,  
$$
\lim_{n\rightarrow\infty}\Pr(|\bar{X}-\mu|\ge\epsilon)=0, \text{ for any } \epsilon>0.
$$

### ***Simulation plots*** {data-height=600}

```{r,echo=FALSE}
renderPlot({
    plot_wlln_Bernoulli(sig2=input$sig2_Bern,
                     N=input$N_Bern,
                     fresh=input$fresh_Bern)
  })
```





# Exponential distribution  
```{r,echo=FALSE}
plot_wlln_exponential <- function(  sig2=1.0, N=1000, fresh=0) {
  
  ell <- 1.0/sqrt(sig2)
  mu <- 1.0/ell ; 
  
  K = 100 ;
  
  n <- seq(1,N) ;
  
  # write( rnorm(10^5,0,1) , file = "./z_rnorm.dat")
  
  if (fresh==1) {
    z <- runif(N*K);
    # write(z,file = "./z_runif.dat") ;
  }
  else {
    z <- scan(file = "./z_runif.dat",nmax=N*K) ;
  }
  z = -log(1-z) ;
  
  x <- matrix(1/ell*z,N,K) ;
  
  x_bar <- matrix(rep(NaN,N*K),N,K) ;
  
  for (j in 1:K) x_bar[,j] <- cumsum(x[,j])/n ;
  CI_x_bar <- t(t(apply(x_bar, 1, sort)) [ ,c( ceiling(2.5/100*K), ceiling(97.5/100*K))] )
  
  x_min_exp <- 0.0 ;
  x_max_exp <- 6.0 ;
  
  x_min <-  0.0 ;
  x_max <- 6.0 ;
  
  n_min = 1 ;
  n_max = 400
  
  x_eval <- seq(x_min,x_max,length=100) ;
  Pr_x <- dexp(x_eval,rate=ell) ;
  
  layout( matrix( c(1,1,2,2,2,2,2,2),nrow=2,ncol=4) )
  
  plot(Pr_x, x_eval, 
       ylim=c(x_min,x_max), 
       xlim=rev(range(Pr_x)), 
       type='l',
       ylab='X',
       xlab='PDF f(X)')
  abline(a=mu, b=0, col='red')
  
  plot(n, x_bar[,1], 
       type='l',
       main="Trace plot of the sample mean",
       xlab="Sample size",
       ylab= 'Sample average of X',
       xlim=c(n_min,n_max), 
       ylim=c(x_min,x_max),
       col='lightgrey'
  )
  #for (j in 2:min(10,K))
  for (j in 2:K)
    lines(n, x_bar[,j], 
          type='l',
          col='lightgrey')
  lines(n, CI_x_bar[1,], 
        type='l',
        col='blue', 
        lwd=2
  )
  lines(n, CI_x_bar[2,], 
        type='l',
        col='blue', 
        lwd=2
  )
  abline(a=mu, b=0, col='red', 
         lwd=2
  )
  
  legend('topright',
         bg="transparent",
         c('Sample mean','95% CI of sample mean','Population mean'),
         lty=c(1,1),
         lwd=c(2.5,2.5),
         col=c('lightgrey','blue','red')) 
  
}
```

Column {.sidebar}
-----------------------------------------------------------------------

***Sample***

$$X_i\sim\mathcal{E}(\lambda) $$
$$\text{E}(X_i) = 1/\lambda$$
$$\text{Var}(X_i) = 1/\lambda^2$$
$$ i=1,...,N$$

***Parameters***
```{r,echo=FALSE}
      sliderInput("sig2_exp",
                  "$$\\text{Var}(X):$$",
                  min = 10^(-5),
                  max = 16.0,
                  value = 1.0,
                  step = 0.1)
      sliderInput("N_exp",
                  "N :",
                  min = 2,
                  max = 400,
                  value = 200)
      selectInput("fresh_exp", "Fresh:",
                  c("Yes" = 1,
                    "No" = 0))
```

Column
-----------------------------------------------------------------------

### ***Theorems*** {data-height=400}

***Chebyshev's inequality:*** 
Let $X$ be a r.v. with mean $\text{E}(X)=\mu$ and variance
$\text{Var}(X)=\sigma^{2}<\infty$. Then, 
$$
\Pr(|X-\mu|\ge \epsilon)\le\frac{\sigma^{2}}{\epsilon^{2}}, \text{ for any } \epsilon>0.
$$

***Chebyshev's inequality (For the average):*** 
Let $X_{1},...,X_{N}$ be i.i.d. r.v., with
 mean $\text{E}(X_{i})=\mu$ and variance $\text{Var}(X_{i})=\sigma^{2}<\infty$.
Let $\bar{X}=\frac{1}{N}\sum_{i=1}^{N}X_{i}$. Then, 
$$
\Pr(|\bar{X}-\mu|\ge \epsilon)\le\frac{\sigma^{2}/N}{\epsilon^{2}}, \text{ for any } \epsilon>0.
$$

***Weak Law of Large Numbers (WLLN):*** 
Let $X_{1},...,X_{N}$ be a sequence of i.i.d. r.v., with
 mean $\text{E}(X_{i})=\mu$ and variance $\text{Var}(X_{i})=\sigma^{2}<\infty$.
Let $\bar{X}=\frac{1}{N}\sum_{i=1}^{N}X_{i}$. Then as $N\rightarrow\infty$,  
$$
\lim_{n\rightarrow\infty}\Pr(|\bar{X}-\mu|\ge\epsilon)=0, \text{ for any } \epsilon>0.
$$

### ***Simulation plots*** {data-height=600}

```{r,echo=FALSE}
renderPlot({
    plot_wlln_exponential(sig2=input$sig2_exp,
                     N=input$N_exp,
                     fresh=input$fresh_exp)
  })
```











