<!-- -------------------------------------------------------------------------------- -->

<!-- Copyright 2019 Georgios Karagiannis -->

<!-- georgios.karagiannis@durham.ac.uk -->
<!-- Assistant Professor -->
<!-- Department of Mathematical Sciences, Durham University, Durham,  UK  -->

<!-- This file is part of Bayesian_Statistics (MATH3341/4031 Bayesian Statistics III/IV) -->
<!-- which is the material of the course (MATH3341/4031 Bayesian Statistics III/IV) -->
<!-- taught by Georgios P. Katagiannis in the Department of Mathematical Sciences   -->
<!-- in the University of Durham  in Michaelmas term in 2019 -->

<!-- Bayesian_Statistics is free software: you can redistribute it and/or modify -->
<!-- it under the terms of the GNU General Public License as published by -->
<!-- the Free Software Foundation version 3 of the License. -->

<!-- Bayesian_Statistics is distributed in the hope that it will be useful, -->
<!-- but WITHOUT ANY WARRANTY; without even the implied warranty of -->
<!-- MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the -->
<!-- GNU General Public License for more details. -->

<!-- You should have received a copy of the GNU General Public License -->
<!-- along with Bayesian_Statistics  If not, see <http://www.gnu.org/licenses/>. -->

<!-- -------------------------------------------------------------------------------- -->

---
title: "Conjugate priors/posteriors"
output: flexdashboard::flex_dashboard
runtime: shiny
---

<!-- Conjugate prior Applet -->
<!-- Copyright (C) 2017  Georgios Karagiannis -->
<!-- georgios.karagiannis@durham.ac.uk -->

<!-- This program is free software: you can redistribute it and/or modify -->
<!-- it under the terms of the GNU General Public License as published by -->
<!-- the Free Software Foundation, either version 3 of the License, or -->
<!-- (at your option) any later version. -->

<!-- This program is distributed in the hope that it will be useful, -->
<!-- but WITHOUT ANY WARRANTY; without even the implied warranty of -->
<!-- MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the -->
<!-- GNU General Public License for more details. -->

<!-- You should have received a copy of the GNU General Public License -->
<!-- along with this program. If not, see <http://www.gnu.org/licenses/>. -->

```{r}
get_randnumb <-function(n) {
  u <- seq(0.001,0.999,length.out = n)
#  u <- runif(0,1,n = n)
}
```

```{r}
#library(rmutil)
dbetabinom <- function (y, size, m, s, log = FALSE) {
  # from rmutil package
    if (any(y < 0)) 
        stop("y must contain non-negative values")
    if (any(size < 0)) 
        stop("size must contain non-negative values")
    if (any(m <= 0) || any(m >= 1)) 
        stop("m must lie between 0 and 1")
    if (any(s <= 0)) 
        stop("s must be positive")
    ly <- max(length(y), length(m), length(s), length(size))
    if (length(y) != ly) {
        if (length(y) == 1) 
            y <- rep(y, ly)
        else stop("length of y incorrect")
    }
    if (length(size) != ly) {
        if (length(size) == 1) 
            size <- rep(size, ly)
        else stop("size must be the same length as y")
    }
    if (any(y > size)) 
        stop("y must be <= size")
    if (length(m) != ly) {
        if (length(m) == 1) 
            m <- rep(m, ly)
        else stop("m and y must have the same length")
    }
    if (length(s) != ly) {
        if (length(s) == 1) 
            s <- rep(s, ly)
        else stop("s and y must have the same length")
    }
    t <- s * m
    u <- s * (1 - m)
    tmp <- lbeta(y + t, size - y + u) - lbeta(t, u) + lchoose(size, 
        y)
    if (!log) 
        tmp <- exp(tmp)
    tmp
}

BinomialBeta_pdf<-function(x,a,b,n) {
  pdf <- dbetabinom(y=x, size=n, m=a/(a+b), s=a+b) ;
  return (pdf) 
}
```


```{r}
StudentT_pdf<-function(x,mu,lam,al) {
  pdf <- 0.5*log(lam) -0.5*log(al) -0.5*log(pi) ;
  pdf <- pdf +lgamma(0.5*(al+1)) -lgamma(0.5*al) ;
  pdf <- pdf -0.5*(al+1)*log(1+lam/al*(x-mu)^2) ;
  pdf <- exp(pdf) ;
  return (pdf)  
}
```


```{r}
dinvgamma <- function(x, a, b){
  return ( exp( a*log(b)-lgamma(a) -(a+1)*log(x) -b/x ) ) ;
}
```



# Bernoulli model


```{r,echo=FALSE}
BernoulliModel_SufStat<-function(x,n) {
 r <- sum(x[1:n]) ; 
}

BernoulliModel_PDF_posterior<-function(theta,r,n,a,b) {
  pdf <- dbeta(theta, a+r, b+n-r) ;
   return (pdf)  
}

BernoulliModel_PDF_prior<-function(theta,a,b) {
  pdf <- dbeta(theta, a, b) ;
   return (pdf)  
}

BernoulliModel_PDF_pred<-function(y,r,n,a,b) {
  pdf <- BinomialBeta_pdf(y, a+r, b+n-r, 1) ;
   return (pdf)  
}

plot_Bernoulli_param <- function(theta=0.5, n=10, a=1, b=1) {
  
  # theta=0.5; n=10; a=1; b=1
  
  fntsz <- 1.5;
  op <- par(cex = fntsz)
  
  obs <-get_randnumb(n) ;
  obs <- 1*(obs<theta) ;
  
  r <- BernoulliModel_SufStat(obs,n) ;

  x_theta <- seq(from = 0, to = 1, length.out = 50) ;
  theta_prior_pdf <- BernoulliModel_PDF_prior(x_theta, a=a, b=b) ;
  theta_post_pdf <- BernoulliModel_PDF_posterior(x_theta, r=r,  n=n, a=a, b=b) ;
  
  x_pred <- c(0,1) ;
  xnew_pred_pdf <- c(BernoulliModel_PDF_pred(0, r=r,  n=n, a=a, b=b),
                     BernoulliModel_PDF_pred(1, r=r,  n=n, a=a, b=b));
  

  #layout( cbind( matrix(1,4,4), matrix(2,4,4), matrix(3,4,4) , matrix(4,4,4) ) )
  
    y_max <- max(theta_post_pdf, theta_prior_pdf)
  
  par(mfrow=c(1,1))
  
  plot(x_theta, 
       theta_post_pdf, 
       type = "l",  
       main='Posterior vs Prior PDF', 
       xlab = expression(theta),
       ylab = expression(paste(pi(theta,"|",x[1:n]), " and " ,pi(theta))), 
       col="blue" ,
      xlim=c(0,1),
      ylim=c(0,y_max+1),
      cex.lab=fntsz, 
      cex.axis=fntsz, 
      cex.main=fntsz, 
      cex.sub=fntsz) ;
  
  lines(x_theta, 
        theta_prior_pdf,
        col= "green",
        cex.lab=fntsz, 
        cex.axis=fntsz, 
        cex.main=fntsz, 
        cex.sub=fntsz) ;
  
  abline(v=theta, 
         col="purple")
    
    legend('topright',
        c('Posterior','Prior',expression(paste(theta, " real value"))),
        bg="transparent",
        lty=c(1,1),
        lwd=c(2.5,2.5),
        col=c('blue','green','purple'),
        cex=fntsz
    )
}


plot_Bernoulli_pred <- function(theta=0.5, n=10, a=1, b=1) {
  
  # theta=0.5; n=10; a=1; b=1
  
  fntsz <- 1.5;
  op <- par(cex = fntsz)
  
  obs <-get_randnumb(n) ;
  obs <- 1*(obs<theta) ;
  
  r <- BernoulliModel_SufStat(obs,n) ;

  x_theta <- seq(from = 0, to = 1, length.out = 50) ;
  
  x_pred <- c(0,1) ;
  xnew_pred_pdf <- c(BernoulliModel_PDF_pred(0, r=r,  n=n, a=a, b=b),
                     BernoulliModel_PDF_pred(1, r=r,  n=n, a=a, b=b));
  

  #layout( cbind( matrix(1,4,4), matrix(2,4,4), matrix(3,4,4) , matrix(4,4,4) ) )
  
  par(mfrow=c(1,1))
  
  barplot(xnew_pred_pdf,  
          names = x_pred, 
          xlab = "x", 
          main = "Predictive PMF", 
          ylab = expression(paste("p(",x[n+1], "|", x["1:n"] , ")")), 
          ylim=c(0,1.1),
          cex.lab=fntsz, 
          cex.axis=fntsz, 
          cex.main=fntsz, 
          cex.sub=fntsz) ;
}  
  
  


```


Inputs {.sidebar} 
-----------------------

***Parametric model parameters***

```{r,echo=FALSE}
sliderInput("Bern_theta",
                 "$$\\theta:$$",
                  min = 0.0001,
                  max = 0.9999,
                  step = 0.05,
                  value = 0.5)

sliderInput("Bern_n",
                 "$$n:$$",
                  min = 1,
                  max = 80,
                  step = 1,
                  value = 10)
```

***Prior hyper-parameters***

```{r,echo=FALSE}
sliderInput("Bern_a",
                 "$$a:$$",
                  min = 0.0,
                  max = 10.0,
                  step = 0.05,
                  value = 1)

sliderInput("Bern_b",
                 "$$b:$$",
                  min = 0.0,
                  max = 10.0,
                  step = 0.05,
                  value = 1)
```


Column {.tabset}
-----------------------


### ***Description*** 

We consider the Bayesian model

\[
\begin{cases}
x_{i}|\theta & \overset{\text{iid}}{\sim}\text{Br}(\theta),\,\,\forall i=1:n\\
\theta & \sim \text{Be}(a,b)
\end{cases}
\]

The posterior distribution for $\theta|x_{1:n}$ is: 
\begin{align*}
\theta|x_{1:n}
&\sim
\text{Be}(a^{*},b^{*})
\end{align*}

The predictive distribution for a future observation $x_{\text{new}}|x_{1:n}$ is

\begin{align*}
x_{\text{new}}|x_{1:n} 
 &\sim
 \text{Bb}(a^{*},b^{*},1) 
\end{align*}

Where

\begin{align*}
 a^{*} 
 &=
 \sum_{i=1}^{n}x_{i}+a \\
 b^{*}
 &=
 n-\sum_{i=1}^{n}x_{i}+b
\end{align*}


### ***Parametric inferece***

```{r,echo=FALSE}
renderPlot({
   plot_Bernoulli_param(theta=input$Bern_theta, 
                  n=input$Bern_n, 
                  a=input$Bern_a, 
                  b=input$Bern_b
                  )
  })
```

### ***Predictive inference***

```{r,echo=FALSE}
renderPlot({
   plot_Bernoulli_pred(theta=input$Bern_theta, 
                  n=input$Bern_n, 
                  a=input$Bern_a, 
                  b=input$Bern_b
                  )
  })
```







# Bernoulli model (parametrisation 2)


```{r,echo=FALSE}

plot_Bernoulli2_param <- function(theta=0.5, n=10,
                                 tau_0=0, tau_1=0) {
  

  a = tau_1+1
  b = tau_0-tau_1+1

  fntsz <- 1.5;
  op <- par(cex = fntsz)
  
  obs <-get_randnumb(n) ;
  obs <- 1*(obs<theta) ;
  
  r <- BernoulliModel_SufStat(obs,n) ;

  x_theta <- seq(from = 0, to = 1, length.out = 50) ;
  theta_prior_pdf <- BernoulliModel_PDF_prior(x_theta, a=a, b=b) ;
  theta_post_pdf <- BernoulliModel_PDF_posterior(x_theta, r=r,  n=n, a=a, b=b) ;
  
  x_pred <- c(0,1) ;
  xnew_pred_pdf <- c(BernoulliModel_PDF_pred(0, r=r,  n=n, a=a, b=b),
                     BernoulliModel_PDF_pred(1, r=r,  n=n, a=a, b=b));
  

  #layout( cbind( matrix(1,4,4), matrix(2,4,4), matrix(3,4,4) , matrix(4,4,4) ) )
  
    y_max <- max(theta_post_pdf, theta_prior_pdf)
  
  par(mfrow=c(1,1))
  
  plot(x_theta, 
       theta_post_pdf, 
       type = "l",  
       main='Posterior vs Prior PDF', 
       xlab = expression(theta),
       ylab = expression(paste(pi(theta,"|",x[1:n]), " and " ,pi(theta))), 
       col="blue" ,
      xlim=c(0,1),
      ylim=c(0,y_max+1),
      cex.lab=fntsz, 
      cex.axis=fntsz, 
      cex.main=fntsz, 
      cex.sub=fntsz) ;
  
  lines(x_theta, 
        theta_prior_pdf,
        col= "green",
        cex.lab=fntsz, 
        cex.axis=fntsz, 
        cex.main=fntsz, 
        cex.sub=fntsz) ;
  
  abline(v=theta, 
         col="purple")
    
    legend('topright',
        c('Posterior','Prior',expression(paste(theta, " real value"))),
        bg="transparent",
        lty=c(1,1),
        lwd=c(2.5,2.5),
        col=c('blue','green','purple'),
        cex=fntsz
    )

}


plot_Bernoulli2_pred <- function(theta=0.5, n=10, 
                                 tau_0=0, tau_1=0) {
  
# theta=0.5
# n=10
# a=10
# b=20
  
  
a = tau_1+1
b = tau_0-tau_1+1


  fntsz <- 1.5;
  op <- par(cex = fntsz)
  
  obs <-get_randnumb(n) ;
  obs <- 1*(obs<theta) ;
  
  r <- BernoulliModel_SufStat(obs,n) ;

  x_theta <- seq(from = 0, to = 1, length.out = 50) ;
  
  x_pred <- c(0,1) ;
  xnew_pred_pdf <- c(BernoulliModel_PDF_pred(0, r=r,  n=n, a=a, b=b),
                     BernoulliModel_PDF_pred(1, r=r,  n=n, a=a, b=b));
  

  #layout( cbind( matrix(1,4,4), matrix(2,4,4), matrix(3,4,4) , matrix(4,4,4) ) )
  
  par(mfrow=c(1,1))
  
  barplot(xnew_pred_pdf,  
          names = x_pred, 
          xlab = "x", 
          main = "Predictive PMF", 
          ylab = expression(paste("p(",x[n+1], "|", x["1:n"] , ")")), 
          ylim=c(0,1.1),
          cex.lab=fntsz, 
          cex.axis=fntsz, 
          cex.main=fntsz, 
          cex.sub=fntsz) ;



}  
  
  
```


Inputs {.sidebar} 
-----------------------

***Parametric model parameters***

```{r,echo=FALSE}
sliderInput("Bern2_theta",
                 "$$\\theta:$$",
                  min = 0.0001,
                  max = 0.9999,
                  step = 0.05,
                  value = 0.5)

sliderInput("Bern2_n",
                 "$$n:$$",
                  min = 1,
                  max = 200,
                  step = 1,
                  value = 10)
```

***Prior hyper-parameters***

```{r,echo=FALSE}
sliderInput("Bern2_tau",
                 "$$\\tau=(\\tau_{1},\\tau_{0}):$$",
                  min = 0.00,
                  max = 18,
                  step = 0.05,
                  value = c(0,0))
```


Column {.tabset}
-----------------------


### ***Description*** 

We consider the Bayesian model

\[
\begin{cases}
x_{i}|\theta & \overset{\text{iid}}{\sim}\text{Br}(\theta),\,\,\forall i=1:n\\
\theta & \sim \text{Be}(\tau_{1}+1,\tau_{0}-\tau_{1}+1)
\end{cases}
\]

Here:

* $\tau_{0}$ is the prior hyperperameter  reflecting the effective number of observations(as replacing the sufficient statistic part $n$ of the likelihood), and   
* $\tau_{1}$ is the  prior hyperparameter reflecting the total amount that these (prior) pseudo-observations contribute to the sufficint statistice (as replacing the sufficient statistic part $\sum_{i=1}^{n}x_{i}$ of the likelihood).


The posterior distribution for $\theta|x_{1:n}$ is: 
\begin{align*}
\theta|x_{1:n}
&\sim
\text{Be}(\tau^{*}_{1}+1,\tau^{*}_{0}-\tau^{*}_{1}+1)
\end{align*}

The predictive distribution for a future observation $x_{\text{new}}|x_{1:n}$ is

\begin{align*}
x_{\text{new}}|x_{1:n} 
 &\sim
 \text{Bb}(\tau^{*}_{1}+1,\tau^{*}_{0}-\tau^{*}_{1}+1,1) 
\end{align*}

Where

\begin{align*}
 \tau^{*}_{0} 
 &=
 n+\tau_{0} \\
 \tau^{*}_{1}
 &=
 \sum_{i=1}^{n}x_{i}+\tau_{1} 
\end{align*}



### ***Parametric inferece***

```{r,echo=FALSE}
renderPlot({
   plot_Bernoulli2_param(theta=input$Bern2_theta, 
                  n=input$Bern2_n, 
                  tau_0=input$Bern2_tau[2], 
                  tau_1=input$Bern2_tau[1]
                  )
  })
```

### ***Predictive inference***

```{r,echo=FALSE}
renderPlot({
   plot_Bernoulli2_pred(theta=input$Bern2_theta, 
                  n=input$Bern2_n,
                  tau_0=input$Bern2_tau[2], 
                  tau_1=input$Bern2_tau[1]
                  )
  })
```











# Normal  model


```{r,echo=FALSE}

NormalModel_SufStat<-function(x,n) {
  xx <- mean(x[1:n])
  yy <- mean((x[1:n]-xx)^2)
 sufstat <- c(xx,yy) ; 
}

NormalModel_PDF_meanvar_prior<-function(mu,vu,mu0,n0,a0,b0) {
  vu0 <- vu/n0 ;
  pdf <- dnorm( mu, mean=mu0, sd=sqrt(vu0)) * dinvgamma(vu, a0, b0) ;
   return (pdf)  
}

NormalModel_PDF_meanvar_post<-function(mu,vu,n,sufstat,mu0,n0,a0,b0) {
  
      if (n==0) {
    pdf <- NormalModel_PDF_meanvar_prior(mu,vu,mu0,n0,a0,b0) ;
    return (pdf) ;
    }
  
  x_mean <- sufstat[1] ;
  
  x_var <- sufstat[2] ;
  
  n_hat <- n0+n ;
  
  a_hat <-a0+0.5*n ;
  
  b_hat <- b0+0.5*n*x_var+0.5*n0*n*((mu0-x_mean)^2)/(n0+n) ;
  
  mu_hat <- (n0*mu0+n*x_mean)/(n0+n) ;
  
  vu_hat <- vu/n_hat
  
  pdf <- dnorm( mu, mean=mu_hat, sd=sqrt(vu_hat) )*dinvgamma(vu,a_hat,b_hat) ;
   return (pdf)  
}

NormalModel_PDF_var_prior<-function(vu,a0,b0) {
  pdf <- dinvgamma(vu,a0,b0) ;
   return (pdf)  
}

NormalModel_PDF_var_post<-function(vu,n,sufstat,mu0,n0,a0,b0) {
  
    if (n==0) {
    pdf <- NormalModel_PDF_var_prior(vu,a0,b0) ;
    return (pdf) ;
    }
  
  x_mean <- sufstat[1] ;
  
  x_var <- sufstat[2] ;
  
  a_hat <-a0+0.5*n ;
  
  b_hat <- b0+0.5*n*x_var+0.5*n0*n*((mu0-x_mean)^2)/(n0+n) ;
  
  pdf <- dinvgamma( vu, a_hat, b_hat) ;
  
  return (pdf)  
}

NormalModel_PDF_mean_prior<-function(mu,mu0,n0,a0,b0) {
  pdf <- StudentT_pdf(mu, mu0, n0*a0/b0, 2*a0) ;
   return (pdf)  
}

NormalModel_PDF_mean_post<-function(mu,n,sufstat,mu0,n0,a0,b0) {
  
  if (n==0) {
    pdf <- NormalModel_PDF_mean_prior(mu,mu0,n0,a0,b0) ;
    return (pdf) ;
    } 
  
  x_mean <- sufstat[1] ;
  
  x_var <- sufstat[2] ;
  
  n_hat <- n0+n ;
  
  a_hat <-a0+0.5*n ;
  
  b_hat <- b0+0.5*n*x_var+0.5*n0*n*((mu0-x_mean)^2)/(n0+n) ;
  
  mu_hat <- (n0*mu0+n*x_mean)/(n0+n) ;
  
  pdf <- StudentT_pdf(mu, mu_hat, (n+n0)*a_hat/b_hat, 2*a_hat) ;
   return (pdf)  
}

NormalModel_PDF_pred<-function(x,n,sufstat,mu0,n0,a0,b0) {
  
    if (n==0) {
      pdf <- StudentT_pdf(x, mu0, n0/(n0+1)*a0/b0, 2*a0) ;    
      return (pdf) ;
    } 
  
  x_mean <- sufstat[1] ;
  
  x_var <- sufstat[2] ;
  
  n_hat <- n0+n ;
  
  a_hat <-a0+0.5*n ;
  
  b_hat <- b0+0.5*n*x_var+0.5*n0*n*((mu0-x_mean)^2)/(n0+n) ;
  
  mu_hat <- (n0*mu0+n*x_mean)/(n0+n) ;
  
  pdf <- StudentT_pdf(x, mu_hat, (n+n0)/(n+n0+1)*a_hat/b_hat, 2*a_hat) ;
   return (pdf)  
}

plot_Normal_param <- function(mu=0.0,
                        vu=1.0,
                        n=10,
                        mu0=0.0,
                        n0=1.0,
                        a0=1.5,
                        b0=1.5) {
 # mu=0.0;
 # vu=1.0;
 # n=10;
 # mu0=0.0;
 # n0=1.0;
 # a0=1.5;
 # b0=1.5;
  
  fntsz <- 1.5;
  op <- par(cex = fntsz)
  
  obs <-get_randnumb(n) ;
  obs <- mu+sqrt(vu)*qnorm(obs,mean=0.0,sd=1.0) ;
  sufstat <- NormalModel_SufStat(obs,n) ;

  x_vu <- seq(from = 0.00, to = 10, length.out = 50) ;
  vu_prior_pdf <- NormalModel_PDF_var_prior(x_vu,a0,b0) ;
  vu_post_pdf <- NormalModel_PDF_var_post(x_vu,n,sufstat,mu0,n0,a0,b0) ;
  
  x_mu <- seq(from = -6, to = 6, length.out = 50) ;
  mu_prior_pdf <- NormalModel_PDF_mean_prior(x_mu,mu0,n0,a0,b0) ;
  mu_post_pdf <- NormalModel_PDF_mean_post(x_mu,n,sufstat,mu0,n0,a0,b0) ;
  
  
  muvu_prior_pdf <- matrix(rep(NaN,50*50),50,50) ;
  muvu_post_pdf <- muvu_prior_pdf ;
  for (i in 1:50)
    for (j in 1:50) {
      muvu_prior_pdf[i,j] <- NormalModel_PDF_meanvar_prior(x_mu[i],x_vu[j],mu0,n0,a0,b0) ;
      muvu_post_pdf[i,j] <- NormalModel_PDF_meanvar_post(x_mu[i],x_vu[j],n,sufstat,mu0,n0,a0,b0) ;
    }
  
  x_pred <- seq(from = -6, to = 6, length.out = 50) ;
  xnew_pred_pdf <- NormalModel_PDF_pred(x_pred,n,sufstat,mu0,n0,a0,b0) ;
  

  #layout( cbind( matrix(1,4,4), matrix(2,4,4), matrix(3,4,4) , matrix(4,4,4) ) )

  
  par(mfrow = c(2 , 2) )
  
  # about mu
  
       y_max <- c(mu_post_pdf, mu_prior_pdf) ;
  y_max <- y_max[is.finite(y_max)]
    y_max <- max(y_max)
  
  plot(x_mu, 
       mu_post_pdf, 
       type = "l",  
       main='Marginal Posterior vs Prior PDF', 
       xlab = expression(mu),
       ylab = expression(paste(pi,"(",mu,"|",x[1:n],")", " and " ,pi,"(",mu,")")), 
       col="blue" ,
       ylim = c(0,y_max+0.3) ,
      cex.lab=fntsz, 
      cex.axis=fntsz, 
      cex.main=fntsz, 
      cex.sub=fntsz) ;
  
  lines(x_mu, 
       mu_prior_pdf, 
       type = "l",  
       xlab = expression(mu),
       ylab = expression(paste(pi,"(",mu,"|",x[1:n],")", " and " ,pi,"(",mu,")")), 
       col="green" ,
      cex.lab=fntsz, 
      cex.axis=fntsz, 
      cex.main=fntsz, 
      cex.sub=fntsz) ;
  
  abline(v=mu, 
         col="purple")
    
  legend('topright',
        c('Posterior','Prior',expression(paste(mu, " real value"))),
        bg="transparent",
        lty=c(1,1),
        lwd=c(2.5,2.5),
        col=c('blue','green','purple'),
        cex=fntsz
    )

  # about sigma
  
  
     y_max <- c(vu_post_pdf, vu_prior_pdf) ;
  y_max <- y_max[is.finite(y_max)]
    y_max <- max(y_max)

  
  plot(x_vu, 
       vu_post_pdf, 
       type = "l",  
       main='Marginal Posterior vs Prior PDF', 
       xlab = expression(sigma^2),
       ylab = expression(paste(pi,"(",sigma^2,"|",x[1:n],")", " and ",pi,"(",sigma^2,")")), 
       col="blue" ,
       ylim = c(0,y_max+0.3) ,
      cex.lab=fntsz, 
      cex.axis=fntsz, 
      cex.main=fntsz, 
      cex.sub=fntsz) ;
  
  lines(x_vu, 
       vu_prior_pdf, 
       type = "l",  
       col="green" ,
      cex.lab=fntsz, 
      cex.axis=fntsz, 
      cex.main=fntsz, 
      cex.sub=fntsz) ;
  
  abline(v=vu, 
         col="purple")
    
  legend('topright',
        c('Posterior','Prior',expression(paste(sigma^2, " real value"))),
        bg="transparent",
        lty=c(1,1),
        lwd=c(2.5,2.5),
        col=c('blue','green','purple'),
        cex=fntsz
    )
 
  
  # about joint prior

  contour(x = x_mu,
         y = x_vu,
         z = muvu_prior_pdf,
         col="green",
         plot.title = title(main = "Joint Prior PDF",
                            xlab = expression(mu),
                            ylab = expression(sigma^2),
                            cex.lab=fntsz, 
                            cex.axis=fntsz, 
                            cex.main=fntsz, 
                            cex.sub=fntsz
                            ),
        cex.lab=fntsz, 
        cex.axis=fntsz, 
        cex.main=fntsz, 
        cex.sub=fntsz,
        labcex=fntsz
        ) ;
  
  
  contour(x = x_mu,
         y = x_vu,
         z = muvu_post_pdf,
         col="blue",
         plot.title = title(main = "Joint Posterior PDF",
                            xlab = expression(mu),
                            ylab = expression(sigma^2),
                            cex.lab=fntsz, 
                            cex.axis=fntsz, 
                            cex.main=fntsz, 
                            cex.sub=fntsz
                            ),
        cex.lab=fntsz, 
        cex.axis=fntsz, 
        cex.main=fntsz, 
        cex.sub=fntsz,
        labcex=fntsz
        ) ;
  points(c(mu),c(vu),pch = "+", col="purple")
  
    legend('topright',
        c('PDF',expression(paste(theta, "=", "(", mu,sigma^2, ")"))),
        bg="transparent",
        lty=c(1,NA),
        pch=c(NA,"+"),
        lwd=c(2.5,2.5),
        col=c('blue','purple'),
        cex=fntsz
    )

}


plot_Normal_pred <- function(mu=0.0,
                        vu=1.0,
                        n=10,
                        mu0=0.0,
                        n0=1.0,
                        a0=1.5,
                        b0=1.5) {
#  mu=0.0;
#  vu=1.0;
#  n=10;
#  mu0=0.0;
#  n0=1.0;
#  a0=1.5;
#  b0=1.5;
  
  fntsz <- 1.5;
  op <- par(cex = fntsz)
  
  obs <-get_randnumb(n) ;
  obs <- mu+sqrt(vu)*qnorm(obs,mean=0.0,sd=1.0) ;
  sufstat <- NormalModel_SufStat(obs,n) ;

  x_vu <- seq(from = 0.00, to = 10, length.out = 50) ;
  
  x_mu <- seq(from = -3, to = 3, length.out = 50) ;
  
  x_pred <- seq(from = -6, to = 6, length.out = 50) ;
  xnew_pred_pdf <- NormalModel_PDF_pred(x_pred,n,sufstat,mu0,n0,a0,b0) ;
  
  par(mfrow = c(1,1) )
  
  # about xnew_pred_pdf
  
  plot(x_pred, 
       xnew_pred_pdf, 
       type = "l",  
       main='Predictive PDF', 
       xlab = expression( x[n+1] ),
       ylab = expression(paste(p(x[n+1], "|",x[1:n]))), 
       col="blue" ,
      cex.lab=fntsz, 
      cex.axis=fntsz, 
      cex.main=fntsz, 
      cex.sub=fntsz) ;
  }


```


Inputs {.sidebar} 
-----------------------

***Parametric model parameters***

```{r,echo=FALSE}
sliderInput("Normal_n",
                 "$$n:$$",
                  min = 0,
                  max = 100,
                  step = 2,
                  value = 5.0)

sliderInput("Normal_mu",
                 "$$\\mu:$$",
                  min = -3.0,
                  max = 3.0,
                  step = 0.05,
                  value = 0.0)

sliderInput("Normal_vu",
                 "$$\\sigma^2:$$",
                  min = 0.01,
                  max = 3.0,
                  step = 0.05,
                  value = 0.1)
```

***Prior hyper-parameters***

```{r,echo=FALSE}

sliderInput("Normal_mu0",
                 "$$\\mu_0:$$",
                  min = -3.0,
                  max = 3.0,
                  step = 0.05,
                  value = 0.0)

sliderInput("Normal_n0",
                 "$$n_0:$$",
                  min = 1.0,
                  max = 100.0,
                  step = 2,
                  value = 1)

sliderInput("Normal_a0",
                 "$$a:$$",
                  min = 0.0,
                  max = 5.0,
                  step = 0.05,
                  value = 1)

sliderInput("Normal_b0",
                 "$$b:$$",
                  min = 0.0,
                  max = 10.0,
                  step = 0.05,
                  value = 1)
```


Column {.tabset}
-----------------------


### ***Description*** 

We consider the Bayesian model

\[
\begin{cases}
x_{i}|\mu,\sigma^2 
& \overset{\text{iid}}{\sim}\text{N}(\mu,\sigma^{2}),\,\,\forall i=1:n\\
\mu|\sigma^2 
&\sim \text{N}(\mu_{0},\frac{\lambda_0}{\sigma^{2}}) \\
\sigma^2 
& \sim \text{IG}(a_0,b_0)
\end{cases}
\]

The joint posterior distribution for $(\mu,\sigma^2)|x_{1:n}$ is: 
\begin{align*}
\begin{cases}
\mu|x_{1:n},\sigma^2 
&\sim \text{N}(\mu^{*},\frac{\lambda^{*}}{\sigma^{2}}) \\
\sigma^2|x_{1:n} 
& \sim \text{IG}(a^{*},b^{*})
\end{cases}
\end{align*}

The marginal posterior distribution for $\mu|x_{1:n}$ is: 
\begin{align*}
\mu|x_{1:n} 
&\sim
\text{St}(\mu^{*},\frac{\lambda^{*}a^{*}}{b^{*}},2a^{*})
\end{align*}

The marginal posterior distribution for $\sigma^2|x_{1:n}$ is: 
\begin{align*}
\sigma^2|x_{1:n} 
& \sim 
\text{IG}(a^{*},b^{*})
\end{align*}

The predictive distribution for a future observation $x_{\text{new}}|x_{1:n}$ is

\begin{align*}
x_{\text{new}}|x_{1:n} 
&\sim
\text{St}(\mu^{*},\frac{\lambda^{*}a^{*}}{(\lambda^{*}+1)b^{*}},2a^{*})
\end{align*}

Where:

\begin{align*}
\mu^{*} 
&= 
\frac{\lambda_{0}\mu_{0}+n\bar{x}}{\lambda_{0}+n} \\
\lambda^{*} 
&=
\lambda_{0}+n \\
a^{*}
&=
a+\frac{1}{2}n \\
b^{*}
&=
b
+\frac{1}{2}ns^{2}
+\frac{1}{2}\frac{\lambda_{0}n(\mu_{0}-\bar{x})^{2}}{\lambda_{0}+n} 
\end{align*}


### ***Parametric inference***

```{r,echo=FALSE}
renderPlot({
  plot_Normal_param(mu=input$Normal_mu,
              vu=input$Normal_vu,
              n=input$Normal_n,
              mu0=input$Normal_mu0,
              n0=input$Normal_n0,
              a0=input$Normal_a0,
              b0=input$Normal_b0)
  })
```

### ***Predictive inference***

```{r,echo=FALSE}
renderPlot({
  plot_Normal_pred(mu=input$Normal_mu,
              vu=input$Normal_vu,
              n=input$Normal_n,
              mu0=input$Normal_mu0,
              n0=input$Normal_n0,
              a0=input$Normal_a0,
              b0=input$Normal_b0)
  })
```







<!-- # Normal  model (parametrisation 2) -->


<!-- ```{r,echo=FALSE} -->

<!-- plot_Normal2_param <- function(mu=0.0, -->
<!--                         vu=1.0, -->
<!--                         n=10, -->
<!--                         tau0=1.0, -->
<!--                         tau1=0.0, -->
<!--                         tau2=2) { -->
<!--  # mu=0.0; -->
<!--  # vu=1.0; -->
<!--  # n=10; -->
<!--  # mu0=0.0; -->
<!--  # n0=1.0; -->
<!--  # a0=1.5; -->
<!--  # b0=1.5; -->

<!--   mu0 <- tau1/tau0 ; -->
<!--   n0 <- tau0 ; -->
<!--   a0 <- 0.5*(tau0+1); -->
<!--   b0 <- 0.5*tau2- tau1*tau1/tau0 ; -->

<!--  plot_Normal_param(mu,vu,n,mu0,n0,a0,b0) -->

<!-- } -->


<!-- plot_Normal2_pred <- function(mu=0.0, -->
<!--                         vu=1.0, -->
<!--                         n=10, -->
<!--                         tau0=1.0, -->
<!--                         tau1=0.0, -->
<!--                         tau2=2) { -->
<!--   mu0 <- tau1/tau0 ; -->
<!--   n0 <- tau0 ; -->
<!--   a0 <- 0.5*(tau0+1); -->
<!--   b0 <- 0.5*tau2- tau1*tau1/tau0 ; -->
<!--   plot_Normal_pred(mu,vu,n,mu0,n0,a0,b0) -->
<!--   } -->


<!-- ``` -->


<!-- Inputs {.sidebar}  -->
<!-- ----------------------- -->

<!-- ***Parametric model parameters*** -->

<!-- ```{r,echo=FALSE} -->
<!-- sliderInput("Normal2_n", -->
<!--                  "$$n:$$", -->
<!--                   min = 0, -->
<!--                   max = 100, -->
<!--                   step = 2, -->
<!--                   value = 5.0) -->

<!-- sliderInput("Normal2_mu", -->
<!--                  "$$\\mu:$$", -->
<!--                   min = -3.0, -->
<!--                   max = 3.0, -->
<!--                   step = 0.05, -->
<!--                   value = 0.0) -->

<!-- sliderInput("Normal2_vu", -->
<!--                  "$$\\sigma^2:$$", -->
<!--                   min = 0.01, -->
<!--                   max = 3.0, -->
<!--                   step = 0.05, -->
<!--                   value = 0.1) -->
<!-- ``` -->

<!-- ***Prior hyper-parameters*** -->

<!-- Note that it should be $2\tau_1^2<\tau_2+\tau_0$. -->

<!-- ```{r,echo=FALSE} -->

<!-- sliderInput("Normal2_tau0", -->
<!--                  "$$\\tau_0:$$", -->
<!--                   min = 1.0, -->
<!--                   max = 100.0, -->
<!--                   step = 0.05, -->
<!--                   value = 1.0) -->

<!-- sliderInput("Normal2_tau1", -->
<!--                  "$$\\tau_1:$$", -->
<!--                   min = -300, -->
<!--                   max = 300.0, -->
<!--                   step = 2, -->
<!--                   value = 0.0) -->

<!-- sliderInput("Normal2_tau2", -->
<!--                  "$$\\tau_2:$$", -->
<!--                   min = 0.0, -->
<!--                   max = 150.0, -->
<!--                   step = 0.1, -->
<!--                   value = 100) -->
<!-- ``` -->


<!-- Column {.tabset} -->
<!-- ----------------------- -->


<!-- ### ***Description***  -->

<!-- We consider the Bayesian model -->

<!-- \[ -->
<!-- \begin{cases} -->
<!-- x_{i}|\mu,\sigma^2  -->
<!-- & \overset{\text{iid}}{\sim}\text{N}(\mu,\sigma^{2}),\,\,\forall i=1:n\\ -->
<!-- \mu|\sigma^2  -->
<!-- &\sim \text{N}(\mu_{0},\frac{\lambda_0}{\sigma^{2}}) \\ -->
<!-- \sigma^2  -->
<!-- & \sim \text{IG}(a_0,b_0) -->
<!-- \end{cases} -->
<!-- \] -->

<!-- The joint posterior distribution for $(\mu,\sigma^2)|x_{1:n}$ is:  -->
<!-- \begin{align*} -->
<!-- \begin{cases} -->
<!-- \mu|x_{1:n},\sigma^2  -->
<!-- &\sim \text{N}(\mu^{*},\frac{\lambda^{*}}{\sigma^{2}}) \\ -->
<!-- \sigma^2|x_{1:n}  -->
<!-- & \sim \text{IG}(a^{*},b^{*}) -->
<!-- \end{cases} -->
<!-- \end{align*} -->

<!-- The marginal posterior distribution for $\mu|x_{1:n}$ is:  -->
<!-- \begin{align*} -->
<!-- \mu|x_{1:n}  -->
<!-- &\sim -->
<!-- \text{St}(\mu^{*},\frac{\lambda^{*}a^{*}}{b^{*}},2a^{*}) -->
<!-- \end{align*} -->

<!-- The marginal posterior distribution for $\sigma^2|x_{1:n}$ is:  -->
<!-- \begin{align*} -->
<!-- \sigma^2|x_{1:n}  -->
<!-- & \sim  -->
<!-- \text{IG}(a^{*},b^{*}) -->
<!-- \end{align*} -->

<!-- The predictive distribution for a future observation $x_{\text{new}}|x_{1:n}$ is -->

<!-- \begin{align*} -->
<!-- x_{\text{new}}|x_{1:n}  -->
<!-- &\sim -->
<!-- \text{St}(\mu^{*},\frac{\lambda^{*}a^{*}}{(\lambda^{*}+1)b^{*}},2a^{*}) -->
<!-- \end{align*} -->

<!-- Where: -->

<!-- \begin{align*} -->
<!-- \mu^{*}  -->
<!-- &=  -->
<!-- \frac{\lambda_{0}\mu_{0}+n\bar{x}}{\lambda_{0}+n} \\ -->
<!-- \lambda^{*}  -->
<!-- &= -->
<!-- \lambda_{0}+n \\ -->
<!-- a^{*} -->
<!-- &= -->
<!-- a+\frac{1}{2}n \\ -->
<!-- b^{*} -->
<!-- &= -->
<!-- b -->
<!-- +\frac{1}{2}ns^{2} -->
<!-- +\frac{1}{2}\frac{\lambda_{0}n(\mu_{0}-\bar{x})^{2}}{\lambda_{0}+n}  -->
<!-- \end{align*} -->


<!-- ### ***Parametric inference*** -->

<!-- ```{r,echo=FALSE} -->
<!-- renderPlot({ -->
<!--   plot_Normal2_param(mu=input$Normal2_mu, -->
<!--               vu=input$Normal2_vu, -->
<!--               n=input$Normal2_n, -->
<!--             tau0=input$Normal2_tau0, -->
<!--               tau1=input$Normal2_tau1, -->
<!--               tau2=input$Normal2_tau2) -->
<!--   }) -->
<!-- ``` -->

<!-- ### ***Predictive inference*** -->

<!-- ```{r,echo=FALSE} -->
<!-- renderPlot({ -->
<!--   plot_Normal2_pred(mu=input$Normal2_mu, -->
<!--               vu=input$Normal2_vu, -->
<!--               n=input$Normal2_n, -->
<!--             tau0=input$Normal2_tau0, -->
<!--               tau1=input$Normal2_tau1, -->
<!--               tau2=input$Normal2_tau2) -->
<!--   }) -->
<!-- ``` -->





