<!-- -------------------------------------------------------------------------------- -->

<!-- Copyright 2019 Georgios Karagiannis -->

<!-- georgios.karagiannis@durham.ac.uk -->
<!-- Assistant Professor -->
<!-- Department of Mathematical Sciences, Durham University, Durham,  UK  -->

<!-- This file is part of Bayesian_Statistics (MATH3341/4031 Bayesian Statistics III/IV) -->
<!-- which is the material of the course (MATH3341/4031 Bayesian Statistics III/IV) -->
<!-- taught by Georgios P. Katagiannis in the Department of Mathematical Sciences   -->
<!-- in the University of Durham  in Michaelmas term in 2019 -->

<!-- Bayesian_Statistics is free software: you can redistribute it and/or modify -->
<!-- it under the terms of the GNU General Public License as published by -->
<!-- the Free Software Foundation version 3 of the License. -->

<!-- Bayesian_Statistics is distributed in the hope that it will be useful, -->
<!-- but WITHOUT ANY WARRANTY; without even the implied warranty of -->
<!-- MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the -->
<!-- GNU General Public License for more details. -->

<!-- You should have received a copy of the GNU General Public License -->
<!-- along with Bayesian_Statistics  If not, see <http://www.gnu.org/licenses/>. -->

<!-- -------------------------------------------------------------------------------- -->

---
title: "Biostatistics (Bayes theorem)"
# output: html_document
output: flexdashboard::flex_dashboard
runtime: shiny
---


```{r}
library(ggplot2)
library(shiny)
```

```{r,echo=FALSE}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```


# Sensitivity \& Specitivity

```{r,echo=FALSE}
plot_posterior_SS <- function(Pr_prior_Dplus = 1.0/270.0, 
                           Pr_Tplus_Dplus = 0.89, 
                           Pr_Tminus_Dminus = 0.75) {
  
  
  # Prior
  #Pr_prior_Dplus <- 1.0/270.0
  Pr_prior_Dminus <- 1.0 - Pr_prior_Dplus
  
  # Sensitivity
  #Pr_Tplus_Dplus <- 0.89 ;
  
  # Specificity
  #Pr_Tminus_Dminus <- 0.75 ;
  
  # False positive
  Pr_Tminus_Dplus <- 1.0 -Pr_Tplus_Dplus  ;
  
  #False negative
  Pr_Tplus_Dminus <- 1.0 -Pr_Tminus_Dminus  ;
  
  
  Pr_post_Dplus_Tplus <- (Pr_Tplus_Dplus*Pr_prior_Dplus)/(Pr_Tplus_Dplus*Pr_prior_Dplus
                                                          +Pr_Tplus_Dminus*Pr_prior_Dminus) ;
  Pr_post_Dminus_Tplus <- 1.0 - Pr_post_Dplus_Tplus ;
  
  Pr_post_Dplus_Tminus <- (Pr_Tminus_Dplus*Pr_prior_Dplus)/(Pr_Tminus_Dplus*Pr_prior_Dplus
                                                            + Pr_Tminus_Dminus*Pr_prior_Dminus) ;
  Pr_post_Dminus_Tminus <- 1.0 - Pr_post_Dplus_Tminus ;
  
  data_Tplus <- data.frame( Disease = c("D+","D-"),
                  Pr = c(Pr_post_Dplus_Tplus, Pr_post_Dminus_Tplus) )
  
  plt1 <- ggplot(data_Tplus, aes(factor(Disease), Pr) )+ 
            geom_bar(stat="identity", position = "dodge") + 
            ggtitle("Given test result: T+") +
            xlab( "Disease" ) +
            ylab("Post. Prob. Pr(.|T+)")+
            ylim(0,1)+
            theme(text = element_text(size=20)) 
  
  data_Tminus <- data.frame(Disease = c("D+","D-"),
                Pr = c(Pr_post_Dplus_Tminus, Pr_post_Dminus_Tminus) )
  
  plt2 <- ggplot(data_Tminus, aes(factor(Disease), Pr)) + 
            geom_bar(stat="identity", position = "dodge") + 
            ggtitle("Given test result: T-") +
            xlab( "Disease" ) +
            ylab("Post. Prob. Pr(.|T-)")+
            ylim(0,1)+
            theme(text = element_text(size=20)) 
  
    multiplot(plt1, plt2, cols=2)
  
}
```

Column {.sidebar}
-----------------------------------------------------------------------

***Notation:*** 

Has the disease?
$$
D =
\begin{cases}
D+ & \text{, if yes} \\
D- & \text{, if no} 
\end{cases}
$$ 

Test result ?
$$
T =
\begin{cases}
T+ & \text{, if positive } \\
T- & \text{, if negative } 
\end{cases}
$$

***Probabilities***


```{r,echo=FALSE}
      sliderInput("Pr_prior_Dplus_SS",
                  "$$\\text{Prior } \\Pr(D+):$$",
                  min = 0.0,
                  max = 1.0,
                  value = 1.0/2.0)
      sliderInput("Pr_Tplus_Dplus_SS",
                  "$$\\text{Sensitivity } \\Pr(T+|D+):$$",
                  min = 0.0,
                  max = 1.0,
                  value = 0.8)
      sliderInput("Pr_Tminus_Dminus_SS",
                   "$$\\text{Specificity } \\Pr(T-|D-):$$",
                  min = 0.0,
                  max = 1.0,
                  value = 0.7) 
```

Column 
-----------------------------------------------------------------------

### ***Posterior distribution*** {data-height=700}

```{r,echo=FALSE}
renderPlot({
    plot_posterior_SS(Pr_prior_Dplus=input$Pr_prior_Dplus_SS, 
                   Pr_Tplus_Dplus=input$Pr_Tplus_Dplus_SS, 
                   Pr_Tminus_Dminus=input$Pr_Tminus_Dminus_SS)
  })
```

### ***Bayes formula*** {data-height=300}

<!-- ***Symbols:***  -->
<!-- $$ -->
<!-- D = -->
<!-- \begin{cases} -->
<!-- D+ & \text{, if diseased} \\ -->
<!-- D- & \text{, if not diseased}  -->
<!-- \end{cases} -->
<!-- $$  -->
<!-- $$ -->
<!-- T = -->
<!-- \begin{cases} -->
<!-- T+ & \text{, if positive test result} \\ -->
<!-- T- & \text{, if negative test result}  -->
<!-- \end{cases} -->
<!-- $$ -->

***Positive test result (T+):*** 
$$
\Pr(D|T+) = 
\begin{cases}
\frac{\Pr(T+|D+)\Pr(D+)}{\Pr(T+|D+)\Pr(D+)\ \ +\ \ \Pr(T+|D-)\Pr(D-)} & \text{, if } D=D+\\
\frac{\Pr(T+|D-)\Pr(D-)}{\Pr(T+|D+)\Pr(D+)\ \ +\ \ \Pr(T+|D-)\Pr(D-)} & \text{, if } D=D-
\end{cases}
$$

***Negative test result (T-):*** 
$$
\Pr(D|T-) = 
\begin{cases}
\frac{\Pr(T-|D+)\Pr(D+)}{\Pr(T-|D+)\Pr(D+)\ \ +\ \ \Pr(T-|D-)\Pr(D-)} & \text{, if } D=D+\\
\frac{\Pr(T-|D-)\Pr(D-)}{\Pr(T-|D+)\Pr(D+)\ \ +\ \ \Pr(T-|D-)\Pr(D-)} & \text{, if } D=D-
\end{cases}
$$




# False Negetive/Positive

```{r,echo=FALSE}
plot_posterior_FNP <- function(Pr_prior_Dplus = 1.0/270.0, 
                           Pr_Tminus_Dplus = 0.89, 
                           Pr_Tplus_Dminus = 0.75) {
  
  
  # Prior
  #Pr_prior_Dplus <- 1.0/270.0
  Pr_prior_Dminus <- 1.0 - Pr_prior_Dplus
  
  # Sensitivity
  Pr_Tplus_Dplus <- 1.0 - Pr_Tminus_Dplus ;
  
  # Specificity
  Pr_Tminus_Dminus <- 1.0 - Pr_Tplus_Dminus ;
  
  Pr_post_Dplus_Tplus <- (Pr_Tplus_Dplus*Pr_prior_Dplus)/(Pr_Tplus_Dplus*Pr_prior_Dplus
                                                          +Pr_Tplus_Dminus*Pr_prior_Dminus) ;
  Pr_post_Dminus_Tplus <- 1.0 - Pr_post_Dplus_Tplus ;
  
  Pr_post_Dplus_Tminus <- (Pr_Tminus_Dplus*Pr_prior_Dplus)/(Pr_Tminus_Dplus*Pr_prior_Dplus
                                                            + Pr_Tminus_Dminus*Pr_prior_Dminus) ;
  Pr_post_Dminus_Tminus <- 1.0 - Pr_post_Dplus_Tminus ;
  
  data_Tplus <- data.frame(
    Disease = c("D-","D+"),
    Pr = c(Pr_post_Dminus_Tplus, Pr_post_Dplus_Tplus) )
  
  plt1 <- ggplot(data_Tplus, aes(factor(Disease), Pr)) + 
    geom_bar(stat="identity", position = "dodge") + 
    ggtitle("Given test result: T+") +
    xlab( "Disease" ) +
    ylab("Post. Prob. Pr(.|T+)")+
    ylim(0,1)+
    theme(text = element_text(size=20)) 
  
  data_Tminus <- data.frame(
    Disease = c("D-","D+"),
    Pr = c(Pr_post_Dminus_Tminus, Pr_post_Dplus_Tminus) )
  
  plt2 <- ggplot(data_Tminus, aes(factor(Disease), Pr)) + 
    geom_bar(stat="identity", position = "dodge") + 
    ggtitle("Given test result: T-") +
    xlab( "Disease" ) +
    ylab("Post. Prob. Pr(.|T-)")+
    ylim(0,1)+
    theme(text = element_text(size=20)) 
  
  multiplot(plt1, plt2, cols=2)
  
}
```

Column {.sidebar}
-----------------------------------------------------------------------
***Notation:*** 

Has the disease?
$$
D =
\begin{cases}
D+ & \text{, if yes} \\
D- & \text{, if no} 
\end{cases}
$$ 

Test result ?
$$
T =
\begin{cases}
T+ & \text{, if positive } \\
T- & \text{, if negative } 
\end{cases}
$$

***Probabilities***


```{r,echo=FALSE}
      sliderInput("Pr_prior_Dplus_FNP",
                  "$$\\text{Prior } \\Pr(D+):$$",
                  min = 0.0,
                  max = 1.0,
                  value = 1.0/2.0)
      sliderInput("Pr_Tminus_Dplus_FNP",
                  "$$\\text{False negative } \\Pr(T-|D+):$$",
                  min = 0.0,
                  max = 1.0,
                  value = 0.2) 
      sliderInput("Pr_Tplus_Dminus_FNP",
                  "$$\\text{False positive } \\Pr(T+|D-):$$",
                  min = 0.0,
                  max = 1.0,
                  value = 0.3)
```

Column
-----------------------------------------------------------------------

### ***Posterior distribution*** {data-height=700}

```{r,echo=FALSE}
renderPlot({
    plot_posterior_FNP(Pr_prior_Dplus=input$Pr_prior_Dplus_FNP, 
                   Pr_Tminus_Dplus=input$Pr_Tminus_Dplus_FNP, 
                   Pr_Tplus_Dminus=input$Pr_Tplus_Dminus_FNP)
  })
```

### ***Bayes formula*** {data-height=300}

<!-- ***Symbols:***  -->
<!-- $$ -->
<!-- D = -->
<!-- \begin{cases} -->
<!-- D+ & \text{, if diseased} \\ -->
<!-- D- & \text{, if not diseased}  -->
<!-- \end{cases} -->
<!-- $$  -->
<!-- $$ -->
<!-- T = -->
<!-- \begin{cases} -->
<!-- T+ & \text{, if positive test result} \\ -->
<!-- T- & \text{, if negative test result}  -->
<!-- \end{cases} -->
<!-- $$ -->


***Positive test result (T+):*** 
$$
\Pr(D|T+) = 
\begin{cases}
\frac{\Pr(T+|D+)\Pr(D+)}{\Pr(T+|D+)\Pr(D+)\ \ +\ \ \Pr(T+|D-)\Pr(D-)} & \text{, if } D=D+\\
\frac{\Pr(T+|D-)\Pr(D-)}{\Pr(T+|D+)\Pr(D+)\ \ +\ \ \Pr(T+|D-)\Pr(D-)} & \text{, if } D=D-
\end{cases}
$$

***Negative test result (T-):*** 
$$
\Pr(D|T-) = 
\begin{cases}
\frac{\Pr(T-|D+)\Pr(D+)}{\Pr(T-|D+)\Pr(D+)\ \ +\ \ \Pr(T-|D-)\Pr(D-)} & \text{, if } D=D+\\
\frac{\Pr(T-|D-)\Pr(D-)}{\Pr(T-|D+)\Pr(D+)\ \ +\ \ \Pr(T-|D-)\Pr(D-)} & \text{, if } D=D-
\end{cases}
$$










